---
title: "Edits Notebook"
author: "Cassandra Etter"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the data behind the EDITS work. It follows four steps. For Step 1 I look at uptake rates between 2010 and 2021 in three countries: US, UK and Germany. For the US, I use the PSID which tracks households between 2011- 2021 on a biannual basis. For Germany, I use SOEP which tracks households between 2010- 2021 on an annual basis, and for the UK I use Understanding Society which tracks households between 2010 - 2021 on an annual basis. All data was downloaded in September 2024 and cleaning files are on github.

8 technologies selected are :Internet Access, own a smartphone, Home has Solar, Home switched from Gas/Coal/Wood to electric (fuel switching), Home renovation/ retrofitting, EV/ Hybrid car, Commuting habit (walking, biking), digital skills/ capabilities (daily use of internet).

Note: Significant limitations on technologies surveyed. With better data, I could also do this for smart-meters, air conditioners, etc.

```{r libraries,include=FALSE}
library(DiagrammeR)
library(psidread)
library(ggplot2)
library(dplyr)
library(tidyr)
library(survey)
library(priceR)
library(readr)
library(tidyverse)
library(nls2)
library(drc)
library(forecast)
library(nls2)
library(quantmod)
library(scales)
```


```{r load UK DEU, include=FALSE}

UK <- read_csv("~/Documents/Github/PhD/Data/UK_Clean.csv")
DEU <- read_csv("~/Documents/Github/PhD/Data/DEU_Clean.csv", show_col_types = FALSE)
US <- read_csv("~/Documents/Github/PhD/Data/US_EDITS.csv", show_col_types = FALSE)

```

## Step 1: Uptake Rates

This provides a snapshot for current uptake across the households assessing heterogeneity across income level.

### U.S. Uptake Rates

For the US, I have information for solar ownership, EV/ hybrid ownership, smartphone ownership, internet access, and commute habit. I create an additional variable on fuel switching by looking at heating method and seeing where moved from wood, oil, gas, coal to electric heating. I also create a variable for retrofitting based on home renovations and skills based on survey responses where RP responds they use the internet daily.

Note: Smartphone ownership and digital skills questions were only added to the survey in 2015.

```{r US Uptake Rates, echo=FALSE}
US <- US %>%
  group_by(pid) %>%
  arrange(pid, year) %>%
  mutate(
    # Create a combined solar indicator
    is_solar = heat_method == 6 | heat_method_2 == 6 | heat_method_3 == 6,
    
    # Track those who switched to electric
    switched_to_electric = lag(heat_method) %in% c(1,3,4,5,10,11) & 
                          heat_method == 2,
    
    # Track those always on electric
    always_electric = all(heat_method == 2, na.rm = TRUE),
    
    # Track those who switched to solar (using any of the three methods)
    switched_to_solar = lag(is_solar) == FALSE & is_solar == TRUE,
    
    # Track those always on solar
    always_solar = all(is_solar, na.rm = TRUE),
    
    # Combined measures: either switched or always been on that method
    using_electric = heat_method == 2,
    using_solar = is_solar,
    
    # Switch years
    switch_year_electric = case_when(
      switched_to_electric ~ year,
      TRUE ~ NA_real_
    ),
    switch_year_solar = case_when(
      switched_to_solar ~ year,
      TRUE ~ NA_real_
    )
  ) %>%
  ungroup() %>%
  filter(year > 2010)
# For decile analysis:
uptake_rates <- US %>%
  group_by(year, decile) %>%
  summarise(
    total_households = n_distinct(pid),
    
    # Electric heating metrics
    new_electric_switches = sum(switched_to_electric, na.rm = TRUE),
    total_electric_users = sum(using_electric, na.rm = TRUE),
    
    # Solar heating metrics
    new_solar_switches = sum(switched_to_solar, na.rm = TRUE),
    total_solar_users = sum(using_solar, na.rm = TRUE),
    
    # Calculate rates
    electric_switch_rate = (new_electric_switches / total_households) * 100,
    total_electric_rate = (total_electric_users / total_households) * 100,
    solar_switch_rate = (new_solar_switches / total_households) * 100,
    total_solar_rate = (total_solar_users / total_households) * 100
  ) %>%
  ungroup()

us_props <- US %>%
  group_by(year, decile) %>%
  summarise(
    n_total = n(),
    smartphone_prop = mean(smartphone == 1, na.rm = TRUE),
    hybrid_prop = mean(hybrid == 1 | hybrid_2 == 1 | hybrid_3 == 1, na.rm = TRUE), # includes EVs 
    solar_prop = mean(heat_method == 6 | heat_method_2 ==6 | heat_method_3 == 6, na.rm = TRUE),
    internet_prop = mean(computer == 1, na.rm = TRUE),     # Computer ownership
    repairs_prop = mean(hh_repair > 1, na.rm = TRUE), 
    skills_prop = mean(internet == 1, na.rm = TRUE),       # Internet usage
    fuel_switch_prop = mean(heat_method == 2 & 
                           !(heat_method_2 %in% c(1,3,4,5,10,11) | 
                             heat_method_3 %in% c(1,3,4,5,10,11)), na.rm = TRUE)
  ) %>%
  mutate(
    smartphone_percent = smartphone_prop * 100,            # Remove duplicate line
    hybrid_percent = hybrid_prop * 100,
    solar_percent = solar_prop * 100, 
    internet_percent = internet_prop * 100, 
    repairs_percent = repairs_prop * 100, 
    skills_percent = skills_prop * 100,
    fuel_switch_percent = fuel_switch_prop * 100 
  )


us_props_long <- us_props %>%
  dplyr::select(year, decile, smartphone_percent, hybrid_percent, solar_percent, 
         internet_percent, repairs_percent, skills_percent, 
         fuel_switch_percent) %>%
  pivot_longer(
    cols = c(smartphone_percent, hybrid_percent, solar_percent, 
             internet_percent, repairs_percent, skills_percent, 
             fuel_switch_percent),
    names_to = "technology",
    values_to = "percentage"
  ) %>%
  mutate(technology = factor(technology, 
    levels = c("smartphone_percent", "hybrid_percent", "solar_percent", 
               "internet_percent", "repairs_percent", "skills_percent", 
               "fuel_switch_percent"),
    labels = c("Smartphone", "Hybrid/EV", "Solar Panels", "Internet", 
               "Home Renovations", "Digital Skills", "Clean Heating Switch")))
# Create the plot
ggplot(us_props_long, aes(x = decile, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year) +
  theme_minimal() +
  labs(
    title = "Technology Adoption by Income Decile in US",
    x = "Income Decile",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  scale_x_continuous(breaks = 1:10) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2)) # Makes legend more compact

# Plot across time
ggplot(us_props_long, aes(x = year, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~decile) +
  theme_minimal() +
  labs(
    title = "Technology Adoption Over Time by Income Decile",
    x = "Year",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2))
```

Clean heating is interesting in that high income households had less electric only and marginally increased over 2011 - 2021 time period, though they still did increase within decile 9 and 10. Solar panel infiltration is lower than national average at 1% with EIA estimating 4%.

### U.K. Uptake Rates

For the UK, I have solar, EV/ hybrid, smartphone, internet, commute and skills. I create fuel switching in the same manner as for US, buy looking at home energy bills and looking for those households that switch from gas/coal to electric only.

Solar question is only asked in 2009, 2012, 2018, and 2021. EV is asked in 2012, 2015, 2018, and 2021. Smartphone begins in 2013. Skills starts in 2011.

```{r UK Uptake Rates, echo=FALSE}
# UK proportions
uk_props <- UK %>%
  group_by(year, weighted_decile) %>%
  summarise(
    n_total = n(),
    solar_prop = mean(solar1 %in% c("yes - fitted", "Yes - fitted"), na.rm = TRUE),
    ev_prop = mean(carfuel1 %in% c("Electric / battery", "electric / battery", "Hybrid (petrol/electric)", "hybrid (petrol/electric)"), na.rm = TRUE),
    smartphone_prop = mean(smartmob %in% c("Yes", "yes"), na.rm = TRUE ),
    skills_prop = mean(netpuse %in% c("Every day", "every day") | netpusenew %in% c("Every day", "every day"), na.rm = TRUE),
    internet_prop = mean(pcnet %in% c("yes", "Yes"), na.rm = TRUE),
    commute_prop = mean(wktrvfar %in% c( "Cycle", "Bus/coach", "Walk", "Train", "Underground/Metro/Tram/Light railway", "underground/metro/tram/light railway", "Motorcycle/moped/scooter") | worktrav %in% c("Underground/Metro/Tram/Light railway", "Train", "cycle", "Walk", "underground/metro/tram/light railway (if england/scotland/wales)", "bus/coach", "Bus/coach", "train", "underground/metro/tram/light railway", "underground/metro/tram/light railway {if region = gb}", "Motorcycle/moped/scooter") , na.rm = TRUE),
    fuel_switch_prop = mean(gaspay == "inapplicable" & duelpay == "inapplicable", na.rm = TRUE)
  ) %>%
  mutate(
    solar_percent = solar_prop * 100,
    ev_percent = ev_prop * 100,
    smartphone_percent = smartphone_prop * 100, 
    internet_percent = internet_prop * 100, 
    commute_percent = commute_prop * 100,
    skills_percent = skills_prop * 100, 
    fuel_switch_percent = fuel_switch_prop * 100
  )

# Convert to long format with all technologies
uk_props_long <- uk_props %>%
  dplyr::select(year, weighted_decile, solar_percent, ev_percent, smartphone_percent, 
         internet_percent, skills_percent, commute_percent, fuel_switch_percent) %>%
  pivot_longer(
    cols = c(solar_percent, ev_percent, smartphone_percent, 
             internet_percent, skills_percent, commute_percent, fuel_switch_percent),
    names_to = "technology",
    values_to = "percentage"
  ) %>%
  mutate(technology = factor(technology, 
    levels = c("ev_percent", "solar_percent", "smartphone_percent", 
               "internet_percent", "skills_percent", "commute_percent", "fuel_switch_percent"),
    labels = c("Electric Vehicle", "Solar Panels", "Smartphone", 
               "Internet", "Digital Skills", "Sustainable Transport", "Clean Heating Switch")))

ggplot(uk_props_long, aes(x = weighted_decile, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year) +
  theme_minimal() +
  labs(
    title = "Technology Adoption by Income Decile in UK",
    x = "Income Decile",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  scale_x_continuous(breaks = 1:10) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2)) # Makes legend more compact

ggplot(uk_props_long[uk_props_long$percentage != 0, ], aes(x = year, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~weighted_decile) +
  theme_minimal() +
  labs(
    title = "Technology Adoption Over Time by Income Decile UK",
    x = "Year",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2))

# Alternative visualization: Faceted plot for clearer viewing
ggplot(uk_props_long, aes(x = weighted_decile, y = percentage)) +
  geom_line(color = "grey70") +
  geom_point(aes(color = technology), size = 3) +
  facet_grid(technology ~ year) +
  theme_minimal() +
  labs(
    title = "Technology Adoption by Income Decile in UK",
    x = "Income Decile",
    y = "Percentage of Households (%)"
  ) +
  scale_x_continuous(breaks = seq(2, 10, 2)) +
  theme(
    legend.position = "none",
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold"),
    panel.spacing.y = unit(1, "lines")
  )

uk_props_long <- uk_props_long %>%
  rename(decile = weighted_decile)
```

### Germany Uptake Rates

For Germany I have information on solar, EV, retrofitting, smartphone, internet.

EV only exists in 2015 and 2020. (note it is full EV or biodiesel, no hybrid category) Retrofitting is asked from 2010 - 2015, and 2019 Smartphone

```{r DEU Uptake Rates, echo=FALSE}

deu_props <- DEU %>%
  group_by(syear, weighted_decile) %>%
  summarise(
    n_total = n(),
    hybrid_prop = sum(hli0121 == 1 | hli0123 == 1 | hli0124 == 1 | 
                     hli0114 == 1 | hli0115 == 1 | hli0116 == 1 | hli0117 == 1, 
                     na.rm = TRUE) / n(),
    solar_prop = sum(hgeqpsol == 1, na.rm = TRUE) / n(),
    renovation_prop = sum(hgcondit == 2 | hgcondit == 3, na.rm = TRUE) /n(),
    smartphone_prop = sum(hgeqptel == 1, na.rm = TRUE) / n(),
    internet_new = sum(hlf0169_h == 1 , na.rm = TRUE) / n()
  ) %>%
  group_by(weighted_decile) %>%
  arrange(syear) %>%
  mutate(
    internet_prop = pmin(cumsum(internet_new), 1),
    smartphone_prop = if_else(syear > 2015, 1, smartphone_prop)
  ) %>%
  mutate(
    hybrid_percent = hybrid_prop * 100,
    solar_percent = solar_prop * 100, 
    renovation_percent = renovation_prop * 100, 
    smartphone_percent = smartphone_prop * 100,
    internet_percent = internet_prop * 100
  )


deu_props_long <- deu_props %>%
  dplyr::select(syear, hybrid_percent, solar_percent, renovation_percent, smartphone_percent, internet_percent) %>%  
  pivot_longer(
    cols = c(hybrid_percent, solar_percent, renovation_percent, smartphone_percent, internet_percent),
    names_to = "technology",
    values_to = "percentage"
  ) %>%
  mutate(technology = factor(technology, 
    levels = c("hybrid_percent", "solar_percent", "renovation_percent", "smartphone_percent", "internet_percent"),
    labels = c("Hybrid/EV", "Solar Panels", "Home Renovations", "Owns a Smartphone", "Internet access")))

ggplot(deu_props_long, aes(x = weighted_decile, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~syear) +
  theme_minimal() +
  labs(
    title = "Technology Adoption by Income Decile in Germany",
    x = "Income Decile",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  scale_x_continuous(breaks = 1:10) +
  theme(legend.position = "bottom")

ggplot(deu_props_long[deu_props_long$percentage != 0, ], aes(x = syear, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~weighted_decile) +
  theme_minimal() +
  labs(
    title = "Technology Adoption Over Time by Income Decile Germany",
    x = "Year",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2))

deu_props_long <- deu_props_long %>%
  rename(year = syear) %>%
  rename(decile = weighted_decile)

```

### Combined Uptake Rates
Not all technologies in each country have the same year of start and end data. I now get data as close to 2015 start and 2021 end as possible. Look at year ranges available for uptake rates for each country and technology at decile level. 

```{r combined uptake rates, echo=FALSE}

# Calculate uptake rates for US
us_uptake_rates <- map_df(unique(us_props_long$technology), function(tech) {
  us_props_long %>%
    filter(
      technology == tech,
      year %in% c(2015, 2021)
    ) %>%
    dplyr::select(year, decile, technology, percentage) %>%
    pivot_wider(
      names_from = year,
      values_from = percentage
    ) %>%
    mutate(
      start_year = 2015,
      uptake_rate = !!sym(as.character(2021)) - !!sym(as.character(2015)),
      country = "US"
    )
})

# Calculate uptake rates for UK
uk_uptake_rates <- map_df(unique(uk_props_long$technology), function(tech) {
  # Set start year based on technology
  start_year <- if(tech == "Solar Panels") {
    2012
  } else {
    2015
  }
  
  # Set end year based on technology
  end_year <- if(tech == "Digital Skills") {
    2019
  } else {
    2021
  }
  
  uk_props_long %>%
    filter(
      technology == tech,
      year %in% c(start_year, end_year)
    ) %>%
    dplyr::select(year, decile, technology, percentage) %>%
    pivot_wider(
      names_from = year,
      values_from = percentage
    ) %>%
    mutate(
      start_year = start_year,
      end_year = end_year,
      uptake_rate = !!sym(as.character(end_year)) - !!sym(as.character(start_year)),
      country = "UK"
    )
})

deu_uptake_rates <- map_df(unique(deu_props_long$technology), function(tech) {
  # Set start year based on technology
  start_year <- if(tech == "Home Renovations") {
    2010
  } else {
    2015  
  }
  
  # Set end year based on technology
  end_year <- case_when(
    tech == "Home Renovations" ~ 2015, 
    tech == "Hybrid/EV" ~ 2020,
    TRUE ~ 2021
  )
  
  deu_props_long %>%
    filter(
      technology == tech,
      year %in% c(start_year, end_year)
    ) %>%
    dplyr::select(year, decile, technology, percentage) %>%
    pivot_wider(
      names_from = year,
      values_from = percentage
    ) %>%
    mutate(
      start_year = start_year,
      end_year = end_year,
      uptake_rate = !!sym(as.character(end_year)) - !!sym(as.character(start_year)),
      country = "Germany"
    )
})
# Combine all datasets
all_uptake_rates <- bind_rows(us_uptake_rates, uk_uptake_rates, deu_uptake_rates)

# Print summary of start years used
print("Start years used for each technology by country:")
all_uptake_rates %>%
  distinct(country, technology, start_year) %>%
  arrange(country, technology) %>%
  print(n = Inf)

# Create heatmap
heatmap_all <- ggplot(all_uptake_rates, 
  aes(x = factor(decile), 
      y = paste0(technology, " (", start_year, "-", ifelse(country == "Germany" & technology == "Home Renovation", "2019", "2021"), ")"), 
      fill = uptake_rate)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "red",
    mid = "white",
    high = "darkgreen",
    midpoint = 0,
    limits = c(-5, 5),
    oob = scales::squish
  ) +
  theme_minimal() +
  facet_wrap(~country) +
  labs(
    title = "Technology Uptake Rates by Income Decile",
    subtitle = "Percentage Point Change in Adoption (Capped at ±5 percentage points)",
    x = "Income Decile",
    y = NULL,
    fill = "Percentage\nPoint Change"
  ) +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(size = 10),
    axis.text.y = element_text(hjust = 1),
    legend.position = "right"
  )

# Create boxplot
boxplot_all <- ggplot(all_uptake_rates, 
  aes(y = paste0(technology, " (", start_year, "-", ifelse(country == "Germany" & technology == "Home Renovations", "2019", "2021"), ")"), 
      fill = country)) +
  geom_vline(xintercept = 0, color = "grey80", linetype = "dashed") +
  geom_boxplot(aes(x = uptake_rate), width = 0.7, alpha = 0.7, position = position_dodge(width = 0.8)) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal() +
  labs(
    title = "Technology Uptake Distribution by Type and Country",
    subtitle = "Boxplots show distribution across income deciles",
    x = "Percentage Point Change",
    y = NULL,
    fill = "Country"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    legend.position = "bottom"
  )

# Display plots
heatmap_all
boxplot_all
```

There are two outlier technologies for uptake rates. The first is Sustainable Transit in the UK. UK commute (sustainable transport) only shows commuting patterns for those who work, so potentially hides significant commuting preferences, particularly for those out of work or job seeking and is being compared against COVID so less people were commuting to work in 2021, no matter their method. The second is home renovations in Germany. If i change the year to later, the figure does go up, but 2010 shows a renovation boom that is lost in later years.

For fun, I'm also looking at US state variation, but this is for PhD, please ignore in meeting.

```{r US Regional Adoption, include=FALSE}
# Calculate state adoption rates
state_adoption <- US %>%
  group_by(year, hh_state) %>%
  summarise(
    n_total = n(),
    smartphone_prop = mean(smartphone == 1, na.rm = TRUE),
    hybrid_prop = mean(hybrid == 1, na.rm = TRUE),
    solar_prop = mean(heat_method == 6, na.rm = TRUE),
    internet_prop = mean(computer == 1, na.rm = TRUE),
    repairs_prop = mean(hh_repair > 1, na.rm = TRUE),
    skills_prop = mean(internet == 1, na.rm = TRUE),
    fuel_switch_prop = mean(switched_to_electric | switched_to_solar, na.rm = TRUE)
  ) %>%
  mutate(
    smartphone_percent = smartphone_prop * 100,
    hybrid_percent = hybrid_prop * 100,
    solar_percent = solar_prop * 100,
    internet_percent = internet_prop * 100,
    repairs_percent = repairs_prop * 100,
    skills_percent = skills_prop * 100,
    fuel_switch_percent = fuel_switch_prop * 100
  )

# Convert to long format
state_adoption_long <- state_adoption %>%
  pivot_longer(
    cols = ends_with("percent"),
    names_to = "technology",
    values_to = "percentage"
  ) %>%
  mutate(technology = factor(technology,
    levels = c("smartphone_percent", "hybrid_percent", "solar_percent",
               "internet_percent", "repairs_percent", "skills_percent",
               "fuel_switch_percent"),
    labels = c("Smartphone", "Hybrid/EV", "Solar Panels",
               "Internet", "Home Repairs", "Digital Skills",
               "Clean Heating Switch")
  ))

# Calculate state growth (2015-2020)
state_growth <- state_adoption_long %>%
  filter(year %in% c(2015, 2021)) %>%
  dplyr::select(year, hh_state, technology, percentage) %>%
  pivot_wider(
    names_from = year,
    values_from = percentage,
    names_prefix = "year_"
  ) %>%
  mutate(
    growth_rate = pmax(year_2021 - year_2015, 0)
  )

# Create heatmap of state growth (top 20 states by average growth)
top_states <- state_growth %>%
  group_by(hh_state) %>%
  summarise(mean_growth = mean(growth_rate, na.rm = TRUE)) %>%
  arrange(desc(mean_growth)) %>%
  head(20) %>%
  pull(hh_state)

ggplot(state_growth %>% filter(hh_state %in% top_states),
       aes(y = reorder(hh_state, growth_rate), x = technology, fill = growth_rate)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "darkgreen") +
  theme_minimal() +
  labs(
    title = "Technology Adoption Growth by State (2015-2020)",
    subtitle = "Top 20 states by average growth rate across technologies",
    x = "Technology",
    y = "State",
    fill = "Growth Rate (%)"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

# Create boxplot of state distributions by technology
ggplot(state_growth, aes(y = reorder(technology, growth_rate, FUN = median))) +
  geom_boxplot(aes(x = growth_rate), fill = "lightgreen", alpha = 0.5) +
  geom_jitter(aes(x = growth_rate), height = 0.2, alpha = 0.4) +
  theme_minimal() +
  labs(
    title = "Distribution of Technology Adoption Growth Across States (2015-2020)",
    subtitle = "Each point represents a state",
    x = "Percentage Point Growth",
    y = NULL
  )

# Calculate summary statistics for each state
state_summary <- state_growth %>%
  group_by(hh_state) %>%
  summarise(
    mean_growth = mean(growth_rate, na.rm = TRUE),
    max_growth = max(growth_rate, na.rm = TRUE),
    top_technology = technology[which.max(growth_rate)],
    n_technologies = sum(growth_rate > 0, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_growth))

print("Top 10 States by Average Growth Rate:")
print(head(state_summary, 10))

```
### CAGR

Two point estimate is not preferable so use logistic cagr. 
```{r CAGR, echo=FALSE}
calculate_adoption_rate <- function(data) {
  data %>%
    group_by(technology, decile) %>%
    # Remove NaN and zero values
    filter(!is.na(percentage) & percentage > 0) %>%
    # Get number of observations
    mutate(n_obs = n()) %>%
    arrange(year) %>%
    summarise(
      start_year = first(year),
      end_year = last(year),
      start_value = first(percentage),
      end_value = last(percentage),
      n_years = end_year - start_year,
      n_observations = n(),
      # Calculate growth rate based on data availability
      cagr = case_when(
        # If only one observation, use that observation as the rate
        n_observations == 1 ~ start_value,
        # If start and end values are the same, use that value
        start_value == end_value ~ start_value,
        # If we have less than 5 observations, use two most recent years
        n_observations < 5 ~ {
          ((end_value/start_value)^(1/n_years) - 1) * 100
        },
        # If more than 5 observations, calculate standard CAGR
        TRUE ~ ((end_value/start_value)^(1/n_years) - 1) * 100
      ),
      data_type = case_when(
        n_observations == 1 ~ "single_point",
        n_observations < 5 ~ "two_point",
        TRUE ~ "multi_point"
      ),
      .groups = 'drop'
    )
}

# Calculate CAGR
deu_cagr_results <- calculate_adoption_rate(deu_props_long) 
us_cagr_results <- calculate_adoption_rate(us_props_long)
uk_cagr_results <-calculate_adoption_rate(uk_props_long)


# Combine all CAGR results
all_cagr <- bind_rows(
  deu_cagr_results %>% mutate(country = "Germany"),
  us_cagr_results %>% mutate(country = "US"),
  uk_cagr_results %>% mutate(country = "UK")
) %>%
  mutate(technology = case_when(
    technology =="Electric Vehicle" ~ "Hybrid/EV",
    technology =="Internet" ~ "Internet access",
    technology == "Smartphone" ~ "Owns a Smartphone",
    TRUE ~ technology
  ))

# Create main plot comparing technologies across countries and deciles
ggplot(all_cagr, aes(x = factor(decile), y = cagr, color = country)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_line(aes(group = country)) +
  facet_wrap(~technology, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Technology Adoption CAGR by Country and Income Decile",
    x = "Income Decile",
    y = "CAGR (%)",
    color = "Country"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text = element_text(size = 9),
    strip.text = element_text(size = 10, face = "bold")
  )

# Create boxplot to show distribution of CAGR by technology and country
ggplot(all_cagr, aes(x = country, y = cagr, fill = country)) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(~technology, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Distribution of CAGR by Technology and Country",
    x = "",
    y = "CAGR (%)",
    fill = "Country"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 10, face = "bold")
  )

# Calculate summary statistics
cagr_summary <- all_cagr %>%
  group_by(country, technology) %>%
  summarise(
    mean_cagr = mean(cagr, na.rm = TRUE),
    median_cagr = median(cagr, na.rm = TRUE),
    sd_cagr = sd(cagr, na.rm = TRUE),
    min_cagr = min(cagr, na.rm = TRUE),
    max_cagr = max(cagr, na.rm = TRUE),
    .groups = 'drop'
  )

# Create a heatmap of mean CAGR
ggplot(cagr_summary, aes(x = country, y = technology, fill = mean_cagr)) +
  geom_tile() +
  scale_fill_gradient2(
    low = "red", 
    mid = "white",
    high = "blue",
    midpoint = 0,
    name = "Mean CAGR (%)"
  ) +
  theme_minimal() +
  labs(
    title = "Mean CAGR by Technology and Country",
    x = "",
    y = ""
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

# Print summary statistics
print("Summary Statistics of CAGR by Country and Technology")
print(cagr_summary %>% arrange(country, technology))
```

Log based growth

### Log Growth 
```{r log based growth rate, echo=FALSE}
calculate_adoption_rate <- function(data) {
  data %>%
    group_by(technology, decile) %>%
    # Remove NaN and zero values
    filter(!is.na(percentage) & percentage > 0) %>%
    # Get number of observations
    mutate(n_obs = n()) %>%
    arrange(year) %>%
    summarise(
      start_year = first(year),
      end_year = last(year),
      start_value = first(percentage),
      end_value = last(percentage),
      n_years = end_year - start_year,
      n_observations = n(),
      # Calculate log-based growth rate
      cagr = case_when(
        # If only one observation, use that observation as the rate
        n_observations == 1 ~ start_value,
        # If start and end values are the same, use that value
        start_value == end_value ~ start_value,
        # Use log-based calculation to capture positive growth
        TRUE ~ {
          # Calculate log change
          log_change <- (log(end_value/start_value) / n_years * 100)
          # If log change is negative, use a small positive value
          ifelse(log_change < 0, 0.1, log_change)
        }
      ),
      data_type = case_when(
        n_observations == 1 ~ "single_point",
        n_observations < 5 ~ "two_point",
        TRUE ~ "multi_point"
      ),
      .groups = 'drop'
    )
}

us_growth_results<- calculate_adoption_rate(us_props_long)
uk_growth_results<- calculate_adoption_rate(uk_props_long)
deu_growth_results<- calculate_adoption_rate(deu_props_long)

# Combine the results from different countries
combined_growth <- bind_rows(
  mutate(us_growth_results, country = "US"),
  mutate(uk_growth_results, country = "UK"),
  mutate(deu_growth_results, country = "Germany")
) %>%
    mutate(technology = case_when(
    technology =="Electric Vehicle" ~ "Hybrid/EV",
    technology =="Internet" ~ "Internet access",
    technology == "Smartphone" ~ "Owns a Smartphone",
    TRUE ~ technology
  ))


# Create a boxplot to show distribution of growth rates by technology and country
ggplot(combined_growth, aes(x = technology, y = cagr, fill = technology)) +
  geom_boxplot() +
  facet_wrap(~ country, scales = "free_y") +
  labs(
    title = "Technology Adoption Growth Rates by Country",
    x = "Technology",
    y = "Log-based Compound Annual Growth Rate (%)",
    fill = "Technology"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  scale_fill_brewer(palette = "Set3")

# Print summary statistics for each technology by country
summary_stats <- combined_growth %>%
  group_by(country, technology) %>%
  summarise(
    mean_growth = mean(cagr),
    median_growth = median(cagr),
    max_growth = max(cagr),
    min_growth = min(cagr)
  ) %>%
  arrange(country, desc(mean_growth))

print(summary_stats)

ggplot(combined_growth, aes(x = factor(decile), y = cagr, fill = technology)) +
 geom_boxplot() +
 facet_wrap(~ country, scales = "free_y") +
 labs(
   title = "Technology Adoption Growth Rates by Decile and Country",
   x = "Decile",
   y = "Log-based Compound Annual Growth Rate (%)",
   fill = "Technology"
 ) +
 theme_minimal() +
 theme(
   axis.text.x = element_text(angle = 0),
   legend.position = "bottom"
 ) +
 scale_fill_brewer(palette = "Set3")

summary_stats_decile <- combined_growth %>%
 group_by(country, technology, decile) %>%
 summarise(
   mean_growth = mean(cagr),
   median_growth = median(cagr),
   max_growth = max(cagr),
   min_growth = min(cagr)
 ) %>%
 arrange(country, technology, decile)

print(summary_stats_decile)

```


## Step 2: Baseline Model

In the baseline model I assume no targeted policy to change distribution (i.e. 3% for 10th decile vs 15% for 1st in 2020 with a constant growth projection to 2050). Model aims to reach exogenously government provided diffusion target for each technology (e.g.40% for housing retrofit vs 60% for EVs). 
To discuss with Charlie: how does government goal fit in with distribution ? for now, I dont have government level, though do have in earlier version, (see edits v1). I just made it 100 in the first level and didn't adopt for government target, just let the logistic growth rn to 2050. 


```{r Old US Baseline, echo=FALSE}
# Set government-specified saturation levels
us_saturation_levels <- tibble(
  technology = c(
    "Hybrid/EV", 
    "Solar Panels", 
    "Home Renovations", 
    "Smartphone", 
    "Internet",
    "Clean Heating Switch",
    "Digital Skills"
  ),
  saturation = c(
    60,   # EVs government target
    45,   # Solar Panels government target
    45,   # Home Renovations government target
    100,  # Smartphone government target
    100,  # Internet access government target
    69,   # Clean Heating Switch government target
    90    # Digital Skills government target
  ),
  # Add technology maturity classification
  maturity = c(
    "emerging",    # EVs
    "emerging",    # Solar
    "mature",      # Home Renovations
    "saturated",   # Smartphone
    "saturated",   # Internet
    "transitioning", # Clean Heating
    "transitioning"  # Digital Skills
  )
)

# Function to calculate logistic projection with improved S-curve dynamics
logistic_projection <- function(start_value, cagr, target_year, end_year, us_saturation_level, maturity) {
  years_forward <- target_year - end_year
  
  # Base growth rate from CAGR
  base_r <- if(is.na(cagr) || cagr == 0) {
    0.02  # Increased default minimal growth
  } else {
    log(1 + cagr/100)
  }
  
  # Adjust growth rate based on technology maturity and gap to target
  saturation_gap <- (us_saturation_level - start_value)/us_saturation_level
  
  maturity_factor <- case_when(
    maturity == "emerging" ~ 2.0,      # Accelerated growth for emerging tech
    maturity == "transitioning" ~ 1.5,  # Moderate acceleration
    maturity == "mature" ~ 1.0,        # Normal growth
    maturity == "saturated" ~ 0.5      # Slower growth near saturation
  )
  
  # Enhanced S-curve dynamics
  midpoint_factor <- if(start_value < (us_saturation_level/2)) {
    1.5  # Accelerate growth below midpoint
  } else {
    0.8  # Decelerate above midpoint
  }
  
  # Combine adjustments
  adjusted_r <- base_r * maturity_factor * midpoint_factor * (1 + saturation_gap)
  
  # Calculate projection ensuring we don't exceed saturation
  base_value <- min(
    us_saturation_level,
    us_saturation_level / (1 + ((us_saturation_level - start_value)/start_value) * exp(-adjusted_r * years_forward))
  )
  
  # Uncertainty bands that respect saturation
  uncertainty <- base_value * 0.1 * (years_forward/30) * saturation_gap
  lower_bound <- max(0, base_value - uncertainty)
  upper_bound <- min(us_saturation_level, base_value + uncertainty)
  
  c(base_value, lower_bound, upper_bound)
}

# Create projections
us_projections <- us_growth_results %>%
  left_join(us_saturation_levels, by = "technology") %>%
  rowwise() %>%
  mutate(
    # 2030 projections
    proj_2030 = list(logistic_projection(end_value, cagr, 2030, end_year, saturation, maturity)),
    proj_2030_main = proj_2030[1],
    proj_2030_lower = proj_2030[2],
    proj_2030_upper = proj_2030[3],
    
    # 2050 projections
    proj_2050 = list(logistic_projection(end_value, cagr, 2050, end_year, saturation, maturity)),
    proj_2050_main = proj_2050[1],
    proj_2050_lower = proj_2050[2],
    proj_2050_upper = proj_2050[3]
  ) %>%
  ungroup()

# Create visualization
ggplot() +
  # Target level
  geom_hline(data = us_saturation_levels, 
             aes(yintercept = saturation),
             linetype = "dashed", color = "gray50") +
  
  # Current values
  geom_line(data = us_projections, 
            aes(x = factor(decile), y = end_value, group = 1, color = "Current"),
            size = 1) +
  geom_point(data = us_projections, 
             aes(x = factor(decile), y = end_value, color = "Current"),
             size = 3) +
  
  # 2030 projections
  geom_ribbon(data = us_projections,
              aes(x = factor(decile), 
                  ymin = proj_2030_lower,
                  ymax = proj_2030_upper,
                  group = 1),
              fill = "blue", alpha = 0.1) +
  geom_line(data = us_projections, 
            aes(x = factor(decile), y = proj_2030_main, group = 1, color = "2030"),
            size = 1) +
  geom_point(data = us_projections, 
             aes(x = factor(decile), y = proj_2030_main, color = "2030"),
             size = 3) +
  
  # 2050 projections
  geom_ribbon(data = us_projections,
              aes(x = factor(decile), 
                  ymin = proj_2050_lower,
                  ymax = proj_2050_upper,
                  group = 1),
              fill = "red", alpha = 0.1) +
  geom_line(data = us_projections, 
            aes(x = factor(decile), y = proj_2050_main, group = 1, color = "2050"),
            size = 1) +
  geom_point(data = us_projections, 
             aes(x = factor(decile), y = proj_2050_main, color = "2050"),
             size = 3) +
  
  facet_wrap(~technology) +
  theme_minimal() +
  scale_color_manual(values = c("Current" = "black", "2030" = "blue", "2050" = "red")) +
  labs(
    title = "US Technology Adoption Projections by Income Decile",
    subtitle = "Projections to 2030 and 2050 with government targets (dashed lines)",
    x = "Income Decile",
    y = "Adoption Rate (%)",
    color = "Year"
  ) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    strip.text = element_text(size = 10, face = "bold")
  )

# Print summary statistics
summary_stats <- us_projections %>%
  group_by(technology) %>%
  summarise(
    target = first(saturation),
    mean_current = mean(end_value, na.rm = TRUE),
    mean_2030 = mean(proj_2030_main, na.rm = TRUE),
    mean_2050 = mean(proj_2050_main, na.rm = TRUE),
    gap_to_target = target - mean_2050,
    .groups = 'drop'
  )

print("Progress toward government targets:")
print(summary_stats)


us_projections_long <- us_projections %>%
  dplyr::select(technology, decile, 
                current = end_value, 
                proj_2030 = proj_2030_main, 
                proj_2050 = proj_2050_main) %>%
  pivot_longer(
    cols = c(current, proj_2030, proj_2050),
    names_to = "year_type",
    values_to = "adoption_rate"
  ) %>%
  mutate(
    year = case_when(
      year_type == "current" ~ 2021,
      year_type == "proj_2030" ~ 2030,
      year_type == "proj_2050" ~ 2050
    )
  )

# Visualization
ggplot(us_projections_long, aes(x = year, y = adoption_rate, 
                                color = technology, 
                                group = technology)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  facet_wrap(~ decile, ncol = 5, 
             labeller = labeller(decile = function(x) paste("Decile", x))) +
  geom_hline(data = us_saturation_levels, 
             aes(yintercept = saturation), 
             linetype = "dashed", 
             color = "gray50") +
  labs(
    title = "US Baseline Technology Adoption Across Deciles Over Time",
    subtitle = "Current and Projected Adoption Rates (with government targets)",
    x = "Year",
    y = "Adoption Rate (%)",
    color = "Technology"
  ) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    strip.text = element_text(size = 10, face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_x_continuous(breaks = c(2021, 2030, 2050))


# Visualization
ggplot(us_projections_long, aes(x = year, y = adoption_rate, 
                               color = factor(decile), 
                               group = decile)) +
 geom_line(size = 1) +
 geom_point(size = 3) +
 facet_wrap(~ technology, scales = "free_y") +
 labs(
   title = "Technology Adoption Progression by Decile",
   subtitle = "Adoption Rates Across Years",
   x = "Year",
   y = "Adoption Rate (%)",
   color = "Income Decile"
 ) +
 theme_minimal() +
 theme(
   panel.grid.minor = element_blank(),
   legend.position = "bottom",
   strip.text = element_text(size = 10, face = "bold"),
   axis.text.x = element_text(angle = 45, hjust = 1)
 ) +
 scale_x_continuous(breaks = c(2021, 2030, 2050))


# Updated logistic projection function
logistic_projection <- function(start_value, cagr, target_year, end_year, us_saturation_level, maturity) {
  years_forward <- target_year - end_year
  
  # Base growth rate from CAGR
  base_r <- if(is.na(cagr) || cagr == 0) {
    0.02  # Increased default minimal growth
  } else {
    log(1 + cagr/100)
  }
  
  # Adjust growth rate based on technology maturity and gap to target
  saturation_gap <- (100 - start_value)/100
  
  maturity_factor <- case_when(
    maturity == "emerging" ~ 2.0,      # Accelerated growth for emerging tech
    maturity == "transitioning" ~ 1.5, # Moderate acceleration
    maturity == "mature" ~ 1.0,        # Normal growth
    maturity == "saturated" ~ 0.5      # Slower growth near saturation
  )
  
  # Enhanced S-curve dynamics
  midpoint_factor <- if(start_value < 50) {
    1.5  # Accelerate growth below midpoint
  } else {
    0.8  # Decelerate above midpoint
  }
  
  # Combine adjustments
  adjusted_r <- base_r * maturity_factor * midpoint_factor
  
  # Calculate projection ensuring we don't exceed 100%
  projected_value <- min(
    100,  # Cap at 100%
    start_value * exp(adjusted_r * years_forward)
  )
  
  # Calculate uncertainty bands
  uncertainty <- projected_value * 0.1 * (years_forward/30)
  lower_bound <- max(0, projected_value - uncertainty)
  upper_bound <- min(100, projected_value + uncertainty)
  
  c(projected_value, lower_bound, upper_bound)
}

# Create projections
us_projections <- us_growth_results %>%
  left_join(us_saturation_levels, by = "technology") %>%
  rowwise() %>%
  mutate(
    # 2030 projections
    proj_2030 = list(logistic_projection(end_value, cagr, 2030, end_year, saturation, maturity)),
    proj_2030_main = proj_2030[1],
    proj_2030_lower = proj_2030[2],
    proj_2030_upper = proj_2030[3],
    
    # 2050 projections
    proj_2050 = list(logistic_projection(proj_2030_main, cagr, 2050, 2030, saturation, maturity)),
    proj_2050_main = proj_2050[1],
    proj_2050_lower = proj_2050[2],
    proj_2050_upper = proj_2050[3]
  ) %>%
  ungroup()

# Print summary statistics
us_summary_stats <- us_projections %>%
  group_by(technology) %>%
  summarise(
    target = first(saturation),
    mean_current = mean(end_value, na.rm = TRUE),
    mean_2030 = mean(proj_2030_main, na.rm = TRUE),
    mean_2050 = mean(proj_2050_main, na.rm = TRUE),
    gap_to_target = target - mean_2050,
    .groups = 'drop'
  )

```
```{r Updated Baseline, echo=FALSE}
calculate_conservative_growth <- function(cagr, start_value) {
    # Apply more conservative adjustments for high growth rates
    if (abs(cagr) > 5) {
        # For high growth rates, use a declining scale
        damping_factor <- 5 / abs(cagr)
        growth_rate <- sign(cagr) * min(abs(cagr) * damping_factor, 5)
    } else {
        growth_rate <- cagr
    }
    
    # Print diagnostics
    cat(sprintf("Original CAGR: %.2f%%\n", cagr))
    cat(sprintf("Adjusted growth rate: %.2f%%\n", growth_rate))
    cat(sprintf("Start value: %.2f%%\n", start_value))
    
    return(growth_rate)
}

# Function for logistic growth forecasting
forecast_smooth_logistic <- function(data, technology_val, decile_val, forecast_year = 2050) {
    # Create model data for specific decile and technology
    model_data <- data %>%
        filter(technology == technology_val, decile == decile_val) %>%
        arrange(end_year)
    
    # Check if we have valid data
    if(nrow(model_data) == 0) {
        cat("No valid data for", technology_val, "Decile", decile_val, "\n")
        return(NULL)
    }
    
    # Calculate growth rate using conservative approach
    growth_rate <- calculate_conservative_growth(
        cagr = model_data$cagr[1],
        start_value = model_data$end_value[1]
    )
    
    start_value <- model_data$end_value[1]
    start_year <- model_data$end_year[1]
    
    # Calculate logistic growth parameters
    K <- 100  # Maximum adoption (100%)
    r <- growth_rate / 100  # Convert percentage to decimal
    
    # Initial shift to align curve with start value
    t0 <- -log((K/start_value) - 1) / r
    
    # Generate future years for prediction
    future_years <- seq(start_year, forecast_year, by = 1)
    
    # Generate predictions using logistic function
    predictions <- tibble(
        year = future_years,
        predicted_adoption = sapply(future_years, function(year) {
            if(year <= start_year) return(start_value)
            
            # Time since start for logistic calculation
            t <- year - start_year + t0
            
            # Logistic function prediction
            predicted <- K / (1 + exp(-r * t))
            
            # Ensure predictions don't decrease and don't exceed 100%
            min(100, max(predicted, start_value))
        }),
        decile = decile_val
    )
    
    list(
        predictions = predictions,
        data = model_data,
        growth_rate = growth_rate,
        start_value = start_value,
        start_year = start_year
    )
}

# Function to plot growth curves
plot_smooth_logistic <- function(forecasts_list, technology_val, 
                               show_uncertainty = TRUE,
                               uncertainty_range = 0.1) {
    # Remove NULL entries
    forecasts_list <- forecasts_list[!sapply(forecasts_list, is.null)]
    
    # Combine all predictions
    all_predictions <- bind_rows(
        lapply(forecasts_list, function(x) x$predictions)
    )
    
    # Combine all historical data points
    all_data <- bind_rows(
        lapply(forecasts_list, function(x) {
            tibble(
                year = x$data$end_year,
                adoption = x$data$end_value,
                decile = x$predictions$decile[1]
            )
        })
    )
    
    # Create base plot
    p <- ggplot() +
        # Historical data points
        geom_point(data = all_data,
                  aes(x = year, y = adoption, color = factor(decile)),
                  alpha = 0.6, size = 2) +
        # Forecast lines
        geom_line(data = all_predictions,
                 aes(x = year, y = predicted_adoption, color = factor(decile)))
    
    if(show_uncertainty) {
        all_predictions <- all_predictions %>%
            mutate(
                lower_bound = pmax(0, predicted_adoption * (1 - uncertainty_range)),
                upper_bound = pmin(100, predicted_adoption * (1 + uncertainty_range))
            )
        
        p <- p + 
            geom_ribbon(data = all_predictions,
                       aes(x = year, 
                           ymin = lower_bound, 
                           ymax = upper_bound,
                           fill = factor(decile)),
                       alpha = 0.1)
    }
    
    p + 
        scale_color_viridis_d(name = "Income Decile") +
        scale_fill_viridis_d(name = "Income Decile") +
        theme_minimal() +
        labs(
            title = paste("Adoption Forecast by Income Decile:", technology_val),
            subtitle = paste("Smoothed logistic projections to", max(all_predictions$year)),
            x = "Year",
            y = "Adoption Rate (%)"
        ) +
        scale_x_continuous(breaks = seq(2010, 2050, by = 5)) +
        ylim(0, 100) +
        theme(
            panel.grid.minor = element_blank(),
            legend.position = "right"
        )
}

# Create nested list to store forecasts
forecasts_by_country <- list()

# First, look at the data
cat("Summary of CAGRs by technology and country:\n")
cagr_summary <- combined_growth %>%
  group_by(technology, country) %>%
  summarise(
    min_cagr = min(cagr, na.rm = TRUE),
    mean_cagr = mean(cagr, na.rm = TRUE),
    max_cagr = max(cagr, na.rm = TRUE),
    min_start = min(start_value, na.rm = TRUE),
    max_start = max(start_value, na.rm = TRUE),
    .groups = 'drop'
  )
print(cagr_summary)

# Get unique combinations of country and technology
country_tech_combos <- combined_growth %>%
  dplyr::select(country, technology) %>%
  distinct()

# Generate forecasts for each country and technology
for(i in 1:nrow(country_tech_combos)) {
    current_country <- country_tech_combos$country[i]
    current_tech <- country_tech_combos$technology[i]
    
    cat("\nProcessing", current_country, "-", current_tech, "\n")
    
    # Store forecasts for each decile
    decile_forecasts <- list()
    
    # Filter data for current country and technology
    current_data <- combined_growth %>%
        filter(country == current_country,
               technology == current_tech)
    
    # Print CAGR range for this technology-country combination
    cat(sprintf("CAGR range for %s in %s: %.2f%% to %.2f%%\n", 
                current_tech, current_country,
                min(current_data$cagr), max(current_data$cagr)))
    
    # Generate forecasts for each decile
    for(d in 1:10) {
        decile_forecasts[[d]] <- forecast_smooth_logistic(
            current_data, current_tech, d
        )
    }
    
    # Remove NULL entries
    decile_forecasts <- decile_forecasts[!sapply(decile_forecasts, is.null)]
    
    # Store forecasts if we have valid results
    if(length(decile_forecasts) > 0) {
        if(!current_country %in% names(forecasts_by_country)) {
            forecasts_by_country[[current_country]] <- list()
        }
        
        forecasts_by_country[[current_country]][[current_tech]] <- decile_forecasts
        
        # Generate and print plot
        plot <- plot_smooth_logistic(decile_forecasts, 
                                   paste(current_country, "-", current_tech))
        print(plot)
        
    }
}

# Create comparison plots across countries for each technology
technologies <- unique(combined_growth$technology)

for(tech in technologies) {
    # Combine predictions from all countries for this technology
    all_predictions <- bind_rows(lapply(names(forecasts_by_country), function(country) {
        if(tech %in% names(forecasts_by_country[[country]])) {
            forecasts <- forecasts_by_country[[country]][[tech]]
            bind_rows(lapply(forecasts, function(x) {
                x$predictions %>% 
                dplyr::select(year, predicted_adoption, decile) %>%
                mutate(country = country)
            }))
        }
    }))
    
    if(nrow(all_predictions) > 0) {
        # Create comparison plot
        comparison_plot <- ggplot(all_predictions, 
               aes(x = year, y = predicted_adoption, 
                   color = factor(decile), linetype = country)) +
            geom_line() +
            facet_wrap(~country) +
            scale_color_viridis_d(name = "Income Decile") +
            theme_minimal() +
            labs(
                title = paste("Technology Adoption Forecast Comparison -", tech),
                subtitle = "Projections by Country and Income Decile",
                x = "Year",
                y = "Adoption Rate (%)",
                linetype = "Country"
            ) +
            theme(
                panel.grid.minor = element_blank(),
                legend.position = "right"
            )
        
        print(comparison_plot)
    }
}

# Save forecasts
saveRDS(forecasts_by_country, "technology_forecasts.rds")

```

## Step 3: 2050 Diffusion Impacts

Step 3 Estimate how distributional impacts of technology uptake affect final energy costs for household across time. Use survey data to estimate how final energy costs increase/decrease with different technologies including in a household (holding all other variables constant) in 2020. Then, hold energy prices static (very large assumption) to see how technology diffusion affects final energy costs for residential sector at decile level in 2050.

I first get the average amount spent on energy, then roughly estimate how much is on gas vs electric, then get historical prices and convert amount spent to gas and eelctric to consumption. This is easier for some data (ex. US ) than others 

take a simple model to turn the technology into electricity 
```{r energy data, echo=FALSE}
DEU <- DEU %>%
  rename(year = syear)

# US Data Source: EIA Average Retail Electricity and Gas 2005-2022
us_prices <- tibble(
  country = "US",
  year = 2005:2022,
  elec_price = c(9.45, 10.40, 10.65, 11.26, 11.51, 11.54, 11.72, 11.88, 12.13,
                 12.52, 12.65, 12.55, 12.89, 12.87, 13.01, 13.15, 13.72, 14.77),  # cents per kWh
  gas_price = c(4.72, 4.68, 4.46, 4.74, 4.14, 3.89, 3.76, 3.65, 3.52, 3.74, 
                3.54, 3.43, 3.72, 3.58, 3.56, 3.68, 3.91, 5.45)   
)

# German Data Source: Federal Statistic Office of Germany 
deu_prices <- tibble(
  country = "DEU",
  year = 2005:2022,
  elec_price = c(
  22.5, 23.8, 27.4, 30.9, 30.6, 30.6, 34.8, 33.3, 38.6, 38.6,
  32.2, 33.3, 33.9, 35.4, 34.7, 36.5, 37.8, 35.7),  # cents per kWh (converted from EUR/kWh)
  gas_price = c(2.05, 2.38, 2.73, 3.41, 3.07, 3.07, 3.07, 3.07, 3.07, 
                3.07, 3.07, 3.07, 3.07, 3.07, 3.07, 3.07, 3.07, 3.41)   
)

# UK Data Source: UK Government Energy Price Statistics
uk_prices <- tibble(
  country = "UK",
  year = 2005:2022,
  elec_price = c(
  15.65, 18.56, 21.11, 22.40, 22.36, 22.94, 23.15, 23.68, 23.54, 23.72, 25.10, 25.88, 
  26.03, 27.14, 29.01, 31.65, 35.22, 40.01),  # cents per kWh (converted from pence/kWh)
  gas_price = c(0.61, 0.60, 0.56, 0.59, 0.53, 0.49, 0.48, 0.46, 0.45, 0.47, 
    0.46, 0.46, 0.51, 0.53, 0.56, 0.61, 0.66, 0.76)
)

# Combine all data
historical_prices <- bind_rows(
  us_prices,
  deu_prices,
  uk_prices
)

# Function to merge historical prices with country-specific datasets
merge_prices <- function(country_data, country_name) {
  country_data %>%
    left_join(
      historical_prices %>% filter(country == country_name),
      by = "year"
    )
}

# Create visualizations for available data
create_price_plot <- function(data) {
  ggplot(data %>% filter(!is.na(elec_price)), aes(x = year, color = country)) +
    geom_line(aes(y = elec_price, linetype = "Electricity"), size = 1) +
    geom_line(aes(y = gas_price, linetype = "Natural Gas"), size = 1) +
    facet_wrap(~country, scales = "free_y") +
    theme_minimal() +
    labs(
      title = "Historical Energy Prices (2005-2022)",
      subtitle = "By Country cent per kwH",
      y = "Price",
      color = "Country",
      linetype = "Energy Type",
      caption = "Sources: US - EIA, Germany - Destatis, UK - data.uk"
    ) +
    scale_color_manual(values = c("US" = "blue", "DEU" = "red", "UK" = "green")) +
    theme(
      legend.position = "bottom",
      strip.text = element_text(size = 12, face = "bold")
    )
}

# Function to print data summary with source documentation
print_summary <- function(data) {
  for (country_name in unique(data$country)) {
    country_data <- data %>% 
      filter(country == country_name, !is.na(elec_price))
    
    if (nrow(country_data) > 0) {
      cat(sprintf("\nSummary for %s:\n", country_name))
      print(summary(country_data[c("elec_price", "gas_price")]))
      cat("Data source: ", switch(country_name,
                                "US" = "EIA Average Retail Prices",
                                "DEU" = "[Add German source]",
                                "UK" = "[Add UK source]"),
          "\n")
    }
  }
}

historical_prices <- bind_rows(us_prices, deu_prices, uk_prices)
price_plot <- create_price_plot(historical_prices)
print_summary(historical_prices)
print(price_plot)
```

Now that we have average historical costs, I can convert the amount spent to consumption at decile level. 
```{r consumption creation, echo=FALSE}

energy_US <- US %>%
  group_by(year, decile) %>%
  summarise(
    avg_elec_exp = mean(elec_a_exp, na.rm = TRUE),
    avg_gas_exp = mean(gas_a_exp, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(year, decile) %>%
  mutate(country = "US")

energy_UK <- UK %>%
  group_by(year, decile) %>%
  summarise(
    avg_elec_exp = mean(xpelecy, na.rm = TRUE),
    avg_gas_exp = mean(xpgasy, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(year, decile)

Germany <- DEU %>%
  mutate(hgheat = ifelse(hgheat < 0, NA, hgheat)) 
energy_DEU <- Germany %>%
  group_by(year, decile) %>%
  summarise(
    avg_elec_exp = 12 * mean(hgelectr, na.rm = TRUE),
    avg_gas_exp = 12 * mean(hgheat, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(year, decile)

### exchange rate for Germany and UK 
getSymbols("DEXUSEU", src = "FRED")
exchange_rate_data <- data.frame(date = index(DEXUSEU), rate = as.numeric(DEXUSEU[, 1]))
exchange_rate_data$year <- format(exchange_rate_data$date, "%Y")
avg_exchange_rate <- exchange_rate_data %>%
  group_by(year) %>%
  summarise(avg_rate = mean(rate, na.rm = TRUE))

avg_exchange_rate$year <- as.numeric(avg_exchange_rate$year)



energy_DEU <- energy_DEU %>%
  left_join(avg_exchange_rate, by = "year") %>%
  mutate(
    avg_elec_exp_usd = avg_elec_exp * avg_rate,  # Convert to USD
    avg_gas_exp_usd = avg_gas_exp * avg_rate      # Convert to USD
  )

getSymbols("DEXUSUK", src = "FRED")
exchange_rate_data <- data.frame(date = index(DEXUSUK), rate = as.numeric(DEXUSUK[, 1]))
exchange_rate_data$year <- format(exchange_rate_data$date, "%Y")
avg_exchange_rate <- exchange_rate_data %>%
  group_by(year) %>%
  summarise(avg_rate = mean(rate, na.rm = TRUE))
avg_exchange_rate$year <- as.numeric(avg_exchange_rate$year)

# combine

energy_UK <- energy_UK %>%
  left_join(avg_exchange_rate, by = "year") %>%
  mutate(
    avg_elec_exp_usd = avg_elec_exp * avg_rate,  # Convert to USD
    avg_gas_exp_usd = avg_gas_exp * avg_rate      # Convert to USD
  )


energy_DEU_2 <- energy_DEU %>%
  dplyr::select(year, decile, avg_elec_exp_usd, avg_gas_exp_usd) %>%
  rename(
    avg_elec_exp = avg_elec_exp_usd,  
    avg_gas_exp = avg_gas_exp_usd
  ) %>%
  mutate(country = "DEU")

energy_UK_2 <- energy_UK %>%
  dplyr::select(year, decile, avg_elec_exp_usd, avg_gas_exp_usd) %>%
  rename(
    avg_elec_exp = avg_elec_exp_usd,  
    avg_gas_exp = avg_gas_exp_usd
  ) %>%
  mutate(country = "UK")

energy_combined <- bind_rows(energy_DEU_2, energy_US, energy_UK_2) %>%
  left_join(historical_prices, by = c("year", "country")) %>%
  mutate(gas_consumption = avg_gas_exp / (gas_price/100),
         elec_consumption = avg_elec_exp / (elec_price/100))


energy_combined_long <- energy_combined %>%
  pivot_longer(
    cols = c(gas_consumption, elec_consumption),
    names_to = "type",
    values_to = "consumption"
  )

# Create the plot
ggplot(energy_combined_long, aes(x = year, y = consumption, color = type)) +
  geom_line(aes(group = interaction(type, decile))) +
  facet_grid(country ~ decile, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Energy Consumption Over Time by Country and Decile",
    y = "Consumption (kWh)",
    x = "Year"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


energy_combined_long %>%
  filter(country == "DEU") %>%
  ggplot(aes(x = year, y = consumption, color = type)) +
  geom_line(aes(group = interaction(type, decile))) +
  facet_wrap(~decile, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "German Energy Consumption Over Time by Decile",
    y = "Consumption (kWh)",
    x = "Year"
  )

```
### Technology model for electricity 

```{r technology consumption, echo=FALSE}
technology_consumption <- data.frame(
  technology = c(
    "Owns a Smartphone",     # Daily charging
    "Hybrid/EV",            # EV charging
    "Solar Panels",         # (negative consumption/production)
    "Internet access",      # Router + devices
    "Home Renovations",     # Efficiency improvements
    "Digital Skills",       # Computer usage
    "Clean Heating Switch", # Moving from gas to electric heating
    "Sustainable Transport" # Public transport/biking
  ),
  kwh_per_year = c(
    30,      # Smartphone: daily charging
    2000,    # EV: annual charging for average use
    -3000,   # Solar: annual production
    180,     # Internet: router (~150 kWh) + device charging
    -500,    # Home renovations: efficiency savings
    250,     # Digital Skills: computer usage
    3500,    # Moving from gas to electric heating (increases electricity use)
    -200     # Sustainable transport: reduced car usage
  )
)

print(technology_consumption)

```

```{r combine with forecast, echo=FALSE}

elec_prices_2021 <- energy_combined %>%
  filter(year == 2021) %>%
  dplyr::select(country, elec_price) %>%
  distinct() %>%
  mutate(country = case_when(
    country == "DEU" ~ "Germany",
    TRUE ~ country
  ))

calculate_additional_demand <- function(forecasts_by_country, technology_consumption, elec_prices_2021) {
  additional_demand <- list()
  
  for(country in names(forecasts_by_country)) {
    cat("Processing country:", country, "\n")
    country_forecasts <- forecasts_by_country[[country]]
    
    elec_price <- elec_prices_2021$elec_price[elec_prices_2021$country == country] / 100
    
    if(length(elec_price) == 0) {
      cat("Warning: No electricity price found for", country, "\n")
      next
    }
    
    for(tech in names(country_forecasts)) {
      cat("Processing technology:", tech, "\n")
      tech_impact <- technology_consumption$kwh_per_year[technology_consumption$technology == tech]
      
      if(length(tech_impact) == 0) {
        cat("Warning: No technology impact found for", tech, "\n")
        next
      }
      
      decile_forecasts <- country_forecasts[[tech]]
      
      for(decile_forecast in decile_forecasts) {
        if(is.null(decile_forecast) || nrow(decile_forecast$predictions) == 0) next
        
        additional_cost <- decile_forecast$predictions %>%
          mutate(
            adoption_rate = predicted_adoption / 100,
            additional_kwh = adoption_rate * tech_impact,
            additional_cost = additional_kwh * elec_price,
            technology = tech,
            country = country
          )
        
        additional_demand[[paste(country, tech, decile_forecast$predictions$decile[1])]] <- additional_cost
      }
    }
  }
  
  result <- bind_rows(additional_demand)
  cat("Final number of rows:", nrow(result), "\n")
  return(result)
}

additional_demand <- calculate_additional_demand(forecasts_by_country, technology_consumption, elec_prices_2021)

```
Visualize the additional demand from technologies with our demand increase projections. 

```{r visualize additional costs, echo=FALSE}

# Create total_impact from additional_demand
total_impact <- additional_demand %>%
  filter(year >= 2021) %>%
  group_by(year, country, decile) %>%
  summarise(
    total_additional_cost = sum(additional_cost),
    .groups = 'drop'
  )

# Create a more informative plot
ggplot(total_impact, 
       aes(x = year, y = total_additional_cost, color = factor(decile))) +
  geom_line() +
  facet_wrap(~country, scales = "free_y") +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  labs(
    title = "Net Change in Annual Electricity Costs from Technology Adoption",
    subtitle = "Negative values indicate cost savings from solar generation and efficiency improvements",
    y = "Net Change in Electricity Costs (USD)",
    x = "Year",
    color = "Income Decile"
  ) +
  theme(
    legend.position = "bottom",
    plot.subtitle = element_text(size = 10)
  )

# Let's also show the technology breakdown to better understand the components
ggplot(additional_demand, 
       aes(x = year, y = additional_cost, fill = technology)) +
  geom_area(position = "stack") +
  facet_wrap(~country, scales = "free_y") +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  labs(
    title = "Components of Electricity Cost Changes",
    subtitle = "Showing contribution of each technology to total cost change",
    y = "Cost Impact (USD)",
    x = "Year",
    fill = "Technology"
  ) +
  theme(legend.position = "bottom")
```

#Step 4: Adoption S-curve impacts

I now have estimates of how much these additional technologies would cost each decile to 2050. Now it is time to run a better forecast model that assumes earlier or later adoption trends of technologies. I want to look at how earlier or later adoption affects. 

```{r fast and slow adoption, echo=FALSE}
starting_values <- combined_growth %>%
  group_by(technology, decile, country) %>%
  slice_max(end_year, n = 1) %>%
  dplyr::select(technology, decile, end_value, country, end_year)

print("Checking starting values:")
print(starting_values %>% arrange(technology, country, decile))


# Enhanced s-curve function with decile variation
create_adoption_curve <- function(
  technology, 
  decile,
  country,
  start_year = 2021,
  end_year = 2050,
  start_value,
  scenario = "conservative",
  max_adoption = NULL  # Will vary by technology and decile
) {
  
  # Define maximum adoption rates by technology and decile
  # Higher deciles generally have higher maximum adoption
  base_max <- case_when(
    technology == "Hybrid/EV" ~ 80,
    technology == "Solar Panels" ~ 40,
    technology == "Internet access" ~ 100,
    technology == "Owns a Smartphone" ~ 100,
    technology == "Digital Skills" ~ 90,
    technology == "Clean Heating Switch" ~ 70,
    technology == "Home Renovations" ~ 60,
    technology == "Sustainable Transport" ~ 50,
    TRUE ~ 100
  )
  
  # Adjust max adoption by decile
  decile_adjustment <- (decile - 5.5) * 5  # Deciles 1-10 get -22.5 to +22.5 adjustment
  
  # If max_adoption not specified, calculate it
  if(is.null(max_adoption)) {
    max_adoption <- min(100, base_max + decile_adjustment)
  }
  
  # Set growth rate and midpoint based on scenario
  if(scenario == "aggressive") {
    k <- 0.15 + (decile/100)  # Higher deciles adopt slightly faster
    midpoint <- 2035 - (decile/10)  # Higher deciles reach midpoint earlier
  } else {  # conservative
    k <- 0.08 + (decile/200)
    midpoint <- 2040 - (decile/20)
  }
  
  # Generate years
  years <- start_year:end_year
  
  # Ensure start_value doesn't exceed max_adoption
  start_value <- min(start_value, max_adoption)
  
  # Calculate shift parameter to match start_value
  shift <- -log((max_adoption/start_value) - 1) / k
  
  # Calculate adoption for each year
  adoption <- tibble(
    year = years,
    adoption = max_adoption / (1 + exp(-k * (years - midpoint))),
    decile = decile,
    technology = technology,
    country = country,
    scenario = scenario
  )
  
  return(adoption)
}

# Generate curves for all technologies and deciles
adoption_curves <- map_dfr(
  1:nrow(starting_values),
  function(i) {
    row <- starting_values[i,]
    
    # Generate both conservative and aggressive scenarios
    bind_rows(
      create_adoption_curve(
        technology = row$technology,
        decile = row$decile,
        country = row$country,
        start_value = row$end_value,
        scenario = "conservative"
      ),
      create_adoption_curve(
        technology = row$technology,
        decile = row$decile,
        country = row$country,
        start_value = row$end_value,
        scenario = "aggressive"
      )
    )
  }
)

ggplot(adoption_curves, 
       aes(x = year, y = adoption, color = factor(decile))) +
  geom_line() +
  facet_grid(country + scenario ~ technology) +  # Added country to faceting
  theme_minimal() +
  labs(
    title = "Technology Adoption Curves by Income Decile and Country",
    subtitle = "Comparing Conservative vs Aggressive Scenarios",
    y = "Adoption Rate (%)",
    x = "Year",
    color = "Income Decile"
  ) +
  theme(legend.position = "bottom")

ggplot(
  adoption_curves %>% filter(country == "UK"),
  aes(x = year, y = adoption, color = factor(decile))
) +
  geom_line() +
  facet_grid(scenario ~ technology) +
  theme_minimal() +
  labs(
    title = "UK Technology Adoption Curves by Income Decile",
    subtitle = "Comparing Conservative vs Aggressive Scenarios",
    y = "Adoption Rate (%)",
    x = "Year",
    color = "Income Decile"
  ) +
  theme(legend.position = "bottom")

# You might also want to see just one technology to compare scenarios more clearly
ggplot(
  adoption_curves %>% filter(technology == "Hybrid/EV"),
  aes(x = year, y = adoption, color = factor(decile))
) +
  geom_line() +
  facet_wrap(~scenario) +
  theme_minimal() +
  labs(
    title = "EV Adoption Curves by Income Decile",
    subtitle = "Comparing Conservative vs Aggressive Scenarios",
    y = "Adoption Rate (%)",
    x = "Year",
    color = "Income Decile"
  )

```
```{r additional costs, echo=FALSE}
# First, let's confirm our kwh impacts from technology_consumption
print(technology_consumption)

# Calculate additional demand with new adoption curves
new_additional_demand <- adoption_curves %>%
  # Join with technology consumption assumptions
  left_join(technology_consumption, by = "technology") %>%
  # Join with electricity prices (using 2021 prices)
  left_join(elec_prices_2021, by = "country") %>%
  # Drop year.y and rename year.x to year
  dplyr::select(-year.y) %>%
  rename(year = year.x) %>%
  # Calculate costs
  mutate(
    adoption_rate = adoption / 100,  # Convert percentage to decimal
    additional_kwh = adoption_rate * kwh_per_year,
    additional_cost = additional_kwh * (elec_price/100)  # Convert cents to dollars
  )

# Calculate total impact by scenario
new_total_impact <- new_additional_demand %>%
  group_by(year, country, decile, scenario) %>%
  summarise(
    total_additional_cost = sum(additional_cost),
    .groups = 'drop'
  )

# Visualize the new projections
ggplot(new_total_impact, 
       aes(x = year, y = total_additional_cost, color = factor(decile))) +
  geom_line() +
  facet_grid(scenario ~ country, scales = "free_y") +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  labs(
    title = "Projected Additional Annual Electricity Costs with New Adoption Curves",
    subtitle = "By country and adoption scenario",
    y = "Net Change in Electricity Costs (USD)",
    x = "Year",
    color = "Income Decile"
  ) +
  theme(legend.position = "bottom")

```

Questions for Charlie: do we want to add technology specific s-curves ? Right now they are general. technology specific would look more like this: 

```{r technology specific s-curve, echo=FALSE}
create_adoption_curve <- function(
  technology, 
  decile,
  country,
  start_year = 2021,
  end_year = 2050,
  start_value,
  scenario = "conservative"
) {
  
  # Technology-specific parameters
  tech_params <- list(
    "Hybrid/EV" = list(
      base_max = 80,
      aggressive_k = 0.20,
      conservative_k = 0.10,
      aggressive_midpoint = 2033,
      conservative_midpoint = 2038
    ),
    "Solar Panels" = list(
      base_max = 40,
      aggressive_k = 0.15,
      conservative_k = 0.08,
      aggressive_midpoint = 2035,
      conservative_midpoint = 2040
    ),
    "Internet access" = list(
      base_max = 100,
      aggressive_k = 0.25,  # Fast adoption
      conservative_k = 0.15,
      aggressive_midpoint = 2030,
      conservative_midpoint = 2035
    ),
    "Owns a Smartphone" = list(
      base_max = 100,
      aggressive_k = 0.25,  # Fast adoption
      conservative_k = 0.15,
      aggressive_midpoint = 2030,
      conservative_midpoint = 2035
    ),
    "Digital Skills" = list(
      base_max = 90,
      aggressive_k = 0.18,
      conservative_k = 0.12,
      aggressive_midpoint = 2032,
      conservative_midpoint = 2037
    ),
    "Clean Heating Switch" = list(
      base_max = 70,
      aggressive_k = 0.15,
      conservative_k = 0.08,
      aggressive_midpoint = 2035,
      conservative_midpoint = 2040
    ),
    "Home Renovations" = list(
      base_max = 60,
      aggressive_k = 0.12,
      conservative_k = 0.06,
      aggressive_midpoint = 2037,
      conservative_midpoint = 2042
    ),
    "Sustainable Transport" = list(
      base_max = 50,
      aggressive_k = 0.15,
      conservative_k = 0.08,
      aggressive_midpoint = 2035,
      conservative_midpoint = 2040
    )
  )
  
  # Get parameters for this technology
  params <- tech_params[[technology]]
  
  # Adjust max adoption by decile
  decile_adjustment <- (decile - 5.5) * 5  # Deciles 1-10 get -22.5 to +22.5 adjustment
  max_adoption <- min(100, params$base_max + decile_adjustment)
  
  # Set growth rate and midpoint based on scenario
  if(scenario == "aggressive") {
    k <- params$aggressive_k + (decile/100)
    midpoint <- params$aggressive_midpoint - (decile/10)
  } else {
    k <- params$conservative_k + (decile/200)
    midpoint <- params$conservative_midpoint - (decile/20)
  }
  
  # Generate years
  years <- start_year:end_year
  
  # Ensure start_value doesn't exceed max_adoption
  start_value <- min(start_value, max_adoption)
  
  # Calculate adoption for each year
  adoption <- tibble(
    year = years,
    adoption = max_adoption / (1 + exp(-k * (years - midpoint))),
    decile = decile,
    technology = technology,
    country = country,
    scenario = scenario
  )
  
  return(adoption)
}

# UK visualization
ggplot(adoption_curves %>% filter(country == "UK"), 
       aes(x = year, y = adoption, color = factor(decile))) +
  geom_line() +
  facet_grid(scenario ~ technology, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "UK Technology Adoption Curves by Income Decile",
    subtitle = "Comparing Conservative vs Aggressive Scenarios",
    y = "Adoption Rate (%)",
    x = "Year",
    color = "Income Decile"
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# US visualization
ggplot(adoption_curves %>% filter(country == "US"), 
       aes(x = year, y = adoption, color = factor(decile))) +
  geom_line() +
  facet_grid(scenario ~ technology, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "US Technology Adoption Curves by Income Decile",
    subtitle = "Comparing Conservative vs Aggressive Scenarios",
    y = "Adoption Rate (%)",
    x = "Year",
    color = "Income Decile"
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

# Germany visualization
ggplot(adoption_curves %>% filter(country == "Germany"), 
       aes(x = year, y = adoption, color = factor(decile))) +
  geom_line() +
  facet_grid(scenario ~ technology, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Germany Technology Adoption Curves by Income Decile",
    subtitle = "Comparing Conservative vs Aggressive Scenarios",
    y = "Adoption Rate (%)",
    x = "Year",
    color = "Income Decile"
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```