---
title: "Edits Notebook"
author: "Cassandra Etter"
date: "2024-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the data behind the EDITS work. It follows four steps. For Step 1 I look at uptake rates between 2010 and 2021 in three countries: US, UK and Germany. For the US, I use the PSID which tracks households between 2011- 2021 on a biannual basis. For Germany, I use SOEP which tracks households between 2010- 2021 on an annual basis, and for the UK I use Understanding Society which tracks households between 2010 - 2021 on an annual basis. All data was downloaded in September 2024 and cleaning files are on github.

8 technologies selected are :Internet Access, own a smartphone, Home has Solar, Home switched from Gas/Coal/Wood to electric (fuel switching), Home renovation/ retrofitting, EV/ Hybrid car, Commuting habit (walking, biking), digital skills/ capabilities (daily use of internet).

Note: Significant limitations on technologies surveyed. With better data, I could also do this for smart-meters, air conditioners, etc.
# Load Lbraries
```{r libraries,include=FALSE}

if (!require(pacman)) install.packages("pacman")

pacman::p_load(dplyr, tidyr,data.table, readr, purrr, ggplot2, scales, survey, nls2, minpack.lm,forecast, knitr, kableExtra, broom, countrycode, priceR, quantmod, modelsummary, kableExtra, stringr, tseries, boot,
    install = TRUE,
  update = FALSE
)
sapply(c("dplyr", "ggplot2", "tidyr"), packageVersion)
```

Set up environment

```{r setup, include=FALSE}
# Require specific R version
if (getRversion() < "4.1.0") {
  stop("This script requires R version 4.1.0 or higher")
}

# Set global knitr options
knitr::opts_chunk$set(
  echo = TRUE,           # Show code chunks
  warning = FALSE,       # Hide warnings
  message = FALSE,       # Hide messages
  error = FALSE,         # Stop on errors
  cache = TRUE,          # Cache code chunk results
  fig.width = 10,        # Default figure width
  fig.height = 6,        # Default figure height
  dpi = 300              # High-resolution figures
)

# Set seed for reproducibility
set.seed(2024)
```

```{r load UK DEU, include=FALSE}

UK <- read_csv("~/Documents/Github/PhD/Data/UK_Clean.csv", show_col_types = FALSE)
DEU <- read_csv("~/Documents/Github/PhD/Data/DEU_Clean.csv", show_col_types = FALSE, na = "NA")
US <- read_csv("~/Documents/Github/PhD/Data/US_Clean.csv", show_col_types = FALSE, na = "NA")
```
## Income Minimum Threshold ### 
For our data, we have a number of households that state they make way below a ''living salary''. To calculate the floor, I make a grand assumption that at least one individual in the home is working full time on minimum wage salary (or benefits which is included in income). 

If the country has a strong benefits system (DEU), I take whatever the minimum-income benefits expenditure is (20% of average wage) which is the average benefits you'd receive from government. this is EUR695 per month in 2010. The standard benefit in 2010 was 345 euro per month, or 4140 EUR a year.
Germany minimum wage was introduced in 2015 and was €8.50, €8.84 in 2017 and €9.19 in 2019, thus it is 695 for 2010 - 2015, then 1,360 in 2015. To error on the side of caution, I keep it flat at 695 EUR per month or 8340 per year. 

For the US, federal income minimum wage amount doesn't change and is $1,150 net per month, or 13,800 per year. If they are single hh benefits, the maximum federal benefit is USD 901 per month or 10816 per year and there is no minimum, but lowest benefits would be around 292 a month (children of disabled workers in 2010). For long-term low earners there is also the special minimum benefit of SSI, which is 170 per month. To be cautious, I make the floor the benefits amount. This is 2040 a year. 


For UK federal minimum is £5.05 in 2005 and rises to £10 over 18 in 2025, which net is 9,696 a year. Benefits (universal credit starting in 2012) is GBP 2,847 per year (1517 single on universal). I take universal credit to error on the side of caution.

To measure benefits, I use OECD 2010 publication on minimum income benefits in OECD countries <https://www.oecd.org/content/dam/oecd/en/publications/reports/2010/01/minimum-income-benefits-in-oecd-countries_g17a1d77/218402763872.pdf>. 

This eliminates 150 observations from US, no observations from the UK, and 38 observations from DEU. 

```{r income minimum, echo = FALSE}

US<- US %>%
  filter(hh_income >= 2040) %>%
  filter(adjusted_income >= 2040)

UK <- UK %>%
  filter(hh_incomey >= 1517) %>%
  filter(adjusted_income >= 1517)

DEU <- DEU %>%
  filter(hh_incomey >= 4140) %>%
  filter(adjusted_income >= 4140)
  
```

# Step 1: Uptake Rates

This provides a snapshot for current uptake across the households assessing heterogeneity across income level.

### U.S. Uptake Rates

For the US, I have information for solar ownership, EV/ hybrid ownership, smartphone ownership, internet access, and commute habit. I create an additional variable on fuel switching by looking at heating method and seeing where moved from wood, oil, gas, coal to electric heating. I also create a variable for retrofitting based on home renovations and skills based on survey responses where RP responds they use the internet daily.

Note: Smartphone ownership and digital skills questions were only added to the survey in 2015.

```{r US Uptake Rates, echo=FALSE}
US <- US %>%
  group_by(pid) %>%
  arrange(pid, year) %>%
  mutate(
    # Create a combined solar indicator
    is_solar = heat_method == 6 | heat_method_2 == 6 | heat_method_3 == 6,
    
    # Track those who switched to electric
    switched_to_electric = (lag(heat_method) %in% c(1,3,4,5,10,11) |
                    lag(heat_method_2) %in% c(1,3,4,5,10,11) |
                    lag(heat_method_3) %in% c(1,3,4,5,10,11)) & 
                          heat_method == 2,
    
    # Track those using electricity 
    always_electric = all(heat_method == 2, na.rm = TRUE),
        switched_to_solar = lag(is_solar) == FALSE & is_solar == TRUE,
        always_solar = all(is_solar, na.rm = TRUE),
        using_electric = heat_method == 2,
    using_solar = is_solar,
        switch_year_electric = case_when(
      switched_to_electric ~ year,
      TRUE ~ NA_real_
    ),
    switch_year_solar = case_when(
      switched_to_solar ~ year,
      TRUE ~ NA_real_
    )
  ) %>%
  ungroup() %>%
  filter(year > 2010)
# For decile analysis:
uptake_rates <- US %>%
  group_by(year, weighted_decile) %>%
  summarise(
    total_households = n_distinct(pid),
    
    # Electric heating metrics
    new_electric_switches = sum(switched_to_electric, na.rm = TRUE),
    total_electric_users = sum(using_electric, na.rm = TRUE),
    
    # Solar heating metrics
    new_solar_switches = sum(switched_to_solar, na.rm = TRUE),
    total_solar_users = sum(using_solar, na.rm = TRUE),
    
    # Calculate rates
    electric_switch_rate = (new_electric_switches / total_households) * 100,
    total_electric_rate = (total_electric_users / total_households) * 100,
    solar_switch_rate = (new_solar_switches / total_households) * 100,
    total_solar_rate = (total_solar_users / total_households) * 100
  ) %>%
  ungroup()

us_props <- US %>%
  group_by(year, weighted_decile) %>%
  summarise(
    n_total = n(),
    smartphone_prop = mean(smartphone == 1, na.rm = TRUE),
    hybrid_prop = mean(hybrid == 1 | hybrid_2 == 1 | hybrid_3 == 1, na.rm = TRUE), # includes EVs 
    solar_prop = mean(heat_method == 6 | heat_method_2 ==6 | heat_method_3 == 6, na.rm = TRUE),
    internet_prop = mean(computer == 1, na.rm = TRUE),     # Computer ownership
    renovation_prop = mean(hh_repair > 1, na.rm = TRUE), 
    skills_prop = mean(internet == 1, na.rm = TRUE),       # Internet usage
    fuel_switch_prop = mean(heat_method == 2 & 
                                                (heat_method_2 == 0 | heat_method_2 == 2 | heat_method_2 == 6) &
                                                (heat_method_3 == 0 | heat_method_3 == 2 | heat_method_3 == 6), 
                                        na.rm = TRUE)
  ) %>%
  mutate(
    smartphone_percent = smartphone_prop * 100,            # Remove duplicate line
    hybrid_percent = hybrid_prop * 100,
    solar_percent = solar_prop * 100, 
    internet_percent = internet_prop * 100, 
    repairs_percent = renovation_prop * 100, 
    skills_percent = skills_prop * 100,
    fuel_switch_percent = fuel_switch_prop * 100 
  )


us_props_long <- us_props %>%
  dplyr::select(year, weighted_decile, smartphone_percent, hybrid_percent, solar_percent, 
         internet_percent, repairs_percent, skills_percent, 
         fuel_switch_percent) %>%
  pivot_longer(
    cols = c(smartphone_percent, hybrid_percent, solar_percent, 
             internet_percent, repairs_percent, skills_percent, 
             fuel_switch_percent),
    names_to = "technology",
    values_to = "percentage"
  ) %>%
  mutate(technology = factor(technology, 
    levels = c("smartphone_percent", "hybrid_percent", "solar_percent", 
               "internet_percent", "repairs_percent", "skills_percent", 
               "fuel_switch_percent"),
    labels = c("Smartphone", "Hybrid/EV", "Solar Panels", "Internet", 
               "Home Renovations", "Digital Skills", "Clean Heating Switch")))
# Create the plot
ggplot(us_props_long, aes(x = weighted_decile, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year) +
  theme_minimal() +
  labs(
    title = "Technology Adoption by Income Decile in US",
    x = "Income Decile",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  scale_x_continuous(breaks = 1:10) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2)) # Makes legend more compact

# Plot across time
ggplot(us_props_long, aes(x = year, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~weighted_decile) +
  theme_minimal() +
  labs(
    title = "US Technology Adoption Over Time by Income Decile",
    x = "Year",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2))

ggsave("~/Documents/GitHub/EDITS/Output/us_adoption.png")

```

Clean heating is interesting in that high income households had less electric only and marginally increased over 2011 - 2021 time period, though they still did increase within decile 9 and 10. Solar panel infiltration is lower than national average at 1% with EIA estimating 4%.

### U.K. Uptake Rates

For the UK, I have solar, EV/ hybrid, smartphone, internet, commute and skills. I create fuel switching in the same manner as for US, buy looking at home energy bills and looking for those households that switch from gas/coal to electric only.

Solar question is only asked in 2009, 2012, 2018, and 2021. EV is asked in 2012, 2015, 2018, and 2021. Smartphone begins in 2013. Skills starts in 2011.

Commuting habit question changes in 2018 to a new question, so I only track to 2017. 

```{r UK Uptake Rates, echo=FALSE}
# UK proportions
uk_props <- UK %>%
  group_by(year, weighted_decile) %>%
  summarise(
    n_total = n(),
    solar_prop = mean(solar1 %in% c("yes - fitted", "Yes - fitted") | 
                        solar2 %in% c("yes - fitted", "Yes - fitted") |
                        solar3 %in% c("yes - fitted", "Yes - fitted"), na.rm = TRUE),
     hybrid_prop = mean(
            carfuel1 %in% c("Electric / battery", "electric / battery", "Hybrid (petrol/electric)", "hybrid (petrol/electric)") |
            carfuel2 %in% c("Electric / battery", "electric / battery", "Hybrid (petrol/electric)", "hybrid (petrol/electric)") | 
              carfuel3 %in%  c("Electric / battery", "electric / battery", "Hybrid (petrol/electric)", "hybrid (petrol/electric)") |
              carfuel4 %in%  c("Electric / battery", "electric / battery", "Hybrid (petrol/electric)", "hybrid (petrol/electric)"), na.rm = TRUE),
    smartphone_prop = mean(smartmob %in% c("Yes", "yes"), na.rm = TRUE ),
    skills_prop = mean(netpuse %in% c("Every day", "every day") | netpusenew %in% c("Every day", "every day"), na.rm = TRUE),
    internet_prop = mean(pcnet %in% c("yes", "Yes"), na.rm = TRUE),
     commute_prop = mean(
                        case_when(
                                year < 2018 ~ worktrav %in% c(
                                        "Bus/coach", "bus/coach", "Cycle", "cycle", "Walk", "walk",
                                        "Train", "train", "Underground/Metro/Tram/Light railway", 
                                        "underground/metro/tram/light railway",
                                        "underground/metro/tram/light railway (if england/scotland/wales)",
                                        "underground/metro/tram/light railway {if region = gb}",
                                        "Motorcycle/moped/scooter", "motorcycle/moped/scooter"
                                ),
                                year >= 2018 ~ (wktrv10 == "Yes mentioned" |  # Underground/metro
                                                        wktrv6 == "Yes mentioned" |   # Bus
                                                        wktrv7 == "Yes mentioned" |   # Train
                                                        wktrv9 == "Yes mentioned" |   # Walk
                                                        wktrv8 == "Yes mentioned")   # Cycle
                        ),
                        na.rm = TRUE
                ),
    fuel_switch_prop = mean(xpelecy > 0 & 
                                                (xpgasy <= 0 | is.na(xpgasy)) & 
                                                (xpoily <= 0 | is.na(xpoily)), 
                                        na.rm = TRUE)) %>%
  mutate(
    solar_percent = solar_prop * 100,
    ev_percent = hybrid_prop * 100,
    smartphone_percent = smartphone_prop * 100, 
    internet_percent = internet_prop * 100, 
    commute_percent = commute_prop * 100,
    skills_percent = skills_prop * 100, 
    fuel_switch_percent = fuel_switch_prop * 100
  )

# Convert to long format with all technologies
uk_props_long <- uk_props %>%
  dplyr::select(year, weighted_decile, solar_percent, ev_percent, smartphone_percent, 
         internet_percent, skills_percent, commute_percent, fuel_switch_percent) %>%
  pivot_longer(
    cols = c(solar_percent, ev_percent, smartphone_percent, 
             internet_percent, skills_percent, commute_percent, fuel_switch_percent),
    names_to = "technology",
    values_to = "percentage"
  ) %>%
  mutate(technology = factor(technology, 
    levels = c("ev_percent", "solar_percent", "smartphone_percent", 
               "internet_percent", "skills_percent", "commute_percent", "fuel_switch_percent"),
    labels = c("Hybrid/EV", "Solar Panels", "Smartphone", 
               "Internet", "Digital Skills", "Sustainable Transport", "Clean Heating Switch")))

ggplot(uk_props_long, aes(x = weighted_decile, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year) +
  theme_minimal() +
  labs(
    title = "Technology Adoption by Income Decile in UK",
    x = "Income Decile",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  scale_x_continuous(breaks = 1:10) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2)) # Makes legend more compact

ggplot(uk_props_long[uk_props_long$percentage != 0, ], aes(x = year, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~weighted_decile) +
  theme_minimal() +
  labs(
    title = "Technology Adoption Over Time by Income Decile UK",
    x = "Year",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2))

ggsave("~/Documents/GitHub/EDITS/Output/uk_adoption.png")


```

### Germany Uptake Rates

For Germany I have information on solar, EV, retrofitting, smartphone, internet.

EV only exists in 2015 and 2020. (note it is full EV or hydrogen, no hybrid category). Remove this variable as 2015 only has 1 response per decile and 2020 isn't much better. 

Retrofitting is asked from 2010 - 2015, and 2019. 

Smartphone hits 100% adoption in 2016, and SOEP stop asking the question. 
```{r rename syear, include=FALSE}
DEU <- DEU %>%
  rename(year = syear)
```

```{r DEU Uptake Rates, echo=FALSE}

deu_props <- DEU %>%
  group_by(year, weighted_decile) %>%
  summarise(
    n_total = n(),
    hybrid_prop = ifelse(sum(!is.na(c(hli0121, hli0122, hli0123, hli0124, hli0128, hli0129, hli0130, hli0131))) == 0,
                     NA_real_,
                     mean(rowSums(data.frame(hli0121 == 1, hli0122 == 1, hli0123 == 1, hli0124 == 1,
                                           hli0128 == 1, hli0129 == 1, hli0130 == 1, hli0131 == 1), 
                                na.rm = TRUE) > 0, 
                         na.rm = TRUE)),
    solar_prop = sum(hgeqpsol == 1 | hgeqpnrj == 1 , na.rm = TRUE) / 
             sum(hgeqpsol %in% c(1, 2) | hgeqpnrj %in% c(1, 2), na.rm = TRUE),
    renovation_prop = sum(hgcondit == 2 | hgcondit == 3, na.rm = TRUE) /n(),
    smartphone_prop = if(first(year) > 2015) {
                        1
                } else {
                        sum(hgeqptel == 1, na.rm = TRUE) / n()
                },
    internet_prop = sum(hlf0178_h == 1, na.rm = TRUE) / 
                  sum(hlf0178_h %in% c(1, 2), na.rm = TRUE),
    fuel_switch_prop = if(first(year) >= 2016) {
                        sum(energy_combination == 3, na.rm = TRUE) / n()
                } else {
                        NA_real_
                }
  ) %>%
  group_by(weighted_decile) %>%
  arrange(year) %>%
  mutate(
    internet_prop = if_else(year == 2020, NA_real_, internet_prop ) ## 2020 had different methodology in survey
  ) %>%
   mutate(
    hybrid_percent = ifelse(hybrid_prop == 0, NA_real_, hybrid_prop * 100),
    solar_percent = ifelse(solar_prop == 0, NA_real_, solar_prop * 100), 
    renovation_percent = ifelse(renovation_prop == 0, NA_real_, renovation_prop * 100), 
    smartphone_percent = ifelse(smartphone_prop == 0, NA_real_, smartphone_prop * 100),
    internet_percent = ifelse(internet_prop == 0, NA_real_, internet_prop * 100),
    fuel_switch_percent = ifelse(fuel_switch_prop == 0, NA_real_, fuel_switch_prop * 100)
  )


deu_props_long <- deu_props %>%
  dplyr::select(year, solar_percent, renovation_percent, smartphone_percent, internet_percent, fuel_switch_percent, hybrid_percent) %>%  
  pivot_longer(
    cols = c( solar_percent, renovation_percent, smartphone_percent, internet_percent, fuel_switch_percent, hybrid_percent),
    names_to = "technology",
    values_to = "percentage"
  ) %>%
  mutate(technology = factor(technology, 
    levels = c( "solar_percent", "renovation_percent", "smartphone_percent", "internet_percent", "fuel_switch_percent", "hybrid_percent"),
    labels = c("Solar Panels", "Home Renovations", "Owns a Smartphone", "Internet access", "Clean Heating Switch", "Hybrid/EV")))

ggplot(deu_props_long, aes(x = weighted_decile, y = percentage, color = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year) +
  theme_minimal() +
  labs(
    title = "Technology Adoption by Income Decile in Germany",
    x = "Income Decile",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  scale_x_continuous(breaks = 1:10) +
  theme(legend.position = "bottom")

# Remove NA values from the dataset
deu_props_long_clean <- deu_props_long %>%
  filter(!is.na(percentage)) 
      #filter(!(technology == "Hybrid/EV"),
           #!(technology == "Home Renovations" & year == 2019))
  

# Then create the plot
ggplot(deu_props_long_clean, aes(x = year, y = percentage, color = technology, group = technology)) +
  geom_line() +
  geom_point() +
  facet_wrap(~weighted_decile) +
  theme_minimal() +
  labs(
    title = "Technology Adoption Over Time by Income Decile Germany",
    x = "Year",
    y = "Percentage of Households (%)",
    color = "Technology"
  ) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    panel.grid.minor = element_blank()
  ) +
  guides(color = guide_legend(nrow = 2))


ggsave("~/Documents/GitHub/EDITS/Output/de_adoption.png")

```


There are two outlier technologies for uptake rates. The first is Sustainable Transit in the UK. UK commute (sustainable transport) only shows commuting patterns for those who work, and because the other two countries don't have sustainable commute, its harder to look at comparisons. The second is home renovations in Germany. If i change the year to later, the figure does go up, but 2010 shows a renovation boom that is lost in later years.


# Step 2: Projections with S-Curve

### Version 1 - Simple
Simple Non linear Regression. Q1-5 and goes to 2100. 
```{r s-curve, include = FALSE}
# Configure maximum penetration hypothesis
max_penetration <- tribble(
  ~tech, ~max_share,
  'solar', 0.9,
  'smartphone', 1,
  'skills', 0.95,
  'internet', 1,
  'hybrid', 0.5,
  'commute', 0.5,
  'fuel_switch', 0.7,
  'renovation', 0.75,
) %>% as.data.table

# Aggregate all 3 countries to the quintile level,

props_long <- bind_rows(
  deu_props %>% mutate(iso2 = "DE") ,
  uk_props  %>% mutate(iso2 = "GB"),
  us_props  %>% mutate(iso2 = "US") 
) %>%
  mutate(quintile = (weighted_decile - 1) %/% 2 + 1) %>% group_by(iso2, year, quintile) %>%
  summarise(across(ends_with('prop'), ~weighted.mean(., n_total, na.rm = TRUE))) %>% as.data.table %>%
  melt(id.vars = c('iso2', 'year', 'quintile'), value.name = 'share', variable.name = 'tech') %>%
  mutate(tech = str_remove(tech, '_prop')) %>%
  # The logistics curve will be fitted to non-zero, non-NA shares
  filter(share != 0, !is.na(share),
         # There's apparently a problem in the computation of the Commute share over 2018:2021 in the UK
         !(year %in% 2018:2021 & tech == 'commute'),
         # Not enough observations (2) for hybrids in Germany
         !(iso2 == 'DE' & tech == 'hybrid'),
         # Problem with solar in the US: only one observation for quintile 2
         !(iso2 == 'US' & tech == 'solar')
  )


# Function to obtain a fit for a particular iso2 x technology x quintile triplet
nlsTechnology <- function(dt, max_penetration, growth_rate, inflexion_year) {
  .tech <- dt$tech[1]; .quintile <- dt$quintile[1]; .iso2 <- dt$iso2[1]

  target_2100 <- tail(props_long[iso2 == .iso2 & quintile == .quintile & tech == .tech], 1)
  target_2100$year <- 2100
  target_2100$share <- max_penetration
  dt <- rbindlist(list(dt, target_2100))

  dt %>%
    minpack.lm::nlsLM(share ~ max_penetration / (1 + exp(-growth_rate * (year - inflexion_year))), .,
                      start = list(
                        max_penetration = max_penetration,
                        growth_rate = growth_rate,
                        inflexion_year = inflexion_year
                      ),
                      lower = c('max_penetration' = 0, 'growth_rate' = 0, 'inflexion_year' = 1900),
                      upper = c('max_penetration' = max_penetration, 'growth_rate' = 0.15, 'inflexion_year' = 2500),
                      control = nls.lm.control(maxfev = 4000, maxiter = 1024)
    )
}


# Market penetration fits across all countries, technologies and quintiles
tech_penetration_fits <- sapply(unique(props_long$iso2) %>% sort, function(.iso2) {
  print(.iso2)
  lapply(unique(props_long[iso2 == .iso2]$quintile) %>% sort, function(.quintile) {
    print(.quintile)
    sapply(unique(props_long[iso2 == .iso2]$tech), function(.tech) {
      print(.tech)
      props_long[iso2 == .iso2 & quintile == .quintile & tech == .tech] %>%
        nlsTechnology(
          max_penetration = max(max_penetration[tech == .tech]$max_share,
                                max(props_long[iso2 == .iso2 & quintile == .quintile & tech == .tech]$share)),
          growth_rate = 0.1, inflexion_year = 2025
        )
    }, simplify = FALSE, USE.NAMES = TRUE)
  })
}, simplify = FALSE, USE.NAMES = TRUE)


# Plot results for a given country
plotCountry <- function(.iso2) {
  lapply(unique(props_long[iso2 == .iso2]$tech), function(.tech) {
    lapply(1:5, function(.quintile) {
      data.table(year = 2000:2100, quintile = .quintile) %>%
        mutate(predicted_share = predict(tech_penetration_fits[[.iso2]][[.quintile]][[.tech]],
                                         newdata = list(year = year)),
               tech = .tech)
    }) %>% rbindlist
  }) %>% rbindlist %>%
    mutate(iso2 = .iso2) %>%
    left_join(props_long, by = c('iso2', 'year', 'tech', 'quintile')) %>%
    mutate(quintile = str_c('Q', quintile)) %>%
    ggplot(aes(year, predicted_share, color = tech)) +
    geom_line() +
    geom_point(aes(year, share, color = tech), size = 1) +
    facet_grid(vars(quintile), vars(tech), switch = "y") +
    scale_y_continuous(labels = scales::label_percent()) +
    scale_x_continuous(breaks = c(2000, 2050, 2100)) +
    xlab('') + ylab('Penetration rate') +
    ggtitle(countrycode(.iso2, 'iso2c', 'country.name')) +
    theme_minimal() +
    theme(legend.position = 'none',
          text = element_text(size = 14))
}

plotCountry('DE')
plotCountry('GB')
plotCountry('US')

## Regression tables
tech_penetration_fits[['DE']] %>% `names<-`(str_c("Quintile ", 1:5)) %>%
  modelsummary(stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
               shape = 'rbind')

tech_penetration_fits[['GB']] %>% `names<-`(str_c("Quintile ", 1:5)) %>%
  modelsummary(stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
               shape = 'rbind')

tech_penetration_fits[['US']] %>% `names<-`(str_c("Quintile ", 1:5)) %>%
  modelsummary(stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
               shape = 'rbind')


```

### Version 2 - Country and Technology Specifics
In the baseline model  Model aims to reach exogenously government provided diffusion target for each technology (e.g.40% for housing retrofit vs 60% for EVs) in each country. Desk research to find target. 

General findings: Digital technologies (internet and smartphone) are earliest adopters and have highest penetration. Q5 generally leads adoption. Energy technologies are much later. Renovation has a very slow increase compared to technologies. 

Now, to improve with technology and country specific policies. 

GERMANY
Solar -- goal of 80-95% solar installation on homes by 2050
Internat -- goal of 100% by 2030 in digitalstrategie deutschland
smartphone --- Digitalstrategie Deutschland state 2030 everyone should have
skills -- EU have a 80% by 2030 goal in EU Digital Decade Strategy
hybrid/ev -- they have a 100% new sales goal, 
commute --2030 Federal Infrastructure Plan lists expected growth to 2030 to be 20% increase  
fuel switching-- 60% RE energy consumption across the board by 2050 in Buildings Strategy, Energieweshel has focus on heat pumps and district heating at 45% of all heating solutions
renovation -- 1% annual to 2% annual in the Buildings Strategy so 45% renovation in 2019 would be 83.14% renovated in 2050

UNITED STATES
solar- EIA predicts 44% residential pv in 2050, but not official goal. 
smartphones-- no specific target but say they want to be 100% by last year. 
skills -- no specific law, but expect to have 90% of jobs needing digital skills in 2030. 
internet-- IIJA has a universal broadband access goal. 
hybrid-- have a ze sales goal for 2035, assuming current sharee of road, it would be around 75% in 2050. 
commute-- IIJA expands public transit, but not federal goals. Given urban goals around 70-80% by 2050 and rural goals in the 40% range, selecting 50% overall. 
fuel-switch-- No federal regulation or goal, but variety of states have goals on electrification that aveage to around 50% 
renovation-- DOE and IRA have investment in retrofitting but no specific goals. State level interest is high so around 70% 

UNITED KINGDOM
solar- no clear number for residential, but 85GW goal for 2050. Based on today having 17.2GW and 30% of that being residential (1.4M hh), it would be about 20GW in 2050 and an increase of 34% for 10K hh. This figure would increase signficiantly if New Homes (solar generation bill) goes into force
smartphones-- no specific target but say they want to be 100% already. 
skills- Digital skills strategy has a 100% goal and Digital Future roadmap commits to enhancing digital skills for 95% of adulst, with no date assigned. 
internet-- global broadband by 2030 target. 
hybrid-- have a 100% new cars goal for 2035, assuming average lifespan of car and curernt hybrid infiltration, that would be around 85% in 2050
commute-- Gear Change goal is 50% journeys to be walked or cycled in 2030. 
fuel-switch--UK NZ and ECO obligations both look for NZ and heating upgrade starting in 2026, homes will be switchign at higher rate (70%) 
renovation-- EEStrategy asks for 100% retrofitting, so likely around 80% in 2050. 

```{r s-curve improved, include = FALSE}
# Configure maximum penetration hypothesis
max_penetration <- tribble(
        ~iso2, ~tech, ~max_share,
        # Germany
        'DE', 'solar', 0.95,
        'DE', 'smartphone', 1.0,
        'DE', 'skills', 0.8,
        'DE', 'internet', 1.0,
        'DE', 'hybrid', 1.0,
        'DE', 'commute', 0.3,
        'DE', 'fuel_switch', 0.80,
        'DE', 'renovation', 0.83,
        # US 
        'US', 'solar', 0.44,
        'US', 'smartphone', 1.0,
        'US', 'skills', 0.90,
        'US', 'internet', 1.0,
        'US', 'hybrid', 0.75,
        'US', 'commute', 0.5,
        'US', 'fuel_switch', 0.7,
        'US', 'renovation', 0.70,
        # GB 
        'GB', 'solar', 0.34,
        'GB', 'smartphone', 1.0,
        'GB', 'skills', 0.95,
        'GB', 'internet', 1.0,
        'GB', 'hybrid', 0.85,
        'GB', 'commute', 0.5,
        'GB', 'fuel_switch', 0.7,
        'GB', 'renovation', 0.80
) %>% as.data.table

# Aggregate all 3 countries to the quintile level,
# and combine them to a single data.table
props_long <- bind_rows(
        deu_props %>% mutate(iso2 = "DE"),
        uk_props  %>% mutate(iso2 = "GB"),
        us_props  %>% mutate(iso2 = "US")
) %>%
        mutate(quintile = (weighted_decile - 1) %/% 2 + 1) %>% 
        group_by(iso2, year, quintile) %>%
        summarise(across(ends_with('prop'), ~weighted.mean(., n_total, na.rm = TRUE))) %>% 
        as.data.table %>%
        melt(id.vars = c('iso2', 'year', 'quintile'), 
             value.name = 'share', 
             variable.name = 'tech') %>%
        mutate(tech = str_remove(tech, '_prop')) %>%
        filter(share != 0, !is.na(share))

# Function to obtain a fit for a particular iso2 x technology x quintile triplet
nlsTechnology <- function(dt, max_penetration, growth_rate, inflexion_year) {
        .tech <- dt$tech[1]
        .quintile <- dt$quintile[1]
        .iso2 <- dt$iso2[1]
        
        country_max <- max_penetration
        
        target_2050 <- tail(props_long[iso2 == .iso2 & 
                                               quintile == .quintile & 
                                               tech == .tech], 1)
        target_2050$year <- 2050
        target_2050$share <- country_max
        dt <- rbindlist(list(dt, target_2050))
        
        dt %>%
                minpack.lm::nlsLM(
                        share ~ country_max / (1 + exp(-growth_rate * (year - inflexion_year))),
                        .,
                        start = list(
                                country_max = country_max,
                                growth_rate = growth_rate,
                                inflexion_year = inflexion_year
                        ),
                        lower = c('country_max' = 0, 
                                  'growth_rate' = 0, 
                                  'inflexion_year' = 1900),
                        upper = c('country_max' = country_max, 
                                  'growth_rate' = 0.15, 
                                  'inflexion_year' = 2500),
                        control = nls.lm.control(maxfev = 4000, maxiter = 1024)
                )
}

# Market penetration fits across all countries, technologies and quintiles
tech_penetration_fits <- sapply(unique(props_long$iso2) %>% sort, function(.iso2) {
        print(.iso2)
        lapply(unique(props_long[iso2 == .iso2]$quintile) %>% sort, function(.quintile) {
                print(.quintile)
                sapply(unique(props_long[iso2 == .iso2]$tech), function(.tech) {
                        print(.tech)
                        props_long[iso2 == .iso2 & 
                                           quintile == .quintile & 
                                           tech == .tech] %>%
                                nlsTechnology(
                                        max_penetration = max_penetration[iso2 == .iso2 & 
                                                                                  tech == .tech]$max_share,
                                        growth_rate = 0.1, 
                                        inflexion_year = 2025
                                )
                }, simplify = FALSE, USE.NAMES = TRUE)
        })
}, simplify = FALSE, USE.NAMES = TRUE)

# Plot results for a given country
plotCountry <- function(.iso2) {
        lapply(unique(props_long[iso2 == .iso2]$tech), function(.tech) {
                lapply(1:5, function(.quintile) {
                        data.table(year = 2000:2050, quintile = .quintile) %>%
                                mutate(
                                        predicted_share = predict(
                                                tech_penetration_fits[[.iso2]][[.quintile]][[.tech]],
                                                newdata = list(year = year)
                                        ),
                                        tech = .tech
                                )
                }) %>% rbindlist
        }) %>% rbindlist %>%
                mutate(iso2 = .iso2) %>%
                left_join(props_long, 
                          by = c('iso2', 'year', 'tech', 'quintile')) %>%
                mutate(quintile = str_c('Q', quintile)) %>%
                ggplot(aes(year, predicted_share, color = tech)) +
                geom_line() +
                geom_point(aes(year, share, color = tech), size = 1) +
                facet_grid(vars(quintile), vars(tech), switch = "y") +
                scale_y_continuous(labels = scales::label_percent()) +
                scale_x_continuous(breaks = seq(2000, 2050, by = 10)) +
                xlab('') + 
                ylab('Penetration rate') +
                ggtitle(countrycode(.iso2, 'iso2c', 'country.name')) +
                theme_minimal() +
                theme(legend.position = 'none',
                      text = element_text(size = 14))
}

# More robust function to extract NLS results
extract_nls_summary_robust <- function(model_list) {
        result_df <- data.frame()
        
        for(q_idx in seq_along(model_list)) {
                quintile_models <- model_list[[q_idx]]
                
                for(tech_name in names(quintile_models)) {
                        model <- quintile_models[[tech_name]]
                        
                        if(inherits(model, "nls")) {
                                tryCatch({
                                        # Just extract coefficients without standard errors
                                        coefs <- coef(model)
                                        
                                        temp_df <- data.frame(
                                                quintile = paste("Quintile", q_idx),
                                                technology = tech_name,
                                                parameter = names(coefs),
                                                estimate = unname(coefs)
                                        )
                                        
                                        result_df <- rbind(result_df, temp_df)
                                }, error = function(e) {
                                        warning(paste("Could not extract results for quintile", 
                                                      q_idx, "technology", tech_name))
                                })
                        }
                }
        }
        
        return(result_df)
}

# Function to create a formatted table for one country
create_country_table <- function(summary_df, country_name) {
        # Reshape the data to wide format
        summary_wide <- reshape2::dcast(
                summary_df,
                quintile + parameter ~ technology,
                value.var = "estimate"
        )
        
        # Print the results
        print(paste("Results for", country_name))
        print(summary_wide)
        
        return(summary_wide)
}

# Generate plots
plotCountry('DE')
plotCountry('GB')
plotCountry('US')

# Extract and display results for each country
DE_summary <- extract_nls_summary_robust(tech_penetration_fits[['DE']])
GB_summary <- extract_nls_summary_robust(tech_penetration_fits[['GB']])
US_summary <- extract_nls_summary_robust(tech_penetration_fits[['US']])

# Create tables
DE_table <- create_country_table(DE_summary, "Germany")
GB_table <- create_country_table(GB_summary, "United Kingdom")
US_table <- create_country_table(US_summary, "United States")

```
### Version 3 - Paper Version
```{r s-curve take three, echo=FALSE}
# Configure maximum penetration hypothesis
max_penetration <- tribble(
        ~iso2, ~tech, ~max_share,
        # Germany
        'DE', 'solar', 0.95,
        'DE', 'smartphone', 1.0,
        'DE', 'skills', 0.8,
        'DE', 'internet', 1.0,
        'DE', 'hybrid', 1.0,
        'DE', 'commute', 0.3,
        'DE', 'fuel_switch', 0.80,
        'DE', 'renovation', 0.83,
        # US 
        'US', 'solar', 0.44,
        'US', 'smartphone', 1.0,
        'US', 'skills', 0.90,
        'US', 'internet', 1.0,
        'US', 'hybrid', 0.75,
        'US', 'commute', 0.5,
        'US', 'fuel_switch', 0.7,
        'US', 'renovation', 0.70,
        # GB 
        'GB', 'solar', 0.34,
        'GB', 'smartphone', 1.0,
        'GB', 'skills', 0.95,
        'GB', 'internet', 1.0,
        'GB', 'hybrid', 0.85,
        'GB', 'commute', 0.5,
        'GB', 'fuel_switch', 0.7,
        'GB', 'renovation', 0.80
) %>% as.data.table

# Create filtered version of props_long without German hybrid
props_long_filtered <- props_long %>%
    filter(!(iso2 == "DE" & tech == "hybrid"),
           ! (iso2 == "DE" & tech == "renovation" & year == 2019))

# Function to obtain a fit for a particular iso2 x technology x quintile triplet
nlsTechnology <- function(dt, max_penetration, growth_rate, inflexion_year) {
        .tech <- dt$tech[1]
        .quintile <- dt$quintile[1]
        .iso2 <- dt$iso2[1]
        
        country_max <- max_penetration

        # Handle smartphone saturation for US, UK, and DE
        if(.tech == "smartphone" && .iso2 %in% c("US", "GB", "DE") && min(dt$share) > 0.8) {
            if(.iso2 == "DE") {
                # Special handling for German smartphones
                growth_rate <- 0.3  # Faster growth
                inflexion_year <- 2005  # Earlier inflection
            } else {
                # US and UK handling
                growth_rate <- 0.2
                inflexion_year <- min(dt$year) - 5
            }
            # Use observed maximum as lower bound for all saturated cases
            country_max <- min(max(dt$share), country_max)
        }
        
        target_2050 <- tail(props_long_filtered[iso2 == .iso2 &
                                              quintile == .quintile &
                                              tech == .tech], 1)
        target_2050$year <- 2050
        target_2050$share <- country_max
        
        dt <- rbindlist(list(dt, target_2050))
        
        dt %>%
                minpack.lm::nlsLM(
                        share ~ country_max / (1 + exp(-growth_rate * (year - inflexion_year))),
                        .,
                        start = list(
                                country_max = country_max,
                                growth_rate = growth_rate,
                                inflexion_year = inflexion_year
                        ),
                        lower = c('country_max' = 0,
                                   'growth_rate' = 0,
                                   'inflexion_year' = 1980),
                        upper = c('country_max' = country_max,
                                   'growth_rate' = 0.15,
                                   'inflexion_year' = 2500),
                        control = nls.lm.control(maxfev = 4000, maxiter = 1024)
                )
}

# Market penetration fits across all countries, technologies and quintiles
tech_penetration_fits <- sapply(unique(props_long_filtered$iso2) %>% sort, function(.iso2) {
        print(.iso2)
        lapply(unique(props_long_filtered[iso2 == .iso2]$quintile) %>% sort, function(.quintile) {
                print(.quintile)
                sapply(unique(props_long_filtered[iso2 == .iso2]$tech), function(.tech) {
                        print(.tech)
                        props_long_filtered[iso2 == .iso2 &
                                            quintile == .quintile &
                                            tech == .tech] %>%
                                nlsTechnology(
                                        max_penetration = max_penetration[iso2 == .iso2 &
                                                                                   tech == .tech]$max_share,
                                        growth_rate = 0.1, 
                                        inflexion_year = 2025
                                )
                }, simplify = FALSE, USE.NAMES = TRUE)
        })
}, simplify = FALSE, USE.NAMES = TRUE)

# Plot results for a given country
plotCountry <- function(.iso2) {
        lapply(unique(props_long_filtered[iso2 == .iso2]$tech), function(.tech) {
                lapply(1:5, function(.quintile) {
                        data.table(year = 2000:2050, quintile = .quintile) %>%
                                mutate(
                                        predicted_share = predict(
                                                tech_penetration_fits[[.iso2]][[.quintile]][[.tech]],
                                                newdata = list(year = year)
                                        ),
                                        tech = .tech
                                )
                }) %>% rbindlist
        }) %>% rbindlist %>%
                mutate(iso2 = .iso2) %>%
                left_join(props_long_filtered,
                           by = c('iso2', 'year', 'tech', 'quintile')) %>%
                mutate(quintile = str_c('Q', quintile)) %>%
                ggplot(aes(year, predicted_share, color = tech)) +
                geom_line() +
                geom_point(aes(year, share, color = tech), size = 1) +
                facet_grid(vars(quintile), vars(tech), switch = "y") +
                scale_y_continuous(labels = scales::label_percent()) +
                scale_x_continuous(breaks = seq(2000, 2050, by = 10)) +
                xlab('') + 
                ylab('Penetration rate') +
                ggtitle(countrycode(.iso2, 'iso2c', 'country.name')) +
                theme_minimal() +
                theme(legend.position = 'none',
                      text = element_text(size = 14))
}

# More robust function to extract NLS results
extract_nls_summary_robust <- function(model_list) {
        result_df <- data.frame()
        
        for(q_idx in seq_along(model_list)) {
                quintile_models <- model_list[[q_idx]]
                
                for(tech_name in names(quintile_models)) {
                        model <- quintile_models[[tech_name]]
                        
                        if(inherits(model, "nls")) {
                                tryCatch({
                                        # Just extract coefficients without standard errors
                                        coefs <- coef(model)
                                        
                                        temp_df <- data.frame(
                                                quintile = paste("Quintile", q_idx),
                                                technology = tech_name,
                                                parameter = names(coefs),
                                                estimate = unname(coefs)
                                        )
                                        
                                        result_df <- rbind(result_df, temp_df)
                                }, error = function(e) {
                                        warning(paste("Could not extract results for quintile",
                                                       q_idx, "technology", tech_name))
                                })
                        }
                }
        }
        
        return(result_df)
}

# Function to create a formatted table for one country
create_country_table <- function(summary_df, country_name) {
        # Reshape the data to wide format
        summary_wide <- reshape2::dcast(
                summary_df,
                quintile + parameter ~ technology,
                value.var = "estimate"
        )
        
        # Print the results
        print(paste("Results for", country_name))
        print(summary_wide)
        
        return(summary_wide)
}

# Generate plots
plotCountry('DE')
plotCountry('GB')
plotCountry('US')

# Extract and display results for each country
DE_summary <- extract_nls_summary_robust(tech_penetration_fits[['DE']])
GB_summary <- extract_nls_summary_robust(tech_penetration_fits[['GB']])
US_summary <- extract_nls_summary_robust(tech_penetration_fits[['US']])

# Create tables
DE_table <- create_country_table(DE_summary, "Germany")
GB_table <- create_country_table(GB_summary, "United Kingdom")
US_table <- create_country_table(US_summary, "United States")

## Regression tables
tech_penetration_fits[['DE']] %>% `names<-`(str_c("Quintile ", 1:5)) %>%
  modelsummary(stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
               shape = 'rbind',
               output = 'tech_penetration_fits_DE.tex')

tech_penetration_fits[['GB']] %>% `names<-`(str_c("Quintile ", 1:5)) %>%
  modelsummary(stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
               shape = 'rbind')

tech_penetration_fits[['US']] %>% `names<-`(str_c("Quintile ", 1:5)) %>%
  modelsummary(stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
               shape = 'rbind')



```
Export for Overleaf
```{r Latex version, echo=FALSE}
output_dir <- "/Users/cassandraetter-wenzel/Documents/GitHub/EDITS/Output"

create_model_summary <- function(models, country, output_path) {
    # Ensure models are named and valid
    if (is.null(names(models))) {
        names(models) <- str_c("Quintile ", seq_along(models))
    }

    # Filter only valid models
    valid_models <- models[sapply(models, function(x) inherits(x, "nls"))]

    if (length(valid_models) == 0) {
        stop("Error: No valid models found.")
    }

    print(names(valid_models))  # Ensure valid names

    # Manually extract model summaries for nls models
    model_summaries <- lapply(valid_models, function(model) {
        summary(model)$coefficients
    })

    # Convert the summaries into a format modelsummary can use
    model_summary_df <- do.call(rbind, model_summaries)
    rownames(model_summary_df) <- names(valid_models)  # Set names of models

    # Write summary to LaTeX
    modelsummary(
        model_summary_df,
        stars = c('*' = 0.1, '**' = 0.05, '***' = 0.01),
        shape = 'rbind',
        output = "latex_tabular",
        booktabs = TRUE,
        fmt = "%.3f",
        table_format = "latex",
        title = paste("Technology Adoption Parameters:", country),
        gof_omit = 'AIC|BIC|Log|fin|Conv',
        coef_map = c(
          "country_max" = "Maximum Penetration",
          "growth_rate" = "Growth Rate",
          "inflexion_year" = "Inflection Year"
        ),
        note = c("* p < 0.1, ** p < 0.05, *** p < 0.01"),
        tex_wrap = c("begin{table}[!htbp]", "\\centering", "\\small", "\\end{table}"),
        add_rows = NULL
    ) %>% readr::write_file(output_path)
}

# Germany
create_model_summary(tech_penetration_fits[['DE']], "Germany", 
                    file.path(output_dir, "tech_penetration_fits_DE.tex"))

# UK
create_model_summary(tech_penetration_fits[['GB']], "United Kingdom", 
                    file.path(output_dir, "tech_penetration_fits_GB.tex"))

# US
create_model_summary(tech_penetration_fits[['US']], "United States", 
                    file.path(output_dir, "tech_penetration_fits_US.tex"))

```

General Analysis could also be shown like this
```{r model summary, echo = FALSE}
# Function to create table for one country
create_country_table <- function(country_summary, country_name) {
  # Format table
  pub_table <- country_summary %>%
    # Round estimates to 3 decimal places
    mutate(
      estimate = round(estimate, 3)
    ) %>%
    # Pivot wider to get technologies as columns
    pivot_wider(
      id_cols = c(quintile, parameter),
      names_from = technology,
      values_from = estimate
    ) %>%
    # Arrange by parameter and quintile
    arrange(parameter, quintile)
  
  # Create formatted table
  formatted_table <- pub_table %>%
    kbl(format = "html",
        caption = paste("NLS Regression Results for", country_name),
        align = "c",
        escape = FALSE) %>%
    kable_styling(bootstrap_options = c("striped", "hover"),
                  full_width = FALSE) %>%
    # Group by parameters instead of countries
    pack_rows(index = table(pub_table$parameter)) %>%
    add_header_above(c("Quintile Groups" = 2, 
                      "Technologies" = ncol(pub_table) - 2)) %>%
    footnote(
      general = paste("Note: Values show parameter estimates from NLS regression.",
                     "Parameters shown are country_max (maximum penetration),",
                     "growth_rate (speed of adoption), and",
                     "inflexion_year (year of maximum growth).")
    )
  
  return(formatted_table)
}

# Create separate tables for each country
germany_table <- create_country_table(DE_summary, "Germany")
uk_table <- create_country_table(GB_summary, "United Kingdom")
us_table <- create_country_table(US_summary, "United States")

# Return all tables
list(
  germany = germany_table,
  uk = uk_table,
  us = us_table
)

```
### Country combined projections by Q1-Q5

Perhaps most interesting is looking at the technologies by quintile accross the three Countries.
```{r combined quintile view, echo = FALSE}
# Function to create plot for a specific quintile
create_quintile_plot <- function(quintile_num) {
  # Extract predictions for the specified quintile
  quintile_predictions <- lapply(c('DE', 'GB', 'US'), function(.iso2) {
    lapply(unique(props_long_filtered[iso2 == .iso2]$tech), function(.tech) {
      data.table(
        year = 2000:2050,
        predicted_share = predict(
          tech_penetration_fits[[.iso2]][[quintile_num]][[.tech]],
          newdata = list(year = 2000:2050)
        ),
        tech = .tech,
        iso2 = .iso2
      )
    }) %>% rbindlist
  }) %>% rbindlist

  # Create the plot
  p <- ggplot(quintile_predictions, aes(x = year, y = predicted_share, color = iso2)) +
    geom_line() +
    facet_wrap(~tech, scales = "free_y", ncol = 3) +
    scale_y_continuous(labels = scales::label_percent()) +
    scale_x_continuous(breaks = seq(2000, 2050, by = 10)) +
    scale_color_manual(
      values = c("DE" = "#1f77b4", "GB" = "#ff7f0e", "US" = "#2ca02c"),
      labels = c("DE" = "Germany", "GB" = "United Kingdom", "US" = "United States")
    ) +
    labs(
      x = "Year",
      y = "Adoption Rate",
      color = "Country",
      title = paste("Technology Adoption Rates by Country (Q", quintile_num, ")", sep = ""),
      subtitle = "Predicted adoption trajectories for different technologies"
    ) +
    theme_minimal() +
    theme(
      text = element_text(size = 12),
      strip.text = element_text(size = 12, face = "bold"),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    )
  
  ggsave(
    filename = file.path(output_dir, paste0("tech_adoption_q", quintile_num, ".png")),
    plot = p,
    width = 12,
    height = 8,
    dpi = 300
  )
  
  return(p)
}
plots <- lapply(1:5, create_quintile_plot)

library(patchwork)
combined_plot <- (plots[[1]] / plots[[2]] / plots[[3]] / plots[[4]] / plots[[5]])
ggsave(
  filename = file.path(output_dir, "tech_adoption_all_quintiles.pdf"),
  plot = combined_plot,
  width = 15,
  height = 40,
  device = cairo_pdf
)
```

# Step 3: 2050 Diffusion Impacts

Step 3 Estimate how distributional impacts of technology uptake affect final energy costs for household across time. Use survey data to estimate how final energy costs increase/decrease with different technologies including in a household (holding all other variables constant) in 2020. Then, hold energy prices static (very large assumption) to see how technology diffusion affects final energy costs for residential sector at decile level in 2050.

I first get the average amount spent on energy, then roughly estimate how much is on gas vs electric, then get historical prices and convert amount spent to gas and electric to consumption. This is easier for some data (ex. US ) than others. For the US I use state level prices from SEDS. To check my consumption comparison against it, I compare my estimates against the ACS. 

For Germany I use national and compare against XX. 
For the UK, I use national and compare against XX. 


First See what to compare against with international data. 
```{r IEA energy price data, echo=FALSE}
# IEA Household Final Annual Average -- USD/MWh 
us_prices <- tibble(
  country = "US",
  year = 2011:2021,
  elec_price = c(117.169, 118.785, 121.237, 125.179, 126.513, 125.489, 128.871, 128.693, 130.144, 131.959, 137.237),  # UDD per MWh
  gas_price = c(36.415, 35.322, 34.02, 36.134, 34.233, 33.166, 35.874, 34.555, 34.542, 35.773, 40.532)
)



# IEA USD/ MWh in PPP
deu_prices <- tibble(
  country = "DEU",
  year = 2011:2021,
  elec_price = c(320.702, 334.775, 376.879, 387.223, 378.925, 395.026, 409.247, 407.031, 397.152, 409.629, 439.238),  
  gas_price = c(84.405, 89.26, 91.983, 92.8, 90.775, 91.197, 89.122, 88.807, 90.435, 92.317, 97.579)
)


# IEA Household EPT FD USD/MWh in PPP
uk_prices <- tibble(
  country = "UK",
  year = 2011:2021,
  elec_price = c(184.896, 195.878, 211.692, 220.36, 217.323, 218.736, 232.745, 249.677, 272.748, 276.079, 303.6),  
  gas_price = c(59.501, 64.918, 69.562, 73.303, 69.636, 64.073, 62.914, 64.889, 66.345, 60.626, 62.688)

)


# Combine all data
historical_prices <- bind_rows(
  us_prices,
  deu_prices,
  uk_prices
)

# Function to merge historical prices with country-specific datasets
merge_prices <- function(country_data, country_name) {
  country_data %>%
    left_join(
      historical_prices %>% filter(country == country_name),
      by = "year"
    )
}

# Create visualizations for available data
create_price_plot <- function(data) {
  ggplot(data %>% filter(!is.na(elec_price) & !is.na(gas_price)), aes(x = year)) +
    geom_line(aes(y = elec_price, color = country, linetype = "Electricity"), size = 1) +
    geom_line(aes(y = gas_price, color = country, linetype = "Natural Gas"), size = 1) +
    theme_minimal() +
    labs(
      title = "Historical Energy Prices (2011-2021)",
      subtitle = "By Country USD/MWh",
      y = "$/MWh",
      color = "Country",
      linetype = "Energy Type",
      caption = "Sources: IEA Energy Price Data Household Final Residential (2011-2021)"
    ) +
    scale_color_viridis_d() +
    theme(
      legend.position = "bottom",
      strip.text = element_text(size = 12, face = "bold") 
    ) + 
    ylim(0, NA)
}


# Function to print data summary with source documentation
print_summary <- function(data) {
  for (country_name in unique(data$country)) {
    country_data <- data %>% 
      filter(country == country_name, !is.na(elec_price))
    
    if (nrow(country_data) > 0) {
      cat(sprintf("\nSummary for %s:\n", country_name))
      print(summary(country_data[c("elec_price", "gas_price")]))
      cat("Data source: ", switch(country_name,
                                "US" = "IEA",
                                "DEU" = "IEA",
                                "UK" = "IEA"),
          "\n")
    }
  }
}

historical_prices <- bind_rows(us_prices, deu_prices, uk_prices)
price_plot <- create_price_plot(historical_prices)
print_summary(historical_prices)
print(price_plot)
```

The US data from IEA does not work, I need state prices-- otherwise it looks like there was no volatility for gas in particular, when that is very much not the case. Focus on US first with state elec and gas prices. 

## US Consumption Creation 

```{r US energy Consumption Calculation, echo = FALSE}
psid_states <- fread("~/Documents/GitHub/EDITS/Data/psid_states.csv") %>%
  left_join(data.table(StateCode = state.abb, us_state = state.name), by ='us_state') %>%
  mutate(StateCode = ifelse(is.na(StateCode), 'DC', StateCode))

seds <- fread("~/Documents/GitHub/EDITS/Data/Complete_SEDS.csv")[Year %in% unique(US$year) & MSN %chin% c("ESRCD", "NGRCD")] %>%
  left_join(psid_states, by = 'StateCode') %>%
  mutate(v = c("ESRCD" = "elec_price", "NGRCD" = "gas_price")[MSN],
         psid_code = ifelse(is.na(psid_code), -1, psid_code),
         us_state = ifelse(is.na(us_state), 'US', us_state)) %>%
  rename(year = Year) %>%
  dcast(psid_code + us_state + year ~ v, value.var = 'Data') 

US_energy <- US %>% 
  mutate(hh_state = ifelse(is.na(hh_state), -1, hh_state)) %>%
  left_join(seds, by = c('hh_state' = 'psid_code', 'year')) %>%
  mutate(elec_a_mmbtu = elec_a_exp / elec_price,
         gas_a_mmbtu = gas_a_exp / gas_price, 
         elec_a_mwh = elec_a_mmbtu * 0.29307107,
         gas_a_mwh = gas_a_mmbtu * 0.29307107)

US_energy_by_decile_MWH <- US_energy %>%
  as.data.table %>% 
  dplyr::select(year, weighted_decile, weight, elec_a_mwh, gas_a_mwh) %>%
  melt(id.vars = c('year', 'weighted_decile', 'weight')) %>%
  mutate(variable = str_remove(variable, '_a_mwh')) %>%
  .[, .(value = weighted.mean(value, weight, na.rm = TRUE),
        weight = sum(weight)),
    by = .(year, weighted_decile, variable)] %>%
  arrange(year, weighted_decile, variable)

US_energy_by_decile_MWH %>%
  ggplot(aes(year, value, color = factor(weighted_decile))) +
  geom_line() +
  ylim(0, NA) +
  facet_wrap(~variable) +
  theme_minimal() + 
  labs(
    title = "Average US HH Consumption (2011 - 2021)", 
    y = "Annual Energy Consumption (MWh)") +
    theme(legend.position = "bottom",
          strip.text = element_text(size = 12, face = "bold")
      
  )
ggsave("~/Documents/GitHub/EDITS/Output/US_consumption.png", width = 10, height = 7, dpi = 300)
### Average for all deciles 
US_energy_by_decile_MWH[, .(value = weighted.mean(value, weight, na.rm = TRUE)), by = .(year, variable)] %>%
  ggplot(aes(year, value)) +
  geom_line() +
  facet_wrap(~variable) +
  ylim(0, NA) +
  theme_minimal() + 
  labs(
    title = "Average US Energy Consumption (2011- 2021)", 
    y = "Household Energy Consumption (MWh)") + 
  theme(
    legend.position = "bottom",
    strip.text = element_text(size = 12, face = "bold")
  )


as.data.table(US %>% mutate(gas_a_exp = ifelse(is.na(gas_a_exp), 0, gas_a_exp), elec_a_exp = ifelse(is.na(elec_a_exp), 0, elec_a_exp)))[, .(tot_energy = weighted.mean(elec_a_exp + gas_a_exp, weight, na.rm = TRUE)), by = year]

## Compare with ACS Consumption data 
acs <- data.table(
  year = seq(2011, 2021, 2),
  acs_energy = c(1843, 1815, 1881, 1801, 1888, 1998)
)

US_consumption <- US_energy %>%
  group_by(weighted_decile, year) %>%
  summarise(
    baseline_elec = weighted.mean(elec_a_mwh, w = weight, na.rm = TRUE),
    baseline_gas = weighted.mean(gas_a_mwh, w = weight, na.rm = TRUE),
    elec_price = weighted.mean(elec_price, w = weight, na.rm = TRUE),
    gas_price = weighted.mean(gas_price, w = weight, na.rm = TRUE) 
  )


```

## DEU Consumption Creation 
biannual data from Eurstat nrg_pc_204 on prices for electricity and gas. Can add in heating and cooling degree days. 
```{r deu consumption, echo = FALSE}

nrg_gas <-fread("~/Documents/GitHub/EDITS/Data/estat_nrg_pc_202_filtered_en.csv") %>%
        select(geo, TIME_PERIOD, currency, OBS_VALUE) %>%
        rename(gas_price = OBS_VALUE)

nrg <- fread("~/Documents/GitHub/EDITS/Data/estat_nrg_pc_204_filtered_en.csv") %>% 
        select(geo, TIME_PERIOD, currency, OBS_VALUE) %>%
        rename(elec_price = OBS_VALUE) %>%
        left_join(nrg_gas, by = c('geo', 'TIME_PERIOD', 'currency'))%>%
        separate(TIME_PERIOD, into = c("year", "semester"), sep = "-") %>%
        mutate(
                year = as.numeric(year),
                semester_num = if_else(semester == "S1", 1, 2)
        )

DEU_energy <- DEU %>%
        mutate(
                semester_num = if_else(hghmonth <= 6, 1, 2)
        ) %>%
        left_join(nrg, by = c("year", "semester_num")) %>%
        mutate(elec_a_mwh = elec_a_exp / (elec_price * 1000),
               gas_a_mwh = gas_a_exp / (gas_price * 1000))


DEU_energy_by_decile_MWH <- DEU_energy %>%
        as.data.table %>% 
        dplyr::select(year, weighted_decile, weight, elec_a_mwh, gas_a_mwh) %>%
        melt(id.vars = c('year', 'weighted_decile', 'weight')) %>%
        mutate(variable = str_remove(variable, '_a_mwh')) %>%
        .[, .(value = weighted.mean(value, weight, na.rm = TRUE),
              weight = sum(weight)),
          by = .(year, weighted_decile, variable)] %>%
        arrange(year, weighted_decile, variable)

DEU_energy_by_decile_MWH %>%
        ggplot(aes(year, value, color = factor(weighted_decile))) +
        geom_line() +
        ylim(0, NA) +
        facet_wrap(~variable) +
        theme_minimal() + 
        labs(
                title = "Average German HH Consumption (2011 - 2021)", 
                y = "Annual Energy Consumption (MWh)") +
        theme(legend.position = "bottom",
              strip.text = element_text(size = 12, face = "bold")
              
        )
ggsave("~/Documents/GitHub/EDITS/Output/DE_consumption.png", width = 10, height = 7, dpi = 300)
### Average for all deciles 

DEU_energy_by_decile_MWH[, .(value = weighted.mean(value, weight, na.rm = TRUE)), by = .(year, variable)] %>%
        ggplot(aes(year, value)) +
        geom_line() +
        facet_wrap(~variable) +
        ylim(0, NA) +
        theme_minimal() + 
        labs(
                title = "Average German Energy Consumption (2011- 2021)", 
                y = "Household Energy Consumption (MWh)") + 
        theme(
                legend.position = "bottom",
                strip.text = element_text(size = 12, face = "bold")
        )

DEU_consumption <- DEU_energy %>%
  group_by(weighted_decile, year) %>%
  summarise(
    baseline_elec = weighted.mean(elec_a_mwh, w = weight, na.rm = TRUE),
    baseline_gas = weighted.mean(gas_a_mwh, w = weight, na.rm = TRUE),
    elec_price = weighted.mean(elec_price, w = weight, na.rm = TRUE) * 1000,
    gas_price = weighted.mean(gas_price, w = weight, na.rm = TRUE) * 1000
  )

```
## UK Consumption Creation  

UK asks average annual household costs for electric, gas, and duel, so no seasonal adjustment but I can add in heating/ cooling degree days for consumption. 


```{r uk consumption creation, echo=FALSE}
uk_gas <-fread("~/Documents/GitHub/PhD/Data/uk_gas_prices.csv") %>%
        mutate(source = "Gas")

uk_elec <- fread("~/Documents/GitHub/PhD/Data/uk_elec_prices.csv") %>% 
        mutate(source = "Electricity")

region_mapping <- data.frame(
        region_name = c("London", "South East", "West Midlands", "East Midlands", "North West", 
                        "Wales", "North East", "Yorkshire and the Humber", "Scotland", "East of England", 
                        "Northern Ireland", "north east", "north west", "yorkshire and the humber", 
                        "east midlands", "South West", "west midlands", "south west", "east of england", 
                        "london", "south east", "wales", "scotland", "northern ireland"),
        region_code = c(7, 8, 5, 4, 2, 10, 1, 3, 11, 6, 12, 1, 2, 3, 4, 9, 5, 9, 6, 7, 8, 10, 11, 12)
)

UK_energy <- UK %>%
        left_join(region_mapping, by = c("gor_dv" = "region_name")) %>%
        rename(weight = hhdenui_xw, gor_dv_code = region_code) %>%
        mutate(payment = case_when(
                elecpay %in% c("A pre-payment (key/card or token) meter", "Fuel Direct scheme or direct from benefits") |
                        gaspay %in% c("A pre-payment (key/card or token) meter", "Fuel Direct scheme or direct from benefits") |
                        duelpay %in% c("A pre-payment (key/card or token) meter", "Fuel Direct scheme or direct from benefits") ~ "Prepayment",
                
                elecpay %in% c("A monthly bill (by direct debit or other means)", "A quarterly bill (by direct debit or other means)",
                               "A fixed amount each month by standing order", "An annual bill (by direct debit or other means)") |
                        gaspay %in% c("A monthly bill (by direct debit or other means)", "A quarterly bill (by direct debit or other means)",
                                      "A fixed amount each month by standing order", "An annual bill (by direct debit or other means)") |
                        duelpay %in% c("A monthly bill (by direct debit or other means)", "A quarterly bill (by direct debit or other means)",
                                       "A fixed amount each month by standing order", "An annual bill (by direct debit or other means)") ~ "Direct Debit",
                
                elecpay %in% c("Frequent cash payments (ie more frequent than once a month)", "Other", "It's included in the rent", "Staywarm scheme") |
                        gaspay %in% c("Frequent cash payments (ie more frequent than once a month)", "Other", "It's included in the rent", "Staywarm scheme") |
                        duelpay %in% c("Frequent cash payments (ie more frequent than once a month)", "Other", "It's included in the rent", "Staywarm scheme") ~ "Credit",
                
                elecpay %in% c("don't know", "inapplicable", "refusal", "missing") |
                        gaspay %in% c("don't know", "inapplicable", "refusal", "missing") |
                        duelpay %in% c("don't know", "inapplicable", "refusal", "missing") ~ "Overall",
                TRUE ~ "Unknown"
        )) 

UK_energy <- UK_energy %>%
        mutate(gor_dv_code = as.character(gor_dv_code), year = as.integer(year))

uk_elec <- uk_elec %>%
        mutate(gor_dv = as.character(gor_dv), year = as.integer(year))

uk_gas <- uk_gas %>%
        mutate(gor_dv = as.character(gor_dv), year = as.integer(year))

get_energy_price <- function(data, prices, energy_type) {
        data %>%
                left_join(prices, by = c("year", "gor_dv_code" = "gor_dv")) %>%
                mutate(
                        price = case_when(
                                energy_type == "electricity" & xpelecy > 1 & payment == "Credit" ~ `Credit: Unit cost (Pence per kWh)`,
                                energy_type == "electricity" & xpelecy > 1 & payment == "Direct Debit" ~ `Direct debit: Unit cost (Pence per kWh)`,
                                energy_type == "electricity" & xpelecy > 1 & payment == "Prepayment" ~ `Prepayment: Unit cost (Pence per kWh)`,
                                energy_type == "electricity" & xpelecy > 1 ~ `Overall: Unit cost (Pence per kWh)`,
                                
                                energy_type == "gas" & xpgasy > 1 & payment == "Credit" ~ `Credit: Unit cost (Pence per kWh)`,
                                energy_type == "gas" & xpgasy > 1 & payment == "Direct Debit" ~ `Direct debit: Unit cost (Pence per kWh)`,
                                energy_type == "gas" & xpgasy > 1 & payment == "Prepayment" ~ `Prepayment: Unit cost (Pence per kWh)`,
                                energy_type == "gas" & xpgasy > 1 ~ `Overall: Unit cost (Pence per kWh)`,
                                
                                TRUE ~ NA_real_
                        )
                ) %>%
                pull(price)
}

UK_energy <- UK_energy %>%
    mutate(
        elec_price = (get_energy_price(UK_energy, uk_elec, "electricity") * 1000) / 100,  
        gas_price = (get_energy_price(UK_energy, uk_gas, "gas") * 1000) / 100 
        )%>%         
    mutate(
        elec_a_mwh = (xpelecy / elec_price), 
        gas_a_mwh = (xpgasy / gas_price)     
    )


UK_energy_by_decile_MWH <- UK_energy %>%
        as.data.table %>% 
        dplyr::select(year, weighted_decile, weight, elec_a_mwh, gas_a_mwh) %>%
        melt(id.vars = c('year', 'weighted_decile', 'weight')) %>%
        mutate(variable = str_remove(variable, '_a_mwh')) %>%
        .[, .(value = weighted.mean(value, weight, na.rm = TRUE),
              weight = sum(weight)),
          by = .(year, weighted_decile, variable)] %>%
        arrange(year, weighted_decile, variable)

UK_energy_by_decile_MWH %>%
        ggplot(aes(year, value, color = factor(weighted_decile))) +
        geom_line() +
        ylim(0, NA) +
        facet_wrap(~variable) +
        theme_minimal() + 
        labs(
                title = "Average UK HH Consumption (2011 - 2021)", 
                y = "Annual Energy Consumption (MWh)") +
        theme(legend.position = "bottom",
              strip.text = element_text(size = 12, face = "bold")
              
        )
ggsave("~/Documents/GitHub/EDITS/Output/UK_consumption.png", width = 10, height = 7, dpi = 300)

UK_consumption <- UK_energy %>%
  group_by(weighted_decile, year) %>%
  summarise(
    baseline_elec = weighted.mean(elec_a_mwh, w = weight, na.rm = TRUE),
    baseline_gas = weighted.mean(gas_a_mwh, w = weight, na.rm = TRUE),
    elec_price = weighted.mean(elec_price, w = weight, na.rm = TRUE),
    gas_price = weighted.mean(gas_price, w = weight, na.rm = TRUE) 
  )
```


```{r coefficient cleaning, include = FALSE}
UK_energy$solar1 <- case_when(
  UK_energy$solar1 %in% c("Yes - seriously considering", "yes - seriously considering", "Yes - fitted", "yes - fitted") ~ 1,  # Yes
  UK_energy$solar1 %in% c("No", "no", "Not applicable / rented accommodation", "spontaneous not applicable/rented accommodation", "Considered in the past and rejected", "considered in the past and rejected", "don't know", "refusal", "missing") ~ 2,  # No
  TRUE ~ NA_real_  # For any other unrecognized values, set to NA
)

UK_energy$tenure_dv <- case_when(
  UK_energy$ncars > 10 ~ NA_character_,  # If ncars > 10, set tenure_dv to NA (character)
  UK_energy$tenure_dv %in% c("Owned with mortgage", "Owned outright") ~ "Owned",  # Owned
  UK_energy$tenure_dv %in% c("Rented private unfurnished", "Rented private furnished", "Rented from employer") ~ "Rented",  # Rented
  UK_energy$tenure_dv %in% c("Housing assoc rented", "Local authority rent") ~ "Public Housing",  # Public Housing
  UK_energy$tenure_dv %in% c("missing", "Other") ~ NA_character_  # NA for missing/other
)
```

### Technology model for electricity 

Use estimates from academic literature for how much different technologies consume for electricity and gas. 
For vehicles, need to also look at switching from petrol to electric only. 

1. Baseline 2050 consumption
2. Elasticizes of Demand --- For Future 
3. Adjusted Consumption with technology uptake 
4. Combine for impacts 

Step 1. Panel data regression with individual (household) fixed effects to control for variation within households. 

UK: I use a panel data regression with within fixed-effects to model energy consumption with the dependent variable as electricity consumption in Mwh and independent variables on electricity prices, gas prices, income, household size, urban/ rural divide, owner/ rental status of home, and other socio-economic variables. Overall the model explains 17.92% of the variation. Electricity prices are highly correlated (-0.115)***, meaning that higher electricity prices are associated with lower consumption. Gas prices, hwoever, suggest that increases in gas lead to increases in electricity, potentially through fuel switching, electric heating, or other means. As households have more income, they tend to consume more electricity (0.791)*** and urban homes consume more electricity than rural homes (2.688)***. While this is true in general, effect diminishes with income and household size. Interestingly, results show that public housing tends to have higher consumption than homeonwers, who in turn consume less than renters. London consumes less than other retiongs and north west has higher consumption relative to the national average. Interestingly, deciles have a U-shape relationship, with lowest income consuming a bit more (potentially as a result of poor housing quality and lack of efficient technologies), middle having the largest reductions and uper income having a luxury effect with larger homes and more use. 

Sensititivye analysis tells me that fixed effect model is preferred to random effects (reject the null of the hausman test) and use fixed effect panel to model my electricity consumption to 2050. VIF multicollinearity tests are all below 10 (highest are weighted decile and income at 6, which are correlated)

```{r UK panel with fixed effects, echo = FALSE}
UK_energy$year <- as.numeric(as.character(UK_energy$year))
US_energy$year <- as.numeric(as.character(US_energy$year))
DEU_energy$year <- as.numeric(as.character(DEU_energy$year))
DEU_energy <- DEU_energy %>%
  mutate(across(c(hgtyp1hh, hgowner, weighted_decile), as.factor))

library(plm)
energy_data_us <- pdata.frame(US_energy, index =  "year")
energy_data_de <- pdata.frame(DEU_energy, index =  "year")
energy_data_uk <- pdata.frame(UK_energy, index =  "year")
energy_data_uk$weighted_decile <- as.factor(energy_data_uk$weighted_decile)
energy_data_us$weighted_decile <- as.factor(energy_data_us$weighted_decile)
energy_data_de$weighted_decile <- as.factor(energy_data_de$weighted_decile)

### Electricity #### 
model_uk_elec <- plm(elec_a_mwh ~ elec_price + gas_price + log(adjusted_income_2yr) + hhsize + 
                            urban_dv + age_dv + hsrooms + tenure_dv + gor_dv + weighted_decile + year,
                               data = energy_data_uk, model = "within", index = c("weighted_decile", "year"))

model_us_elec <- plm(elec_a_mwh ~elec_price + gas_price + log(adjusted_income_2yr) + hh_size + 
                       hh_rooms + hh_own + us_state + weighted_decile + year, 
                     data = energy_data_us, model = "within", index = 
                       c("weighted_decile", "year"))

model_de_elec <- plm(elec_a_mwh ~elec_price + gas_price + log(adjusted_income_2yr) + hgtyp1hh +
                       hgsize + hgowner + hgnuts1 + weighted_decile + year, 
                     data = energy_data_de, model = "within", index = 
                       c("weighted_decile", "year"))

summary(model_us_elec)
summary(model_uk_elec)
summary(model_de_elec)

model_uk_gas <- plm(gas_a_mwh ~ elec_price + gas_price + log(adjusted_income_2yr) + hhsize + 
                            urban_dv + age_dv + hsrooms + tenure_dv + gor_dv + weighted_decile + year,
                               data = energy_data_uk, model = "within", index = c("weighted_decile", "year"))

model_us_gas <- plm(gas_a_mwh ~elec_price + gas_price + log(adjusted_income_2yr) + hh_size + 
                       hh_rooms + hh_own + us_state + weighted_decile + year, 
                     data = energy_data_us, model = "within", index = 
                       c("weighted_decile", "year"))

model_de_gas<- plm(gas_a_mwh ~elec_price + gas_price + log(adjusted_income_2yr) + hgtyp1hh +
                       hgsize + hgowner + hgnuts1 + weighted_decile + year, 
                     data = energy_data_de, model = "within", index = 
                       c("weighted_decile", "year"))

summary(model_uk_gas)
summary(model_us_gas)
summary(model_de_gas)

# Fit random effects model
model_uk_elec_random <- plm(formula = elec_a_mwh ~ elec_price + gas_price + log(adjusted_income_2yr),
                    data = energy_data_uk, model = "random", index = c("hidp", "year"))
model_uk_gas_random <- plm(formula = gas_a_mwh ~ elec_price + gas_price + log(adjusted_income_2yr),
                    data = energy_data_uk, model = "random", index = c("hidp", "year"))

# Perform Hausman Test
hausman_test <- phtest(model_uk_gas, model_uk_gas_random)
hausman_test


### Sensitivity Testing ###

plot_residuals <- function(model) {
  std_residuals <- residuals(model) / sd(residuals(model))
  residuals_df <- data.frame(index = 1:length(std_residuals), residuals = std_residuals)
  
  ggplot(residuals_df, aes(x = index, y = residuals)) +
    geom_hline(yintercept = 0, color = "gray50") +
    geom_hline(yintercept = c(-3, 3), color = "red", linetype = "dashed") +
    geom_point(size = 0.5, alpha = 0.5) +
    theme_minimal() +
    labs(title = "Standardized Residuals Plot", x = "Observation Index", y = "Standardized Residuals")
}
 ### look at residuals for each model 
plot_residuals(model_uk_elec)
plot_residuals(model_uk_gas)
plot_residuals(model_us_elec)
plot_residuals(model_us_gas)
plot_residuals(model_de_gas)
plot_residuals(model_de_elec)

# Clustered standard errors for panel data models using plm
library(lmtest)
coeftest(model_uk_elec, vcov = vcovHC(model_uk_elec, type = "HC1", cluster = "group"))
coeftest(model_uk_gas, vcov = vcovHC(model_uk_gas, type = "HC1", cluster = "group"))

### check for multicollinearity 
# Create a standard linear model with the same variables for VIF checking -- good for UK, Germany has some collinearity for income. 
# for the US states are very correlated, which is expected as gas and electricity prices are at the state level. 
library(car)
lm_for_vif_uk <- lm(gas_a_mwh ~ elec_price + gas_price + log(adjusted_income_2yr) + 
                 hhsize + urban_dv + age_dv + hsrooms + tenure_dv + gor_dv + 
                 weighted_decile,
                 data = energy_data_uk)

lm_for_vif_de<- lm(gas_a_mwh ~elec_price + gas_price + log(adjusted_income_2yr) + hgtyp1hh +
                       hgsize + hgowner + hgnuts1 + weighted_decile, 
                     data = energy_data_de)

lm_for_vif_us <- lm(elec_a_mwh ~elec_price + gas_price + log(adjusted_income_2yr) + hh_size + 
                       hh_rooms + hh_own + us_state + weighted_decile, 
                     data = energy_data_us)
# Check VIF
vif(lm_for_vif_us)

```
Basic ARIMA price forecasting for electricity. Gas prices is ARIMA with bootstrap to help with volatility. These become input variables in my future predictions dataframe. 

### Projection for Price

Clean price data 
```{r electricity price forecast, echo = FALSE}
## ensure its in GBP per MWh
uk_gas <- uk_gas %>% 
  rename(gas_price = "Overall: Unit cost (Pence per kWh)") %>%
  mutate(gas_price = as.numeric(gas_price),
         gas_price = (gas_price / 100) * 1000)  # Convert to numeric in the same pipeline
uk_gas <- uk_gas[, c("year", "gor_dv", "gas_price")]


uk_elec <- uk_elec %>%
        rename(elec_price = "Overall: Unit cost (Pence per kWh)" ) %>%
  mutate(elec_price = as.numeric(elec_price),
         elec_price = (elec_price / 100) * 1000)
uk_elec <- uk_elec [, c("year", "gor_dv", "elec_price")]

## ensure its in $/MWh
us_elec <- seds[, c("year", "us_state", "elec_price"), with = FALSE]
us_elec$elec_price <- as.numeric(us_elec$elec_price)

us_gas <- seds[, c("year", "us_state", "gas_price"), with = FALSE]
us_gas$gas_price <- as.numeric(us_gas$gas_price)


de_gas <- nrg %>%
  group_by(year) %>%
  summarise(
    gas_price = mean(gas_price, na.rm = TRUE) * 1000
  )
de_gas$gas_price <- as.numeric(de_gas$gas_price)

de_elec <- nrg %>%
  group_by(year) %>%
  summarise(
    elec_price = mean(elec_price, na.rm = TRUE) * 1000
  )
de_elec$elec_price <- as.numeric(de_elec$elec_price)
```

```{r function for price projection, echo = FALSE}
# ---- Required Packages ----
if (!require(forecast)) install.packages("forecast")
if (!require(dplyr)) install.packages("dplyr")
library(forecast)
library(dplyr)
library(rugarch)

# ---- Clean Data Function ----
clean_data <- function(df, price_var, region_var = NULL) {
    df[[price_var]] <- as.numeric(df[[price_var]])  # Ensure price column is numeric
    
    # If region_var is provided, use it, otherwise clean by year
    if (!is.null(region_var)) {
        df_clean <- df[complete.cases(df[[price_var]], df[[region_var]], df$year), ]
    } else {
        df_clean <- df[complete.cases(df[[price_var]], df$year), ]  # Clean by year only if no region
    }
    
    return(df_clean)
}

forecast_arima <- function(df, price_var, region_var = NULL, years_ahead = 30, min_years = 5) {
    
    # If no region_var is provided or it's not in the dataframe, treat the whole data as one group
    if (is.null(region_var) || !region_var %in% colnames(df)) {
        # Aggregate by year for the entire dataset (no regional split)
        df_clean <- clean_data(df, price_var)
        
        # Aggregate by year, calculate average price
        region_data_agg <- aggregate(df_clean[[price_var]], 
                                     by = list(df_clean$year), 
                                     FUN = mean)
        colnames(region_data_agg) <- c("year", "avg_price")
        region_data_agg <- region_data_agg[order(region_data_agg$year), ]
        
        if (nrow(region_data_agg) >= min_years) {
            price_ts <- ts(region_data_agg$avg_price, start = min(region_data_agg$year), frequency = 1)
            arima_model <- auto.arima(price_ts, stepwise = FALSE, approximation = FALSE, trace = TRUE)
            forecast_results <- forecast(arima_model, h = years_ahead)
        } else {
            warning("Not enough data to forecast for Germany")
            forecast_results <- NULL
        }
    } else {
        # Otherwise, if region_var is provided, loop over regions (for UK, US, etc.)
        regions <- unique(df[[region_var]])
        forecast_results <- list()
        
        for (region in regions) {
            region_data <- df[df[[region_var]] == region, ]
            region_data_clean <- clean_data(region_data, price_var, "year")
            
            # Aggregate by year, calculate average price
            region_data_agg <- aggregate(region_data_clean[[price_var]], 
                                         by = list(region_data_clean$year), 
                                         FUN = mean)
            colnames(region_data_agg) <- c("year", "avg_price")
            region_data_agg <- region_data_agg[order(region_data_agg$year), ]
            
            if (nrow(region_data_agg) >= min_years) {
                price_ts <- ts(region_data_agg$avg_price, start = min(region_data_agg$year), frequency = 1)
                arima_model <- auto.arima(price_ts, stepwise = FALSE, approximation = FALSE, trace = TRUE)
                forecast_results[[as.character(region)]] <- forecast(arima_model, h = years_ahead)
            } else {
                warning(paste("Not enough data to forecast for region:", region))
            }
        }
    }
    
    return(forecast_results)
}

# ---- ARIMA + GARCH + Bootstrap Forecast (Gas) ----
forecast_arima_garch_bootstrap <- function(df, price_var, region_var = NULL, years_ahead = 30, min_years = 5, n_paths = 100) {

    # Ensure rugarch package is loaded
    if (!require("rugarch")) {
        install.packages("rugarch")
        library(rugarch)
    }

    # If no region_var is provided or it's not in the dataframe, treat the whole data as one group
    if (is.null(region_var) || !region_var %in% colnames(df)) {
        # Aggregate by year for the entire dataset (no regional split)
        df_clean <- clean_data(df, price_var)
        
        # Aggregate by year, calculate average price
        region_data_agg <- aggregate(df_clean[[price_var]], 
                                     by = list(df_clean$year), 
                                     FUN = mean)
        colnames(region_data_agg) <- c("year", "avg_price")
        region_data_agg <- region_data_agg[order(region_data_agg$year), ]
        
        if (nrow(region_data_agg) >= min_years) {
            price_ts <- ts(region_data_agg$avg_price, start = min(region_data_agg$year), frequency = 1)
            
            # Fit ARIMA model
            arima_model <- auto.arima(price_ts, stepwise = FALSE, approximation = FALSE, trace = TRUE)
            
            # Print out the ARIMA model selection (for debugging)
            print(arima_model)
            
            # Get residuals and fit GARCH model
            garch_spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
                                     mean.model = list(armaOrder = arima_model$arma[c(1, 6)], include.mean = TRUE),
                                     distribution.model = "norm")
            garch_fit <- ugarchfit(garch_spec, residuals(arima_model))
            
            # Generate forecast using ARIMA and GARCH
            arima_forecast <- forecast(arima_model, h = years_ahead)
            garch_forecast <- ugarchforecast(garch_fit, n.ahead = years_ahead)
            
            # Combine the forecasts: ARIMA mean and GARCH volatility
            forecast_results <- list(
                arima_forecast = arima_forecast,
                garch_forecast = garch_forecast
            )
        } else {
            warning("Not enough data to forecast for the whole dataset")
            forecast_results <- NULL
        }
    } else {
        # If region_var is provided, loop over regions (for UK, US, etc.)
        regions <- unique(df[[region_var]])
        forecast_results <- list()
        
        for (region in regions) {
            region_data <- df[df[[region_var]] == region, ]
            region_data_clean <- clean_data(region_data, price_var, "year")
            
            # Aggregate by year, calculate average price
            region_data_agg <- aggregate(region_data_clean[[price_var]], 
                                         by = list(region_data_clean$year), 
                                         FUN = mean)
            colnames(region_data_agg) <- c("year", "avg_price")
            region_data_agg <- region_data_agg[order(region_data_agg$year), ]
            
            if (nrow(region_data_agg) >= min_years) {
                price_ts <- ts(region_data_agg$avg_price, start = min(region_data_agg$year), frequency = 1)
                
                # Fit ARIMA model
                arima_model <- auto.arima(price_ts, stepwise = FALSE, approximation = FALSE, trace = TRUE)
                
                # Print out the ARIMA model selection (for debugging)
                print(arima_model)
                
                # Get residuals and fit GARCH model
                garch_spec <- ugarchspec(variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
                                         mean.model = list(armaOrder = arima_model$arma[c(1, 6)], include.mean = TRUE),
                                         distribution.model = "norm")
                garch_fit <- ugarchfit(garch_spec, residuals(arima_model))
                
                # Generate forecast using ARIMA and GARCH
                arima_forecast <- forecast(arima_model, h = years_ahead)
                garch_forecast <- ugarchforecast(garch_fit, n.ahead = years_ahead)
                
                # Store the forecast results for each region
                forecast_results[[as.character(region)]] <- list(
                    arima_forecast = arima_forecast,
                    garch_forecast = garch_forecast
                )
            } else {
                warning(paste("Not enough data to forecast for region:", region))
            }
        }
    }
    
    return(forecast_results)
}

        
  

# ---- Plot Function ----
plot_forecast <- function(forecast_list, title) {
    if (length(forecast_list) > 0) {
        for (region in names(forecast_list)) {
            plot(forecast_list[[region]],
                 main = paste(title, "for", region),
                 ylab = "Price (National Currency/MWh)",
                 xlab = "Year")
        }
    } else {
        print("No regions had enough data to forecast.")
    }
}


# ---- Forecasting ----
# Electricity (ARIMA Only)
forecasted_DE_elec_prices <- forecast_arima(de_elec, "elec_price", 30)
forecasted_UK_elec_prices <- forecast_arima(uk_elec, "elec_price", "gor_dv", 30)
forecasted_US_elec_prices <- forecast_arima(us_elec, "elec_price", "us_state", 30)
# Gas (ARIMA + Bootstrap)
forecasted_UK_gas_prices <- forecast_arima_garch_bootstrap(uk_gas, "gas_price", "gor_dv", 30)
forecasted_US_gas_prices <- forecast_arima_bootstrap(us_gas, "gas_price", "us_state", 30)
forecasted_DE_gas_prices <- forecast_arima_garch_bootstrap(de_gas, "gas_price", 30)
# ---- Plotting ----
plot_forecast(forecasted_DE_elec_prices, "Forecasted DE Electricity Prices")
plot_forecast(forecasted_DE_gas_prices, "Forecasted DE Gas Prices")
plot_forecast(forecasted_UK_elec_prices, "Forecasted UK Electricity Prices")
plot_forecast(forecasted_UK_gas_prices, "Forecasted UK Gas Prices")
plot_forecast(forecasted_US_gas_prices, "Forecasted US Gas Prices")

```

### Projection for Consumption

```{r future consumption, echo=FALSE}
### simple growth rate for coefficents that need it 
hhsize_growth_rate <- mean(diff(energy_data_uk$hhsize), na.rm = TRUE)
projected_hhsize <- rep(tail(energy_data_uk$hhsize, 1) * (1 + hhsize_growth_rate)^(1:29),
                        each = length(unique(energy_data_uk$weighted_decile)))
income_growth_rate <- mean(diff(log(energy_data_uk$adjusted_income_2yr)), na.rm = TRUE)
projected_income <- rep(tail(energy_data_uk$adjusted_income_2yr, 1) * (1 + income_growth_rate)^(1:29), 
                        each = length(unique(energy_data_uk$weighted_decile)))
```


```{r prepare data, include = FALSE}
DE_summary <- DE_summary %>%
  pivot_wider(names_from = parameter, values_from = estimate)
US_summary <- US_summary %>%
  pivot_wider(names_from = parameter, values_from = estimate)
UK_summary <- GB_summary %>%
  pivot_wider(names_from = parameter, values_from = estimate)



```
```{r technology impacts, echo=FALSE}
technology_impacts <- data.frame(
        technology = c("solar", "smartphone", "internet", "renovation", "fuel_switch"),
        elec_generation_mwh = c(2.5, 0.000, 0.000, 0.000, 0.000),   
        elec_consumption_mwh = c(0.030, 0.05, 0.180, -0.500, 3.500),
        gas_generation_mwh = c(0.000, 0.000, 0.000, 0.000, 0.000),  
        gas_consumption_mwh = c(0.000, 0.000, 0.000, -0.200, -3.500) 
) 

```

```{r Baseline Scenario function, include = FALSE}
# Energy projection function that works for multiple countries
project_energy_time <- function(consumption_data, tech_impacts, adoption_data, 
                                country_name = "DEU",
                                start_year = 2025, end_year = 2050,
                                n_simulations = 100) {
        
        # Ensure correct data types
        consumption_data <- consumption_data %>%
                mutate(weighted_decile = as.integer(weighted_decile))
        
        # Function to run a single projection with enhanced randomness
        single_projection <- function(seed) {
                set.seed(seed)
                
                adoption_params <- adoption_data %>%
                        mutate(
                                technology = tolower(technology),
                                weighted_decile = case_when(
                                        quintile == "Quintile 1" ~ list(1:2),
                                        quintile == "Quintile 2" ~ list(3:4),
                                        quintile == "Quintile 3" ~ list(5:6),
                                        quintile == "Quintile 4" ~ list(7:8),
                                        quintile == "Quintile 5" ~ list(9:10)
                                ),
                                # Enhanced randomness for more visible confidence intervals
                                country_max = pmin(1, pmax(0, country_max + rnorm(n(), 0, 0.08))),
                                growth_rate = growth_rate + rnorm(n(), 0, 0.02),
                                inflexion_year = inflexion_year + rnorm(n(), 0, 1.5)
                        ) %>%
                        unnest(weighted_decile) %>%
                        # Force weighted_decile to be integer
                        mutate(weighted_decile = as.integer(weighted_decile))
                
                # Get baseline for the last year
                last_year <- max(consumption_data$year)
                baseline <- consumption_data %>%
                        filter(year == last_year)
                
                # Create time projection
                years <- data.frame(projection_year = seq(start_year, end_year, by = 1))
                
                # Force identical types before join
                results <- baseline %>%
                        crossing(years) %>%
                        crossing(technology = unique(tech_impacts$technology)) %>%
                        # Force integer type before joining
                        mutate(weighted_decile = as.integer(weighted_decile)) %>%
                        left_join(adoption_params, by = c("weighted_decile", "technology")) %>%
                        left_join(tech_impacts, by = "technology") %>%
                        mutate(
                                # Calculate adoption and impacts
                                adoption_factor = country_max / (1 + exp(-growth_rate * (projection_year - inflexion_year))),
                                tech_gas_impact = gas_consumption_mwh * adoption_factor,
                                tech_elec_impact = elec_consumption_mwh * adoption_factor
                        )
                
                # Aggregate results
                aggregated_results <- results %>%
                        group_by(weighted_decile, projection_year) %>%
                        summarise(
                                projected_gas = first(baseline_gas) + sum(tech_gas_impact, na.rm = TRUE),
                                projected_elec = first(baseline_elec) + sum(tech_elec_impact, na.rm = TRUE),
                                
                                gas_price = first(gas_price),
                                elec_price = first(elec_price),
                                
                                annual_gas_bill = projected_gas * gas_price,
                                annual_elec_bill = projected_elec * elec_price,
                                total_annual_bill = annual_gas_bill + annual_elec_bill,
                                
                                baseline_gas = first(baseline_gas),
                                baseline_elec = first(baseline_elec),
                                baseline_annual_gas_bill = baseline_gas * gas_price,
                                baseline_annual_elec_bill = baseline_elec * elec_price,
                                baseline_total_annual_bill = baseline_annual_gas_bill + baseline_annual_elec_bill,
                                
                                total_adoption_factor = sum(adoption_factor, na.rm = TRUE),
                                .groups = 'drop'
                        ) %>%
                        # Add country identifier
                        mutate(country = country_name)
                
                return(aggregated_results)
        }
        
        # Run multiple simulations
        message(paste("Running", n_simulations, "simulations for", country_name, "..."))
        simulation_results <- list()
        for(i in 1:n_simulations) {
                if(i %% 10 == 0) message(sprintf("Completed %d of %d simulations", i, n_simulations))
                simulation_results[[i]] <- single_projection(i)
        }
        
        # Calculate confidence intervals
        combined_results <- bind_rows(simulation_results, .id = "simulation") %>%
                group_by(country, weighted_decile, projection_year) %>%
                summarise(
                        # Central values (using median for robustness)
                        projected_gas = median(projected_gas),
                        projected_elec = median(projected_elec),
                        total_annual_bill = median(total_annual_bill),
                        annual_gas_bill = median(annual_gas_bill),
                        annual_elec_bill = median(annual_elec_bill),
                        
                        # Confidence intervals - using wider interval for visibility
                        total_annual_bill_lower = quantile(total_annual_bill, 0.2),
                        total_annual_bill_upper = quantile(total_annual_bill, 0.8),
                        annual_gas_bill_lower = quantile(annual_gas_bill, 0.2),
                        annual_gas_bill_upper = quantile(annual_gas_bill, 0.8),
                        annual_elec_bill_lower = quantile(annual_elec_bill, 0.2),
                        annual_elec_bill_upper = quantile(annual_elec_bill, 0.8),
                        
                        # Also add confidence intervals for consumption
                        projected_gas_lower = quantile(projected_gas, 0.2),
                        projected_gas_upper = quantile(projected_gas, 0.8),
                        projected_elec_lower = quantile(projected_elec, 0.2),
                        projected_elec_upper = quantile(projected_elec, 0.8),
                        
                        # Baselines (should be the same across all simulations)
                        baseline_gas = first(baseline_gas),
                        baseline_elec = first(baseline_elec),
                        baseline_annual_gas_bill = first(baseline_annual_gas_bill),
                        baseline_annual_elec_bill = first(baseline_annual_elec_bill),
                        baseline_total_annual_bill = first(baseline_total_annual_bill),
                        
                        # Tracking adoption
                        total_adoption_factor = median(total_adoption_factor),
                        .groups = 'drop'
                )
        
        return(combined_results)
}

# Run projections for all three countries
deu_projections <- project_energy_time(
        consumption_data = DEU_consumption,
        tech_impacts = technology_impacts,
        adoption_data = DE_summary,
        country_name = "DEU",
        start_year = 2025,
        end_year = 2050,
        n_simulations = 100
)

uk_projections <- project_energy_time(
        consumption_data = UK_consumption,
        tech_impacts = technology_impacts,
        adoption_data = UK_summary,
        country_name = "UK",
        start_year = 2025,
        end_year = 2050,
        n_simulations = 100
)

us_projections <- project_energy_time(
        consumption_data = US_consumption,
        tech_impacts = technology_impacts,
        adoption_data = US_summary,
        country_name = "US",
        start_year = 2025,
        end_year = 2050,
        n_simulations = 100
)

# Function to prepare historical data in consistent format
prepare_historical_data <- function(consumption_data, country_name) {
        consumption_data %>%
                mutate(
                        country = country_name,
                        weighted_decile = as.integer(weighted_decile),
                        total_consumption = baseline_gas + baseline_elec,
                        annual_gas_bill = baseline_gas * gas_price,
                        annual_elec_bill = baseline_elec * elec_price,
                        total_annual_bill = annual_gas_bill + annual_elec_bill
                ) %>%
                rename(projection_year = year) %>%
                select(country, projection_year, weighted_decile,
                       baseline_gas, baseline_elec, total_consumption,
                       annual_gas_bill, annual_elec_bill, total_annual_bill)
}

# Prepare historical data for all countries
historical_deu <- prepare_historical_data(DEU_consumption, "DEU")
historical_uk <- prepare_historical_data(UK_consumption, "UK")
historical_us <- prepare_historical_data(US_consumption, "US")

# Combine historical data
historical_combined <- bind_rows(historical_deu, historical_uk, historical_us) %>%
        pivot_longer(
                cols = c(baseline_gas, baseline_elec, total_consumption,
                         annual_gas_bill, annual_elec_bill, total_annual_bill),
                names_to = "metric",
                values_to = "value"
        ) %>%
        mutate(
                type = case_when(
                        metric %in% c("baseline_gas", "baseline_elec", "total_consumption") ~ "Consumption (MWh)",
                        TRUE ~ "Annual Bills (€)"
                ),
                source = case_when(
                        metric %in% c("baseline_gas", "annual_gas_bill") ~ "Gas",
                        metric %in% c("baseline_elec", "annual_elec_bill") ~ "Electricity",
                        TRUE ~ "Total"
                )
        )

# Prepare projection data in same format
prepare_projection_data <- function(projection_data) {
        # Calculate total consumption
        projection_data %>%
                mutate(
                        projected_total = projected_gas + projected_elec,
                        projected_total_lower = projected_gas_lower + projected_elec_lower,
                        projected_total_upper = projected_gas_upper + projected_elec_upper
                ) %>%
                # Reshape for visualization
                pivot_longer(
                        cols = c(
                                # Consumption columns
                                projected_gas, projected_elec, projected_total,
                                projected_gas_lower, projected_elec_lower, projected_total_lower,
                                projected_gas_upper, projected_elec_upper, projected_total_upper,
                                # Bill columns
                                annual_gas_bill, annual_elec_bill, total_annual_bill,
                                annual_gas_bill_lower, annual_elec_bill_lower, total_annual_bill_lower,
                                annual_gas_bill_upper, annual_elec_bill_upper, total_annual_bill_upper
                        ),
                        names_to = "metric",
                        values_to = "value"
                ) %>%
                mutate(
                        # Categorize by type
                        type = case_when(
                                grepl("projected", metric) ~ "Consumption (MWh)",
                                grepl("bill", metric) ~ "Annual Bills (€)"
                        ),
                        # Categorize by source
                        source = case_when(
                                grepl("gas", metric) ~ "Gas",
                                grepl("elec", metric) ~ "Electricity",
                                TRUE ~ "Total"
                        ),
                        # Identify central values vs confidence intervals
                        ci_type = case_when(
                                grepl("lower", metric) ~ "lower",
                                grepl("upper", metric) ~ "upper",
                                TRUE ~ "central"
                        )
                )
}

# Process all projection data
projection_combined <- bind_rows(
        prepare_projection_data(deu_projections),
        prepare_projection_data(uk_projections),
        prepare_projection_data(us_projections)
)

# Function to create visualization for a single country
create_country_vis <- function(country_code) {
        # Filter data for the selected country
        country_historical <- historical_combined %>% 
                filter(country == country_code)
        
        country_projections <- projection_combined %>% 
                filter(country == country_code)
        
        # Create the plot
        ggplot() +
                # Historical data
                geom_line(data = country_historical,
                          aes(x = projection_year, 
                              y = value,
                              color = factor(weighted_decile),
                              group = weighted_decile),
                          size = 1) +
                # Transition point
                geom_point(data = country_historical %>% 
                                   filter(projection_year == max(projection_year)),
                           aes(x = projection_year,
                               y = value,
                               color = factor(weighted_decile)),
                           size = 2) +
                # Projections with confidence intervals
                geom_ribbon(data = country_projections %>% 
                                    filter(ci_type != "central") %>%
                                    pivot_wider(names_from = ci_type, values_from = value, values_fill = NA),
                            aes(x = projection_year,
                                ymin = lower,
                                ymax = upper,
                                fill = factor(weighted_decile),
                                group = weighted_decile),
                            alpha = 0.3,  # Increased for visibility
                            color = NA) +  # Remove ribbon border
                # Projected central lines
                geom_line(data = country_projections %>% 
                                  filter(ci_type == "central"),
                          aes(x = projection_year,
                              y = value,
                              color = factor(weighted_decile),
                              group = weighted_decile),
                          linetype = "dashed",
                          size = 1) +
                # Transition line
                geom_vline(xintercept = 2025, 
                           linetype = "dashed", 
                           color = "gray50") +
                # Faceting
                facet_grid(type ~ source, scales = "free_y") +
                # Customize appearance
                labs(
                        title = paste(country_code, "- Historical and Projected Energy Consumption and Bills"),
                        subtitle = "Solid lines = historical data, Dashed lines = projections with 20-80% confidence intervals",
                        x = "Year",
                        y = NULL,
                        color = "Income Decile",
                        fill = "Income Decile"
                ) +
                theme_minimal() +
                scale_color_viridis_d() +
                scale_fill_viridis_d() +
                scale_y_continuous(limits = c(0, NA), expand = c(0, 0)) +
                theme(
                        legend.position = "right",
                        plot.title = element_text(face = "bold"),
                        plot.subtitle = element_text(size = 9),
                        strip.text = element_text(face = "bold"),
                        panel.spacing = unit(2, "lines"),
                        axis.text = element_text(margin = margin(t = 5, r = 5))
                )
}

# Create visualizations for each country
deu_vis <- create_country_vis("DEU")
uk_vis <- create_country_vis("UK") 
us_vis <- create_country_vis("US")

# Display the visualizations
print(deu_vis)
print(uk_vis)
print(us_vis)

# Compare results across countries - summary statistics
country_comparison <- bind_rows(
        deu_projections %>% 
                filter(projection_year %in% c(2025, 2050)) %>%
                select(country, weighted_decile, projection_year, projected_gas, projected_elec, 
                       annual_gas_bill, annual_elec_bill, total_annual_bill),
        uk_projections %>% 
                filter(projection_year %in% c(2025, 2050)) %>%
                select(country, weighted_decile, projection_year, projected_gas, projected_elec, 
                       annual_gas_bill, annual_elec_bill, total_annual_bill),
        us_projections %>% 
                filter(projection_year %in% c(2025, 2050)) %>%
                select(country, weighted_decile, projection_year, projected_gas, projected_elec, 
                       annual_gas_bill, annual_elec_bill, total_annual_bill)
) %>%
        group_by(country, weighted_decile) %>%
        pivot_wider(
                names_from = projection_year,
                values_from = c(projected_gas, projected_elec, annual_gas_bill, annual_elec_bill, total_annual_bill)
        ) %>%
        mutate(
                gas_change_pct = (projected_gas_2050 - projected_gas_2025) / projected_gas_2025 * 100,
                elec_change_pct = (projected_elec_2050 - projected_elec_2025) / projected_elec_2025 * 100,
                bill_change_pct = (total_annual_bill_2050 - total_annual_bill_2025) / total_annual_bill_2025 * 100
        ) %>%
        arrange(country, weighted_decile)

print("Cross-Country Comparison:")
print(country_comparison)


# Save projection results
saveRDS(deu_projections, "deu_projections.rds")
saveRDS(uk_projections, "uk_projections.rds")
saveRDS(us_projections, "us_projections.rds")

# Save visualizations
ggsave("deu_energy_projection.png", deu_vis, width = 12, height = 8, dpi = 300)
ggsave("uk_energy_projection.png", uk_vis, width = 12, height = 8, dpi = 300)
ggsave("us_energy_projection.png", us_vis, width = 12, height = 8, dpi = 300)

# Save summary statistics
write.csv(country_comparison, "country_comparison.csv", row.names = FALSE)
```

```{r Baseline Scenario Improved, echo = FALSE}
# Fix issue with data
DEU_consumption <- DEU_consumption %>% 
        filter(!year == 2010)
UK_consumption <- UK_consumption %>%
        filter(!year == 2009)

# Enhanced energy projection function with improved baseline methodologies
project_energy_time_enhanced <- function(consumption_data, tech_impacts, adoption_data, 
                                         household_projections = NULL,
                                         country_name = "DEU",
                                         start_year = 2025, end_year = 2050,
                                         n_simulations = 100,
                                         regional_variation = TRUE,
                                         include_demographics = TRUE) {
        
        # Ensure correct data types
        consumption_data <- consumption_data %>%
                mutate(weighted_decile = as.integer(weighted_decile))
        
        # 1. Smoothing approach with conversion to numeric
        historical_trend <- consumption_data %>%
                group_by(weighted_decile) %>%
                arrange(year) %>%
                mutate(
                        # Convert the smoothed output to numeric vectors
                        smoothed_elec = as.numeric(stats::smooth(baseline_elec, kind = "3R")),
                        smoothed_gas = as.numeric(stats::smooth(baseline_gas, kind = "3R"))
                ) %>%
                ungroup()
        
        # Get last year of historical data for baseline
        last_year <- max(consumption_data$year)
        baseline <- historical_trend %>%
                filter(year == last_year) %>%
                # Use smoothed values for projection base
                mutate(
                        baseline_elec = smoothed_elec,
                        baseline_gas = smoothed_gas
                )
        
        # Function to run a single projection with enhanced modeling
        single_projection <- function(seed) {
                set.seed(seed)
                
                # 2. Enhanced adoption parameters with regional variation
                adoption_params <- adoption_data %>%
                        mutate(
                                technology = tolower(technology),
                                weighted_decile = case_when(
                                        quintile == "Quintile 1" ~ list(1:2),
                                        quintile == "Quintile 2" ~ list(3:4),
                                        quintile == "Quintile 3" ~ list(5:6),
                                        quintile == "Quintile 4" ~ list(7:8),
                                        quintile == "Quintile 5" ~ list(9:10)
                                ),
                                # Add regional variation if enabled
                                region_factor = if(regional_variation) {
                                        # Simulate regional differences in adoption
                                        sample(c(0.85, 0.95, 1.0, 1.05, 1.15), n(), replace = TRUE,
                                               prob = c(0.2, 0.3, 0.3, 0.15, 0.05))
                                } else {
                                        1.0  # No regional variation
                                },
                                # Income-based adoption modification
                                income_factor = case_when(
                                        quintile == "Quintile 1" ~ 0.8,  # Lower income adopts slower
                                        quintile == "Quintile 2" ~ 0.9,
                                        quintile == "Quintile 3" ~ 1.0,  # Middle income is baseline
                                        quintile == "Quintile 4" ~ 1.1,
                                        quintile == "Quintile 5" ~ 1.2   # Higher income adopts faster
                                ),
                                # Technology saturation - not all households will adopt
                                # Different max adoption rates by technology and income
                                adjusted_max = pmin(1, country_max * income_factor * region_factor),
                                
                                # Add random variation with income-dependent noise
                                # Higher income has more variation in adoption patterns
                                variation_factor = case_when(
                                        quintile == "Quintile 1" ~ 0.04,
                                        quintile == "Quintile 2" ~ 0.05,
                                        quintile == "Quintile 3" ~ 0.06,
                                        quintile == "Quintile 4" ~ 0.07,
                                        quintile == "Quintile 5" ~ 0.08
                                ),
                                
                                # Apply variations
                                country_max = pmin(1, pmax(0, adjusted_max + rnorm(n(), 0, variation_factor))),
                                growth_rate = growth_rate * income_factor + rnorm(n(), 0, 0.02),
                                inflexion_year = inflexion_year - (log(income_factor) * 5) + rnorm(n(), 0, 1.5)
                        ) %>%
                        unnest(weighted_decile) %>%
                        # Force weighted_decile to be integer
                        mutate(weighted_decile = as.integer(weighted_decile))
                
                # Create time projection
                years <- data.frame(projection_year = seq(start_year, end_year, by = 1))
                
                # 3. Add demographic effects if enabled
                if(include_demographics && !is.null(household_projections)) {
                        # Join with household size projections to adjust consumption
                        years <- years %>%
                                left_join(household_projections, by = c("projection_year" = "year")) %>%
                                mutate(
                                        # Default to no change if missing
                                        household_size_factor = if_else(is.na(household_size_factor), 1.0, household_size_factor)
                                )
                } else {
                        # No demographic adjustment
                        years <- years %>%
                                mutate(household_size_factor = 1.0)
                }
                
                # Calculate technology impacts with enhanced modeling
                results <- baseline %>%
                        crossing(years) %>%
                        crossing(technology = unique(tech_impacts$technology)) %>%
                        # Force integer type before joining
                        mutate(weighted_decile = as.integer(weighted_decile)) %>%
                        left_join(adoption_params, by = c("weighted_decile", "technology")) %>%
                        left_join(tech_impacts, by = "technology") %>%
                        group_by(weighted_decile, projection_year) %>%
                        # 4. Track technology effects separately
                        mutate(
                                year_in_projection = projection_year - min(projection_year),
                                random_walk = cumsum(rnorm(n(), 0, 0.01)),  # Add autocorrelation
                                                                raw_adoption = country_max / (1 + exp(-growth_rate * (projection_year - inflexion_year))),
                                
                                adoption_factor = pmin(country_max, pmax(0, raw_adoption + random_walk)),
                                
                                # 5. Natural efficiency improvements in baseline technologies
                                baseline_efficiency_factor = 1 - (0.003 * year_in_projection),  # ~0.3% annual efficiency gain
                                
                                tech_gas_impact = gas_consumption_mwh * adoption_factor,
                                tech_elec_impact = elec_consumption_mwh * adoption_factor,
                                
                                tech_id = paste(technology, row_number(), sep = "_")
                        ) %>%
                        ungroup()
                
                aggregated_results <- results %>%
                        group_by(weighted_decile, projection_year) %>%
                        summarise(
                                # Apply household size factor and baseline efficiency to consumption
                                household_size_factor = first(household_size_factor),
                                baseline_efficiency_factor = first(baseline_efficiency_factor),
                                
                                projected_gas = (first(baseline_gas) * baseline_efficiency_factor * household_size_factor) +
                                        sum(tech_gas_impact, na.rm = TRUE),
                                projected_elec = (first(baseline_elec) * baseline_efficiency_factor * household_size_factor) +
                                        sum(tech_elec_impact, na.rm = TRUE),
                                
                                projected_gas = pmax(0, projected_gas),
                                projected_elec = pmax(0, projected_elec),
                                
                                gas_price = first(gas_price),
                                elec_price = first(elec_price),
                                
                                annual_gas_bill = projected_gas * gas_price,
                                annual_elec_bill = projected_elec * elec_price,
                                total_annual_bill = annual_gas_bill + annual_elec_bill,
                                
                                baseline_gas = first(baseline_gas),
                                baseline_elec = first(baseline_elec),
                                baseline_annual_gas_bill = baseline_gas * gas_price,
                                baseline_annual_elec_bill = baseline_elec * elec_price,
                                baseline_total_annual_bill = baseline_annual_gas_bill + baseline_annual_elec_bill,
                                
                                total_adoption_factor = sum(adoption_factor, na.rm = TRUE),
                                unique_technologies = n_distinct(tech_id),
                                .groups = 'drop'
                        ) %>%
                        # Add country identifier
                        mutate(country = country_name)
                
                return(aggregated_results)
        }
        
        message(paste("Running", n_simulations, "simulations for", country_name, "..."))
        simulation_results <- list()
        pb <- txtProgressBar(min = 0, max = n_simulations, style = 3)  # Better progress bar
        for(i in 1:n_simulations) {
                simulation_results[[i]] <- single_projection(i)
                setTxtProgressBar(pb, i)
        }
        close(pb)
        
        combined_results <- bind_rows(simulation_results, .id = "simulation") %>%
                group_by(country, weighted_decile, projection_year) %>%
                summarise(
                        projected_gas = median(projected_gas),
                        projected_elec = median(projected_elec),
                        total_annual_bill = median(total_annual_bill),
                        annual_gas_bill = median(annual_gas_bill),
                        annual_elec_bill = median(annual_elec_bill),

                        total_annual_bill_lower = quantile(total_annual_bill, 0.1),
                        total_annual_bill_upper = quantile(total_annual_bill, 0.9),
                        annual_gas_bill_lower = quantile(annual_gas_bill, 0.1),
                        annual_gas_bill_upper = quantile(annual_gas_bill, 0.9),
                        annual_elec_bill_lower = quantile(annual_elec_bill, 0.1),
                        annual_elec_bill_upper = quantile(annual_elec_bill, 0.9),
                        
                        projected_gas_lower = quantile(projected_gas, 0.1),
                        projected_gas_upper = quantile(projected_gas, 0.9),
                        projected_elec_lower = quantile(projected_elec, 0.1),
                        projected_elec_upper = quantile(projected_elec, 0.9),
                        
                        projected_gas_sd = sd(projected_gas),
                        projected_elec_sd = sd(projected_elec),
                        total_annual_bill_sd = sd(total_annual_bill),
                        
                        baseline_gas = first(baseline_gas),
                        baseline_elec = first(baseline_elec),
                        baseline_annual_gas_bill = first(baseline_annual_gas_bill),
                        baseline_annual_elec_bill = first(baseline_annual_elec_bill),
                        baseline_total_annual_bill = first(baseline_total_annual_bill),
                        
                        total_adoption_factor = median(total_adoption_factor),
                        avg_technologies = median(unique_technologies),
                        .groups = 'drop'
                )
        
        model_metrics <- data.frame(
                country = country_name,
                n_simulations = n_simulations,
                include_demographics = include_demographics,
                regional_variation = regional_variation,
                convergence_metric = sd(sapply(simulation_results, function(x) mean(x$total_annual_bill))),
                simulation_variability = mean(combined_results$total_annual_bill_sd),
                simulation_time = Sys.time()
        )
        
        list(
                projections = combined_results,
                model_metrics = model_metrics
        )
}

create_household_projection <- function(country, start_year = 2025, end_year = 2050) {
        annual_change <- case_when(
                country == "DEU" ~ -0.004,  # Declining household size
                country == "UK" ~ -0.003,   # Slow decline
                country == "US" ~ -0.002    # Very slow decline
        )
        
        tibble(
                country = country,
                year = start_year:end_year,
                household_size_factor = 1 + annual_change * (year - start_year)
        )
}

run_all_countries <- function() {
        # Prepare country-specific household projections
        deu_households <- create_household_projection("DEU", 2025, 2050)
        uk_households <- create_household_projection("UK", 2025, 2050)
        us_households <- create_household_projection("US", 2025, 2050)
        
        # Germany
        message("Running Germany projections...")
        deu_enhanced <- project_energy_time_enhanced(
                consumption_data = DEU_consumption,
                tech_impacts = technology_impacts,
                adoption_data = DE_summary,
                household_projections = deu_households,
                country_name = "DEU",
                start_year = 2025,
                end_year = 2050,
                n_simulations = 100,
                regional_variation = TRUE,
                include_demographics = TRUE
        )
        
        # UK
        message("Running UK projections...")
        uk_enhanced <- project_energy_time_enhanced(
                consumption_data = UK_consumption,
                tech_impacts = technology_impacts,
                adoption_data = UK_summary,
                household_projections = uk_households,
                country_name = "UK",
                start_year = 2025,
                end_year = 2050,
                n_simulations = 100,
                regional_variation = TRUE,
                include_demographics = TRUE
        )
        
        # US
        message("Running US projections...")
        us_enhanced <- project_energy_time_enhanced(
                consumption_data = US_consumption,
                tech_impacts = technology_impacts,
                adoption_data = US_summary,
                household_projections = us_households,
                country_name = "US",
                start_year = 2025,
                end_year = 2050,
                n_simulations = 100,
                regional_variation = TRUE,
                include_demographics = TRUE
        )
        
        return(list(
                DEU = deu_enhanced,
                UK = uk_enhanced,
                US = us_enhanced
        ))
}

# Run projections for all countries
all_country_results <- run_all_countries()

# Extract projection data for academic presentation
deu_projections <- all_country_results$DEU$projections
uk_projections <- all_country_results$UK$projections
us_projections <- all_country_results$US$projections

# Combine all results
all_projections <- bind_rows(
        deu_projections,
        uk_projections,
        us_projections
)

# Function to create enhanced visualization
create_enhanced_vis <- function(projections, historical_data, country_code) {
        # Prepare historical data
        historical <- historical_data %>%
                filter(country == country_code) %>%
                mutate(
                        type = case_when(
                                metric %in% c("baseline_gas", "baseline_elec", "total_consumption") ~ "Consumption (MWh)",
                                TRUE ~ "Annual Bills (€)"
                        ),
                        source = case_when(
                                metric %in% c("baseline_gas", "annual_gas_bill") ~ "Gas",
                                metric %in% c("baseline_elec", "annual_elec_bill") ~ "Electricity",
                                TRUE ~ "Total"
                        )
                )
        
        # Prepare projection data
        projection <- projections %>%
                filter(country == country_code) %>%
                mutate(
                        projected_total = projected_gas + projected_elec,
                        projected_total_lower = projected_gas_lower + projected_elec_lower,
                        projected_total_upper = projected_gas_upper + projected_elec_upper
                ) %>%
                pivot_longer(
                        cols = c(
                                # Consumption columns
                                projected_gas, projected_elec, projected_total,
                                projected_gas_lower, projected_elec_lower, projected_total_lower,
                                projected_gas_upper, projected_elec_upper, projected_total_upper,
                                # Bill columns
                                annual_gas_bill, annual_elec_bill, total_annual_bill,
                                annual_gas_bill_lower, annual_elec_bill_lower, total_annual_bill_lower,
                                annual_gas_bill_upper, annual_elec_bill_upper, total_annual_bill_upper
                        ),
                        names_to = "metric",
                        values_to = "value"
                ) %>%
                mutate(
                        type = case_when(
                                grepl("projected", metric) ~ "Consumption (MWh)",
                                TRUE ~ "Annual Bills (National Currency)"
                        ),
                        source = case_when(
                                grepl("gas", metric) ~ "Gas",
                                grepl("elec", metric) ~ "Electricity",
                                TRUE ~ "Total"
                        ),
                        ci_type = case_when(
                                grepl("lower", metric) ~ "lower",
                                grepl("upper", metric) ~ "upper",
                                TRUE ~ "central"
                        )
                )
        
        # Create visualization
        ggplot() +
                # Historical data
                geom_line(data = historical,
                          aes(x = projection_year, 
                              y = value,
                              color = factor(weighted_decile),
                              group = weighted_decile),
                          size = 1) +
                # Transition point
                geom_point(data = historical %>% 
                                   filter(projection_year == max(projection_year)),
                           aes(x = projection_year,
                               y = value,
                               color = factor(weighted_decile)),
                           size = 2) +
                # Projections with confidence intervals
                geom_ribbon(data = projection %>% 
                                    filter(ci_type != "central") %>%
                                    pivot_wider(names_from = ci_type, values_from = value, values_fill = NA),
                            aes(x = projection_year,
                                ymin = lower,
                                ymax = upper,
                                fill = factor(weighted_decile),
                                group = weighted_decile),
                            alpha = 0.3, 
                            color = NA) +
                # Projected central lines
                geom_line(data = projection %>% 
                                  filter(ci_type == "central"),
                          aes(x = projection_year,
                              y = value,
                              color = factor(weighted_decile),
                              group = weighted_decile),
                          linetype = "dashed",
                          size = 1) +
                # Transition line
                geom_vline(xintercept = 2025, 
                           linetype = "dashed", 
                           color = "gray50") +
                # Faceting
                facet_grid(type ~ source, scales = "free_y") +
                # Customize appearance
                labs(
                        title = paste(country_code, "- Enhanced Baseline Projection"),
                        subtitle = "Includes demographic effects, regional variation, and natural efficiency improvements",
                        x = "Year",
                        y = NULL,
                        color = "Income Decile",
                        fill = "Income Decile"
                ) +
                theme_minimal() +
                scale_color_viridis_d() +
                scale_fill_viridis_d() +
                scale_y_continuous(limits = c(0, NA), expand = c(0, 0)) +
                theme(
                        legend.position = "right",
                        plot.title = element_text(face = "bold"),
                        plot.subtitle = element_text(size = 9),
                        strip.text = element_text(face = "bold"),
                        panel.spacing = unit(2, "lines")
                )
}

deu_enhanced_vis <- create_enhanced_vis(
        deu_projections, 
        historical_combined, 
        "DEU"
)

us_enhanced_vis <- create_enhanced_vis(
        us_projections,  
        historical_combined, 
        "US"
)

uk_enhanced_vis <- create_enhanced_vis(
        uk_projections,  # Fixed: using the correct variable
        historical_combined, 
        "UK"
)

# Display the enhanced visualization
print(deu_enhanced_vis)
print(uk_enhanced_vis)
print(us_enhanced_vis)

# Sensitivity analysis - run multiple scenarios
run_sensitivity_analysis <- function(country_data, tech_data, adoption_data, country_name) {
        # Define scenarios
        scenarios <- expand.grid(
                regional_variation = c(TRUE, FALSE),
                demographics = c(TRUE, FALSE),
                efficiency_improvement = c("High", "Medium", "Low")
        )
        
        # Adjust household projections based on scenarios
        household_proj <- case_when(
                country_name == "DEU" ~ deu_households,
                country_name == "UK" ~ uk_households,
                country_name == "US" ~ us_households
        )
        
        # Run all scenarios
        scenario_results <- list()
        for(i in 1:nrow(scenarios)) {
                # Get scenario parameters
                regional <- scenarios$regional_variation[i]
                demographics <- scenarios$demographics[i]
                efficiency <- scenarios$efficiency_improvement[i]
                
                # Run projection with this scenario
                message(paste("Running scenario", i, "of", nrow(scenarios)))
                
                # Adjust tech impacts based on efficiency scenario
                modified_tech <- tech_data %>%
                        mutate(
                                efficiency_factor = case_when(
                                        efficiency == "High" ~ 1.2,
                                        efficiency == "Medium" ~ 1.0,
                                        efficiency == "Low" ~ 0.8
                                ),
                                gas_consumption_mwh = gas_consumption_mwh * efficiency_factor,
                                elec_consumption_mwh = elec_consumption_mwh * efficiency_factor
                        )
                
                # Run with reduced simulations for speed
                result <- project_energy_time_enhanced(
                        consumption_data = country_data,
                        tech_impacts = modified_tech,
                        adoption_data = adoption_data,
                        household_projections = household_proj,
                        country_name = country_name,
                        start_year = 2025,
                        end_year = 2050,
                        n_simulations = 50,  # Reduced for sensitivity analysis
                        regional_variation = regional,
                        include_demographics = demographics
                )
                
                # Add scenario details
                result$projections$scenario_id <- i
                result$projections$regional_variation <- regional
                result$projections$demographics <- demographics
                result$projections$efficiency <- efficiency
                
                scenario_results[[i]] <- result$projections
        }
        
        # Combine all scenarios
        all_scenarios <- bind_rows(scenario_results)
        
        # Return with scenario metadata
        list(
                results = all_scenarios,
                scenarios = scenarios
        )
}

# Run sensitivity analysis for Germany
deu_sensitivity <- run_sensitivity_analysis(
        DEU_consumption, 
        technology_impacts, 
        DE_summary, 
        "DEU"
)

uk_sensitivity <- run_sensitivity_analysis(
        UK_consumption, 
        technology_impacts, 
        UK_summary, 
        "UK"
)

us_sensitivity <- run_sensitivity_analysis(
        US_consumption, 
        technology_impacts, 
        US_summary, 
        "US"
)

# Summarize sensitivity analysis
sensitivity_summary <- deu_sensitivity$results %>%
        filter(projection_year == 2050) %>%
        group_by(scenario_id, regional_variation, demographics, efficiency, weighted_decile) %>%
        summarise(
                total_consumption = projected_gas + projected_elec,
                total_bill = total_annual_bill,
                .groups = 'drop'
        ) %>%
        group_by(scenario_id, regional_variation, demographics, efficiency) %>%
        summarise(
                avg_consumption = mean(total_consumption),
                avg_bill = mean(total_bill),
                .groups = 'drop'
        ) %>%
        arrange(desc(avg_consumption))

# Display sensitivity results
print("Sensitivity Analysis Results:")
print(sensitivity_summary)

saveRDS(all_country_results, "all_country_enhanced_projections.rds")

# Or save individual country results
saveRDS(all_country_results$DEU, "deu_enhanced_projections.rds")
saveRDS(all_country_results$US, "us_enhanced_projections.rds")
saveRDS(all_country_results$UK, "uk_enhanced_projections.rds")


saveRDS(deu_sensitivity, "deu_sensitivity_analysis.rds")
ggsave("deu_enhanced_projection.png", deu_enhanced_vis, width = 12, height = 8, dpi = 300)
ggsave("uk_enhanced_projection.png", uk_enhanced_vis, width = 12, height = 8, dpi = 300)
ggsave("us_enhanced_projection.png", us_enhanced_vis, width = 12, height = 8, dpi = 300)


#-----------------------------------
# ADDED: Academic Presentation Code
#-----------------------------------

# Load required packages for academic tables
library(fixest)
library(modelsummary)
library(kableExtra)

# 1. Create regression-style tables with fixed effects
regression_data <- all_projections %>%
        mutate(
                # Create decile categories
                income_group = case_when(
                        weighted_decile <= 3 ~ "Low",
                        weighted_decile <= 7 ~ "Middle",
                        TRUE ~ "High"
                ),
                # Create period indicators
                period = case_when(
                        projection_year <= 2030 ~ "Near-term (2025-2030)",
                        projection_year <= 2040 ~ "Mid-term (2031-2040)",
                        TRUE ~ "Long-term (2041-2050)"
                ),
                # Calculate changes from baseline
                elec_change = projected_elec - baseline_elec,
                gas_change = projected_gas - baseline_gas,
                total_energy_change = elec_change + gas_change,
                bill_change = total_annual_bill - baseline_total_annual_bill,
                pct_bill_change = (total_annual_bill / baseline_total_annual_bill - 1) * 100
        )

# Run regression models
models <- list(
        "Electricity" = feols(projected_elec ~ income_group + period | country, regression_data),
        "Gas" = feols(projected_gas ~ income_group + period | country, regression_data),
        "Total Bill" = feols(total_annual_bill ~ income_group + period | country, regression_data),
        "% Bill Change" = feols(pct_bill_change ~ income_group + period | country, regression_data)
)

# Create regression table (Model 1)
table1 <- modelsummary(models, 
                       stars = TRUE,
                       title = "Energy Consumption and Bills by Income Group and Period",
                       coef_map = c(
                               "income_groupLow" = "Low Income",
                               "income_groupMiddle" = "Middle Income",
                               "income_groupHigh" = "High Income (ref.)",
                               "periodNear-term (2025-2030)" = "2025-2030",
                               "periodMid-term (2031-2040)" = "2031-2040",
                               "periodLong-term (2041-2050)" = "2041-2050"
                       ),
                       gof_omit = "IC|Log|F|RMSE")

# 2. Panel data fixed effects model (Model 2)
panel_models <- list(
        "Gas\nConsumption" = feols(projected_gas ~ period:income_group | country + weighted_decile, regression_data),
        "Electricity\nConsumption" = feols(projected_elec ~ period:income_group | country + weighted_decile, regression_data),
        "Energy\nBill" = feols(total_annual_bill ~ period:income_group | country + weighted_decile, regression_data),
        "Bill\nChange %" = feols(pct_bill_change ~ period:income_group | country + weighted_decile, regression_data)
)

table2 <- modelsummary(panel_models, 
                       stars = TRUE,
                       title = "Energy Consumption and Bills - Panel Data with Fixed Effects",
                       coef_omit = "Int",
                       gof_omit = "IC|Log|F|RMSE")

# 3. Point estimates with confidence intervals (most common in energy economics)
point_estimates <- all_projections %>%
        filter(projection_year %in% c(2030, 2040, 2050)) %>%
        group_by(country, weighted_decile, projection_year) %>%
        summarise(
                gas_consumption = projected_gas,
                gas_consumption_lower = projected_gas_lower,
                gas_consumption_upper = projected_gas_upper,
                
                elec_consumption = projected_elec,
                elec_consumption_lower = projected_elec_lower,
                elec_consumption_upper = projected_elec_upper,
                
                total_bill = total_annual_bill,
                total_bill_lower = total_annual_bill_lower,
                total_bill_upper = total_annual_bill_upper,
                
                # Get baseline values for comparisons
                baseline_gas = baseline_gas,
                baseline_elec = baseline_elec,
                baseline_bill = baseline_total_annual_bill,
                
                # Calculate percentage changes
                gas_change_pct = (projected_gas / baseline_gas - 1) * 100,
                elec_change_pct = (projected_elec / baseline_elec - 1) * 100,
                bill_change_pct = (total_annual_bill / baseline_total_annual_bill - 1) * 100,
                
                .groups = "drop"
        ) %>%
        # Group by income group for table
        mutate(
                income_group = case_when(
                        weighted_decile <= 3 ~ "Low Income",
                        weighted_decile <= 7 ~ "Middle Income",
                        TRUE ~ "High Income"
                )
        ) %>%
        arrange(country, projection_year, weighted_decile)

# Create point estimate table for select years
point_table <- point_estimates %>%
        filter(weighted_decile %in% c(1, 5, 10)) %>%  # Representative deciles
        mutate(
                # Format values with confidence intervals
                gas_value = sprintf("%.1f [%.1f, %.1f]", 
                                    gas_consumption, gas_consumption_lower, gas_consumption_upper),
                elec_value = sprintf("%.1f [%.1f, %.1f]", 
                                     elec_consumption, elec_consumption_lower, elec_consumption_upper),
                bill_value = sprintf("%.0f [%.0f, %.0f]", 
                                     total_bill, total_bill_lower, total_bill_upper),
                
                # Format percentage changes
                gas_change = sprintf("%.1f%%", gas_change_pct),
                elec_change = sprintf("%.1f%%", elec_change_pct),
                bill_change = sprintf("%.1f%%", bill_change_pct)
        ) %>%
        select(country, projection_year, income_group, weighted_decile,
               gas_value, gas_change, elec_value, elec_change, bill_value, bill_change) %>%
        # Make the table more readable
        rename(
                "Country" = country,
                "Year" = projection_year,
                "Income Group" = income_group,
                "Decile" = weighted_decile,
                "Gas (MWh) [95% CI]" = gas_value,
                "Gas Change" = gas_change,
                "Electricity (MWh) [95% CI]" = elec_value,
                "Electricity Change" = elec_change,
                "Annual Bill (€) [95% CI]" = bill_value,
                "Bill Change" = bill_change
        )

# Create table with kableExtra styling
kable(point_table, "html", caption = "Table 1: Projected Energy Consumption and Bills with Confidence Intervals") %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
        column_spec(1, bold = TRUE) %>%
        collapse_rows(columns = 1:3, valign = "middle") %>%
        add_header_above(c(" " = 4, "Gas" = 2, "Electricity" = 2, "Energy Bill" = 2))

# 4. Decomposition analysis (common in energy economics papers)
decomposition <- all_projections %>%
        filter(projection_year %in% c(2025, 2030, 2040, 2050)) %>%
        group_by(country, weighted_decile) %>%
        mutate(
                # Calculate changes between periods
                total_consumption = projected_gas + projected_elec,
                baseline_total = baseline_gas + baseline_elec,
                
                # Decompose changes
                total_change = total_consumption - first(total_consumption),
                gas_contribution = projected_gas - first(projected_gas),
                elec_contribution = projected_elec - first(projected_elec),
                
                # Calculate percentage contributions
                gas_share = gas_contribution / total_change * 100,
                elec_share = elec_contribution / total_change * 100
        ) %>%
        filter(projection_year > 2025) %>%  # Remove base year
        ungroup()

# Create decomposition table
decomp_table <- decomposition %>%
        group_by(country, projection_year) %>%
        summarise(
                avg_total_change = mean(total_change, na.rm = TRUE),
                avg_gas_contribution = mean(gas_contribution, na.rm = TRUE),
                avg_elec_contribution = mean(elec_contribution, na.rm = TRUE),
                avg_gas_share = mean(gas_share, na.rm = TRUE),
                avg_elec_share = mean(elec_share, na.rm = TRUE),
                
                # By income group
                low_total_change = mean(total_change[weighted_decile <= 3], na.rm = TRUE),
                mid_total_change = mean(total_change[weighted_decile > 3 & weighted_decile <= 7], na.rm = TRUE),
                high_total_change = mean(total_change[weighted_decile > 7], na.rm = TRUE),
                
                .groups = "drop"
        ) %>%
        arrange(country, projection_year)

# Format decomposition table
decomp_formatted <- decomp_table %>%
        mutate(
                # Format values
                total_change = sprintf("%.2f", avg_total_change),
                gas_contribution = sprintf("%.2f (%.1f%%)", avg_gas_contribution, avg_gas_share),
                elec_contribution = sprintf("%.2f (%.1f%%)", avg_elec_contribution, avg_elec_share),
                
                # Format by income group
                low_income = sprintf("%.2f", low_total_change),
                mid_income = sprintf("%.2f", mid_total_change),
                high_income = sprintf("%.2f", high_total_change)
        ) %>%
        select(country, projection_year, total_change, gas_contribution, elec_contribution,
               low_income, mid_income, high_income) %>%
        rename(
                "Country" = country,
                "Year" = projection_year,
                "Total Change (MWh)" = total_change,
                "Gas Contribution" = gas_contribution,
                "Electricity Contribution" = elec_contribution,
                "Low Income" = low_income,
                "Middle Income" = mid_income,
                "High Income" = high_income
        )

# Create formatted table
kable(decomp_formatted, "html", caption = "Table 2: Decomposition of Energy Consumption Changes") %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
        column_spec(1, bold = TRUE) %>%
        collapse_rows(columns = 1, valign = "middle") %>%
        add_header_above(c(" " = 2, "Overall Change Decomposition" = 3, "Change by Income Group" = 3))

# 5. Distributional impact analysis (for energy policy papers)
distributional <- all_projections %>%
        filter(projection_year %in% c(2025, 2050)) %>%
        select(country, weighted_decile, projection_year, total_annual_bill, baseline_total_annual_bill) %>%
        pivot_wider(
                names_from = projection_year,
                values_from = c(total_annual_bill, baseline_total_annual_bill)
        ) %>%
        mutate(
                # Calculate absolute and percentage changes
                absolute_change = total_annual_bill_2050 - total_annual_bill_2025,
                percentage_change = (total_annual_bill_2050 / total_annual_bill_2025 - 1) * 100,
                
                # Calculate share of income (using decile as proxy)
                income_proxy = 11 - weighted_decile,  # Invert decile as income proxy
                bill_share_2025 = total_annual_bill_2025 / income_proxy,
                bill_share_2050 = total_annual_bill_2050 / income_proxy,
                share_change = bill_share_2050 - bill_share_2025
        ) %>%
        arrange(country, weighted_decile)

# Calculate distributional indices
gini_coef <- function(x) {
        x <- sort(x)
        n <- length(x)
        return(sum(x * 1:n) / (n * sum(x)) - (n + 1) / (2 * n))
}

distributional_indices <- distributional %>%
        group_by(country) %>%
        summarise(
                # Gini coefficient for bills in 2025 and 2050
                gini_2025 = gini_coef(total_annual_bill_2025),
                gini_2050 = gini_coef(total_annual_bill_2050),
                gini_change = gini_2050 - gini_2025,
                
                # Ratio of top to bottom decile
                top_bottom_2025 = total_annual_bill_2025[weighted_decile == 10] / 
                        total_annual_bill_2025[weighted_decile == 1],
                top_bottom_2050 = total_annual_bill_2050[weighted_decile == 10] / 
                        total_annual_bill_2050[weighted_decile == 1],
                top_bottom_change = top_bottom_2050 - top_bottom_2025,
                
                # Average change percentage
                avg_pct_change = mean(percentage_change),
                
                # Low income vs high income change
                low_income_pct = mean(percentage_change[weighted_decile <= 3]),
                high_income_pct = mean(percentage_change[weighted_decile >= 8]),
                
                .groups = "drop"
        )

# Format distributional indices table
dist_formatted <- distributional_indices %>%
        mutate(
                # Format values
                gini_2025 = sprintf("%.3f", gini_2025),
                gini_2050 = sprintf("%.3f", gini_2050),
                gini_change = sprintf("%.3f", gini_change),
                
                top_bottom_2025 = sprintf("%.2f", top_bottom_2025),
                top_bottom_2050 = sprintf("%.2f", top_bottom_2050),
                top_bottom_change = sprintf("%.2f", top_bottom_change),
                
                avg_pct_change = sprintf("%.1f%%", avg_pct_change),
                low_income_pct = sprintf("%.1f%%", low_income_pct),
                high_income_pct = sprintf("%.1f%%", high_income_pct)
        ) %>%
        rename(
                "Country" = country,
                "Gini 2025" = gini_2025,
                "Gini 2050" = gini_2050,
                "Gini Change" = gini_change,
                "Top/Bottom 2025" = top_bottom_2025,
                "Top/Bottom 2050" = top_bottom_2050,
                "Top/Bottom Change" = top_bottom_change,
                "Average Change" = avg_pct_change,
                "Low Income Change" = low_income_pct,
                "High Income Change" = high_income_pct
        )

# Create formatted table
kable(dist_formatted, "html", caption = "Table 3: Distributional Impact Analysis") %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
        column_spec(1, bold = TRUE) %>%
        add_header_above(c(" " = 1, "Gini Coefficient" = 3, "Top-to-Bottom Ratio" = 3, "Percentage Change" = 3))

# Save all results for future use
saveRDS(all_country_results, "all_country_enhanced_projections.rds")
saveRDS(point_estimates, "point_estimates.rds")
saveRDS(decomposition, "energy_decomposition.rds")
saveRDS(distributional, "distributional_analysis.rds")
saveRDS(distributional_indices, "distributional_indices.rds")

# Print summary of available tables
cat("Academic tables created:\n")
cat("1. Regression-style table with fixed effects\n")
cat("2. Panel data fixed effects model\n")
cat("3. Point estimates with confidence intervals\n")
cat("4. Decomposition analysis of energy changes\n")
cat("5. Distributional impact analysis\n")

```

```{r export to Overleaf, include = FALSE}
library(knitr)
library(kableExtra)

# Function to convert an RDS table to a LaTeX file
convert_rds_to_latex <- function(rds_file, caption = NULL, digits = 2) {
  data <- readRDS(rds_file)
    base_name <- tools::file_path_sans_ext(basename(rds_file))
    latex_table <- kable(data, format = "latex", booktabs = TRUE, 
                       caption = caption, digits = digits) %>%
    kable_styling(latex_options = c("striped", "scale_down", "hold_position"))
  
  output_file <- paste0(base_name, ".tex")
  write(latex_table, output_file)
  
  cat("Converted", rds_file, "to", output_file, "\n")
}

# Convert all your tables
convert_rds_to_latex("all_country_enhanced_projections.rds", 
                     caption = "Enhanced Country Projections")
convert_rds_to_latex("point_estimates.rds", 
                     caption = "Point Estimates")
convert_rds_to_latex("energy_decomposition.rds", 
                     caption = "Energy Decomposition")
convert_rds_to_latex("distributional_analysis.rds", 
                     caption = "Distributional Analysis")
convert_rds_to_latex("distributional_indices.rds", 
                     caption = "Distributional Indices")


library(knitr)
library(kableExtra)

# Load your distributional analysis data
distributional <- readRDS("distributional_analysis.rds")

# Create just the tabular content without the table environment
latex_content <- kable(distributional, format = "latex", booktabs = TRUE,
                    caption = NULL) %>%  # Important: no caption here
    kable_styling(latex_options = c("striped"))

# Save without table environment 
write(latex_content, "distributional_analysis_content.tex")

# If you want the full table with environment to be directly included
full_table <- kable(distributional, format = "latex", booktabs = TRUE,
                   caption = "Distributional Analysis") %>%
    kable_styling(latex_options = c("striped"))
write(full_table, "distributional_analysis_table.tex")
```


## Scenario 3: Front-loaded adoption

```{r frontloaded adoption scenario, echo = FALSE}
# Function to modify adoption data for front-loaded scenario
create_frontloaded_adoption <- function(adoption_data, shift_years = 3, growth_boost = 1.5) {
  
  frontloaded_adoption <- adoption_data %>%
    mutate(
      # Move inflexion point earlier (shift adoption curve to the left)
      inflexion_year = inflexion_year - shift_years,
      
      # Increase growth rate to make adoption happen faster initially
      growth_rate = growth_rate * growth_boost,
      
      # Keep the maximum adoption levels the same
      # This ensures we're comparing speed of adoption, not total adoption
      country_max = country_max
    )
  
  return(frontloaded_adoption)
}

# Function to run frontloaded scenario
run_frontloaded_scenario <- function() {
  # Create frontloaded adoption data
  deu_frontloaded <- create_frontloaded_adoption(DE_summary)
  uk_frontloaded <- create_frontloaded_adoption(UK_summary)
  us_frontloaded <- create_frontloaded_adoption(US_summary)
  
  # Prepare country-specific household projections (reusing from original code)
  deu_households <- create_household_projection("DEU", 2025, 2050)
  uk_households <- create_household_projection("UK", 2025, 2050)
  us_households <- create_household_projection("US", 2025, 2050)
  
  # Germany frontloaded scenario
  message("Running Germany frontloaded projections...")
  deu_frontloaded_results <- project_energy_time_enhanced(
    consumption_data = DEU_consumption,
    tech_impacts = technology_impacts,
    adoption_data = deu_frontloaded,
    household_projections = deu_households,
    country_name = "DEU",
    start_year = 2025,
    end_year = 2050,
    n_simulations = 100,
    regional_variation = TRUE,
    include_demographics = TRUE
  )
  
  # UK frontloaded scenario
  message("Running UK frontloaded projections...")
  uk_frontloaded_results <- project_energy_time_enhanced(
    consumption_data = UK_consumption,
    tech_impacts = technology_impacts,
    adoption_data = uk_frontloaded,
    household_projections = uk_households,
    country_name = "UK",
    start_year = 2025,
    end_year = 2050,
    n_simulations = 100,
    regional_variation = TRUE,
    include_demographics = TRUE
  )
  
  # US frontloaded scenario
  message("Running US frontloaded projections...")
  us_frontloaded_results <- project_energy_time_enhanced(
    consumption_data = US_consumption,
    tech_impacts = technology_impacts,
    adoption_data = us_frontloaded,
    household_projections = us_households,
    country_name = "US",
    start_year = 2025,
    end_year = 2050,
    n_simulations = 100,
    regional_variation = TRUE,
    include_demographics = TRUE
  )
  
  # Return combined results
  return(list(
    DEU = deu_frontloaded_results,
    UK = uk_frontloaded_results,
    US = us_frontloaded_results
  ))
}

# Run frontloaded scenario
frontloaded_results <- run_frontloaded_scenario()
```

## Scenario 4: Back-laoded adoption 

I now have estimates of how much these additional technologies would cost each decile to 2050. Now it is time to run a better forecast model that assumes earlier or later adoption trends of technologies. I want to look at how earlier or later adoption affects. 

```{r slow adoption, echo=FALSE}
# Function to modify adoption data for back-loaded scenario
create_backloaded_adoption <- function(adoption_data, delay_years = 4, growth_reduction = 0.7) {
  
  backloaded_adoption <- adoption_data %>%
    mutate(
      # Move inflexion point later (shift adoption curve to the right)
      inflexion_year = inflexion_year + delay_years,
      
      # Decrease growth rate to make adoption happen slower initially
      growth_rate = growth_rate * growth_reduction,
      
      # Keep the maximum adoption levels similar
      country_max = country_max
    )
  
  return(backloaded_adoption)
}

run_backloaded_scenario <- function() {
  # Create backloaded adoption data
  deu_backloaded <- create_backloaded_adoption(DE_summary)
  uk_backloaded <- create_backloaded_adoption(UK_summary)
  us_backloaded <- create_backloaded_adoption(US_summary)
  
  # Prepare country-specific household projections (reusing from original code)
  deu_households <- create_household_projection("DEU", 2025, 2050)
  uk_households <- create_household_projection("UK", 2025, 2050)
  us_households <- create_household_projection("US", 2025, 2050)
  
  message("Running Germany backloaded projections...")
  deu_backloaded_results <- project_energy_time_enhanced(
    consumption_data = DEU_consumption,
    tech_impacts = technology_impacts,
    adoption_data = deu_backloaded,
    household_projections = deu_households,
    country_name = "DEU",
    start_year = 2025,
    end_year = 2050,
    n_simulations = 100,
    regional_variation = TRUE,
    include_demographics = TRUE
  )
  
  message("Running UK backloaded projections...")
  uk_backloaded_results <- project_energy_time_enhanced(
    consumption_data = UK_consumption,
    tech_impacts = technology_impacts,
    adoption_data = uk_backloaded,
    household_projections = uk_households,
    country_name = "UK",
    start_year = 2025,
    end_year = 2050,
    n_simulations = 100,
    regional_variation = TRUE,
    include_demographics = TRUE
  )
  
  message("Running US backloaded projections...")
  us_backloaded_results <- project_energy_time_enhanced(
    consumption_data = US_consumption,
    tech_impacts = technology_impacts,
    adoption_data = us_backloaded,
    household_projections = us_households,
    country_name = "US",
    start_year = 2025,
    end_year = 2050,
    n_simulations = 100,
    regional_variation = TRUE,
    include_demographics = TRUE
  )
    return(list(
    DEU = deu_backloaded_results,
    UK = uk_backloaded_results,
    US = us_backloaded_results
  ))
}


backloaded_results <- run_backloaded_scenario()

```

```{r compare scenarios, echo = FALSE}


compare_adoption_scenarios <- function(baseline, frontloaded, backloaded, country_code) {
  # Extract projection data for the specified country
  baseline_data <- baseline[[country_code]]$projections %>% 
    filter(weighted_decile %in% c(1, 5, 10)) %>%
    mutate(scenario = "Baseline")
    
  frontloaded_data <- frontloaded[[country_code]]$projections %>%
    filter(weighted_decile %in% c(1, 5, 10)) %>%
    mutate(scenario = "Front-loaded")
    
  backloaded_data <- backloaded[[country_code]]$projections %>%
    filter(weighted_decile %in% c(1, 5, 10)) %>%
    mutate(scenario = "Back-loaded")
  
  # Combine data
  combined_data <- bind_rows(baseline_data, frontloaded_data, backloaded_data)
  
  # Map deciles to meaningful labels
  combined_data <- combined_data %>%
    mutate(income_group = case_when(
      weighted_decile == 1 ~ "Low Income (Q1)",
      weighted_decile == 5 ~ "Middle Income (Q5)",
      weighted_decile == 10 ~ "High Income (Q10)",
      TRUE ~ paste("Decile", weighted_decile)
    ))
  
  # Create visualization comparing scenarios
  ggplot(combined_data, 
         aes(x = projection_year, 
             y = total_annual_bill, 
             color = scenario,
             linetype = scenario)) +
    geom_line(size = 1.2) +
    facet_wrap(~income_group, scales = "free_y") +
    labs(
      title = paste(country_code, "- Impact of Adoption Timing by Income Group"),
      subtitle = "Comparison of baseline, front-loaded, and back-loaded technology adoption",
      x = "Year",
      y = "Annual Energy Bill(log scaled)",
      color = "Scenario",
      linetype = "Scenario"
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      strip.text = element_text(face = "bold", size = 11),
      plot.title = element_text(face = "bold"),
      plot.subtitle = element_text(size = 9)
    ) +
    scale_color_brewer(palette = "Set1") +
    scale_x_continuous(breaks = seq(2025, 2050, 5)) + 
scale_y_continuous(trans = "log1p", breaks = pretty_breaks())}

deu_scenario_vis <- compare_adoption_scenarios(
  all_country_results,
  frontloaded_results,
  backloaded_results,
  "DEU"
)

uk_scenario_vis <- compare_adoption_scenarios(
  all_country_results,
  frontloaded_results,
  backloaded_results,
  "UK"
)

us_scenario_vis <- compare_adoption_scenarios(
  all_country_results,
  frontloaded_results,
  backloaded_results,
  "US"
)
ggsave("deu_income_scenarios.pdf", deu_scenario_vis, width = 10, height = 7)
ggsave("uk_income_scenarios.pdf", uk_scenario_vis, width = 10, height = 7)
ggsave("us_income_scenarios.pdf", us_scenario_vis, width = 10, height = 7)





# Run the relative impacts comparison visualization
compare_relative_impacts <- function(baseline, frontloaded, backloaded, country_code) {
  # Calculate % differences from baseline for each scenario
  baseline_data <- baseline[[country_code]]$projections %>% 
    filter(weighted_decile %in% c(1, 5, 10)) %>%
    select(weighted_decile, projection_year, baseline_bill = total_annual_bill)
  
  # Process frontloaded data with relative changes
  frontloaded_data <- frontloaded[[country_code]]$projections %>%
    filter(weighted_decile %in% c(1, 5, 10)) %>%
    select(weighted_decile, projection_year, frontloaded_bill = total_annual_bill) %>%
    inner_join(baseline_data, by = c("weighted_decile", "projection_year")) %>%
    mutate(
      scenario = "Front-loaded",
      pct_change = (frontloaded_bill / baseline_bill - 1) * 100
    )
  
  # Process backloaded data with relative changes
  backloaded_data <- backloaded[[country_code]]$projections %>%
    filter(weighted_decile %in% c(1, 5, 10)) %>%
    select(weighted_decile, projection_year, backloaded_bill = total_annual_bill) %>%
    inner_join(baseline_data, by = c("weighted_decile", "projection_year")) %>%
    mutate(
      scenario = "Back-loaded",
      pct_change = (backloaded_bill / baseline_bill - 1) * 100
    )
  
  # Combine and prepare for plotting
  combined_data <- bind_rows(frontloaded_data, backloaded_data) %>%
    mutate(income_group = case_when(
      weighted_decile == 1 ~ "Low Income (Q1)",
      weighted_decile == 5 ~ "Middle Income (Q5)",
      weighted_decile == 10 ~ "High Income (Q10)",
      TRUE ~ paste("Decile", weighted_decile)
    ))
  
  # Create visualization
  ggplot(combined_data, 
         aes(x = projection_year, 
             y = pct_change, 
             color = scenario,
             linetype = scenario)) +
    geom_line(size = 1.2) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
    facet_wrap(~income_group) +
    labs(
      title = paste(country_code, "- Relative Impact of Different Adoption Scenarios"),
      subtitle = "Percentage change in annual energy bills compared to baseline scenario",
      x = "Year",
      y = "% Change from Baseline",
      color = "Scenario",
      linetype = "Scenario"
    ) +
    theme_minimal() +
    scale_color_brewer(palette = "Set1")
}

deu_relative_impacts <- compare_relative_impacts(
  all_country_results,
  frontloaded_results,
  backloaded_results,
  "DEU"
)

uk_relative_impacts <- compare_relative_impacts(
  all_country_results,
  frontloaded_results,
  backloaded_results,
  "UK"
)

us_relative_impacts <- compare_relative_impacts(
  all_country_results,
  frontloaded_results,
  backloaded_results,
  "US"
)


# Also save as PDF for vector graphics (better quality in LaTeX)
ggsave("deu_relative_impacts.pdf", deu_relative_impacts, width = 10, height = 7)
ggsave("uk_relative_impacts.pdf", uk_relative_impacts, width = 10, height = 7)
ggsave("us_relative_impacts.pdf", us_relative_impacts, width = 10, height = 7)
```